---
title: "All figures for manuscript"
author: Tormey Reimer
date: last-modified
---

```{r setup}
#| code-summary: Load required packages and set preferences
#| cache: false
#| warning: false
#| echo: false

library(stringr)
library(units)
library(targets)
library(tarchetypes)
library(macrogrow)
library(ozmaps)
library(arrow)
library(gganimate)
library(magrittr)
library(ggplot2)
library(dplyr)
library(terra)
library(tidyterra)
library(here)
library(lubridate)
library(tidyr)
library(cowplot)
library(purrr)
library(patchwork)
library(conflicted)
library(gridExtra)
library(grid)

conflicts_prefer(dplyr::select, dplyr::filter)

source(here("R_scripts/00_functions.R"))
source(here("R_scripts/00_global_displays.R"))
```

```{r special-display}
#| code-summary: Special display parameters for manuscript figures

prettyplot_MS <- function() {
  theme_classic() +
    theme(text = element_text(family = "serif", size = 12, colour = "black"),
          legend.position = "none",
          axis.title.y = element_text(vjust = 1.5),
          axis.title.x = element_text(vjust = 1.5),
          legend.title = element_blank())
}
env_plot_MS <- function() {
  prettyplot_MS() + theme(aspect.ratio = 0.55)
}

rm.x <- function(){
  theme(axis.text.x = element_blank(), axis.ticks.x = element_blank(), axis.title.x = element_blank())
  }
rm.y <- function(){
  theme(axis.text.y = element_blank(), axis.ticks.y = element_blank(), axis.title.y = element_blank())
  }

tlabel <- function(letter, x = 0, y = Inf) {
  annotate("text", label = letter, 
      x = x, y = y, hjust = 0.25, vjust = 1, 
      size = 4, family = "serif", fontface = "bold"
      )
}

rm(prettyplot, env_plot) # Removes previous display defaults
```

# Figure 1: Map of all cells

```{r cell-coords}
#| code-summary: |
#|   Load in all the BARRA-R2 cells being used, with their coordinates.

cell_coords <- find_read(envi_data, "R2_cell_coords") %>% 
  mutate(state = factor(state, levels = states_ord)) %>% 
  dplyr::select(-layer)
```

```{r fig-1}
#| code-summary: Code to generate plot
#| fig-height: 5.5
#| fig-cap: Map showing the area included this analysis, with colours indicating the different states. The ~60 km coastline buffer includes 13,207 cells approximately 12×12 km each.

cell_coords %>% 
  ggplot(aes(x = longitude, y = latitude, fill = state, label = cell_no)) +
  geom_raster() +
  geom_sf(data = ozmap_data(data = "states"), inherit.aes = F) +
  coord_sf(xlim = c(bounds[["AUS"]]["lonmin"], bounds[["AUS"]]["lonmax"]),
           ylim = c(bounds[["AUS"]]["latmin"], bounds[["AUS"]]["latmax"])) +
  # coord_sf(xlim = c(149.5, 151.5),
  #          ylim = c(-39, -37)) +
  # geom_text(size = 2.25) +
  scale_x_continuous(breaks = seq(100, 160, 5)) +
  scale_y_continuous(breaks = seq(-5, -45, -5)) +
  scale_fill_manual(values = states_pal) +
  labs(x = "Longitude", y = "Latitude") +
  prettyplot_MS() +
  theme(legend.position = "none")
```

# Figure 2: Map of N sources

```{r N-data sources locations}
#| code-summary: Load all nitrogen sources and locations

refstation_locations <- here() %>% 
  file.path("data", "nitrogen", "refstation_locations.parquet") %>% 
  arrow::read_parquet() %>% 
  rename(name = StationName) %>% 
  dplyr::select(-code)

outfall_locations <- here() %>% 
  file.path("data_raw", "national-outfall-database", "data-output", "outflow_site_locations.parquet") %>% 
  arrow::read_parquet() %>% 
  rename(lon = longitude,
         lat = latitude) %>% 
  dplyr::select(-state)

N_locations <- rbind(
  outfall_locations %>% mutate(type = "outfall"),
  refstation_locations %>% mutate(type = "refstation")
) %>% 
  mutate(name = as.character(name),
         state = case_when(
    type == "refstation" ~ NA,
    lon > bounds[["SAU"]]["lonmin"] & lon <= bounds[["SAU"]]["lonmax"] & 
      lat > bounds[["SAU"]]["latmin"] & lat <= bounds[["SAU"]]["latmax"]  ~ "SAU",
    lon > bounds[["QLD"]]["lonmin"] & lon <= bounds[["QLD"]]["lonmax"] & 
      lat > bounds[["QLD"]]["latmin"] & lat <= bounds[["QLD"]]["latmax"]  ~ "QLD",
    lon > bounds[["WAS"]]["lonmin"] & lon <= bounds[["WAS"]]["lonmax"] & 
      lat > bounds[["WAS"]]["latmin"] & lat <= bounds[["WAS"]]["latmax"]  ~ "WAS",
    lon > bounds[["WAN"]]["lonmin"] & lon <= bounds[["WAN"]]["lonmax"] & 
      lat > bounds[["WAN"]]["latmin"] & lat <= bounds[["WAN"]]["latmax"]  ~ "WAN",
    lon > bounds[["VIC"]]["lonmin"] & lon <= bounds[["VIC"]]["lonmax"] & 
      lat > bounds[["VIC"]]["latmin"] & lat <= bounds[["VIC"]]["latmax"]  ~ "VIC",
    lon > bounds[["NSW"]]["lonmin"] & lon <= bounds[["NSW"]]["lonmax"] & 
      lat > bounds[["NSW"]]["latmin"] & lat <= bounds[["NSW"]]["latmax"]  ~ "NSW",
    lon > bounds[["NTE"]]["lonmin"] & lon <= bounds[["NTE"]]["lonmax"] & 
      lat > bounds[["NTE"]]["latmin"] & lat <= bounds[["NTE"]]["latmax"]  ~ "NTE",
    lon > bounds[["TAS"]]["lonmin"] & lon <= bounds[["TAS"]]["lonmax"] & 
      lat > bounds[["TAS"]]["latmin"] & lat <= bounds[["TAS"]]["latmax"]  ~ "TAS",
    TRUE ~ NA
  ),
  state = factor(state, levels = states_ord)) %>% 
  relocate(state, .before = name) %>% 
  rename(latitude = lat,
         longitude = lon)

refstation_locations <- filter(N_locations, type == "refstation")
outfall_locations <- filter(N_locations, type == "outfall")
```

```{r fig-2}
#| code-summary: Code to generate plot
#| fig-height: 5.5
#| fig-cap: Map showing the locations of recorded outfalls (points) and the IMOS reference stations (black diamonds) used in this analysis. Outfall data are coloured by state.

ggplot(N_locations, aes(x = longitude, y = latitude, color = state, shape = type)) +
  geom_sf(data = ozmap_data(data = "states"), inherit.aes = F) +
  coord_sf(xlim = c(bounds[["AUS"]]["lonmin"], bounds[["AUS"]]["lonmax"]),
           ylim = c(bounds[["AUS"]]["latmin"], bounds[["AUS"]]["latmax"])) +
  geom_point(size = 3.5) +
  scale_shape_manual(values = c(20, 18)) +
  scale_color_manual(values = states_pal, na.value = "black") +
  scale_x_continuous(breaks = seq(100, 160, 5)) +
  scale_y_continuous(breaks = seq(-5, -45, -5)) +
  labs(x = "Longitude", y = "Latitude") +
  prettyplot_MS() +
  theme(legend.position = "none")
```

# Figure 3: Sensitivity environmental conditions

```{r sens-conditions}
#| code-summary: Get data for sensitivity environmental conditions

culture_depths <- tar_read(culture_depths, store = spec_store)
factors <- tar_read(factors, store = spec_store)

state_input_timeseries <- envi_data %>% 
  list.files(full.names = T) %>% 
  str_subset("state_input_timeseries") %>% 
  purrr::map_dfr(arrow::read_parquet) %>% 
  filter(yday <= 365)

sens_cond <- file.path(spec_data, "sensitivity_conditions.parquet") %>% 
  arrow::read_parquet()
```

```{r fig-3}
#| code-summary: Code to generate plot
#| fig-height: 7
#| fig-cap: Input curves for the model parameter sensitivity runs. Environmental inputs included high (solid) and low (dashed) temperatures (A), surface irradiance (B), water velocity (C), salinity (D), nitrate concentration (E) and ammonium concentration (F).

p1 <- sens_cond %>% 
  dplyr::filter(t <= 365) %>% 
  ggplot(aes(x = t, y = T_input, linetype = as.factor(T_level))) +
  geom_line(linewidth = 0.75) +
  geom_line(data = state_input_timeseries, aes(x = yday, y = T_input_mean, colour = state), linetype = "dotted", inherit.aes = F) +
  scale_y_continuous(breaks = seq(10, 35, 5), limits = c(10, 33.5)) +
  labs(y = expression("Temperature ("*degree*"C)")) +
  prettyplot_MS()

p2 <- sens_cond %>% 
  dplyr::filter(t <= 365 & T_level == 1) %>% 
  ggplot(aes(x = t, y = I_input)) +
  geom_line(linewidth = 0.75) +
  geom_line(data = state_input_timeseries, aes(x = yday, y = I_input_mean, colour = state), linetype = "dotted", inherit.aes = F) +
  scale_y_continuous(breaks = seq(0, 700, 100), limits = c(0, 660)) +
  labs(y = expression("Irradiance ("*mu*"mol m"^-2*" s"^-1*")")) +
  prettyplot_MS()

p3 <- sens_cond %>% 
  dplyr::filter(t <= 365 & T_level == 1) %>% 
  ggplot(aes(x = t, y = U_input)) +
  geom_line(linewidth = 0.75) +
  geom_line(data = state_input_timeseries, aes(x = yday, y = UV_input_mean, colour = state), linetype = "dotted", inherit.aes = F) +
  scale_y_continuous(breaks = seq(0, 0.6, 0.1), limits = c(0, 0.575)) +
  labs(y = expression("Water velocity (m"^-1*" s"^-1*")")) +
  prettyplot_MS()

p4 <- sens_cond %>% 
  dplyr::filter(t <= 365 & T_level == 1) %>% 
  ggplot(aes(x = t, y = S_input)) +
  geom_line(linewidth = 0.75) +
  geom_line(data = state_input_timeseries, aes(x = yday, y = S_input_mean, colour = state), linetype = "dotted", inherit.aes = F) +
  scale_y_continuous(breaks = seq(33.1, 36.8, 0.3), limits = c(34, 36.11)) +
  labs(y = expression("Salinity (g L"^-1*")")) +
  prettyplot_MS()

pdata1 <- sens_cond %>% 
  dplyr::filter(t <= 365 & T_level == 1) %>% 
  mutate(Ni_input = set_units(Ni_input, "mg m-3") %>% set_units("umol L-1") %>% drop_units(),
         Am_input = set_units(Am_input, "mg m-3") %>% set_units("umol L-1") %>% drop_units())
pdata2 <- state_input_timeseries %>% 
  mutate(Ni_input_mean = set_units(Ni_input_mean, "mg m-3") %>% set_units("umol L-1") %>% drop_units(),
         Am_input_mean = set_units(Am_input_mean, "mg m-3") %>% set_units("umol L-1") %>% drop_units())

p5 <- pdata1 %>% 
  ggplot(aes(x = t, y = Ni_input)) +
  geom_line(linewidth = 0.75) +
  geom_line(data = pdata2, aes(x = yday, y = Ni_input_mean, colour = state), linetype = "dotted", inherit.aes = F) +
  scale_y_continuous(breaks = seq(0, 5, 0.5), limits = c(0, 3)) +
  labs(y = nitrate_lab, x = "Day of the year") +
  prettyplot_MS()

p6 <- pdata1 %>% 
  ggplot(aes(x = t, y = Am_input)) +
  geom_line(linewidth = 0.75) +
  geom_line(data = pdata2, aes(x = yday, y = Am_input_mean, colour = state), linetype = "dotted", inherit.aes = F) +
  scale_y_continuous(breaks = seq(0, 10, 0.1), limits = c(0, 0.5)) +
  labs(y = ammonium_lab, x = "Day of the year") +
  prettyplot_MS()

plot_grid(
  plot_grid(
    p1 + rm.x() + tlabel("A"), 
    p2 + rm.x() + tlabel("B"), 
    p3 + rm.x() + tlabel("C"), 
    p4 + rm.x() + tlabel("D"), 
    p5 + tlabel("E"), 
    p6 + tlabel("F"), 
    ncol = 2, align = "hv"
  ),
  ggdraw(),
  # draw_label("Day of the year", size = 12, fontfamily = "serif"),
  ncol = 1,
  rel_heights = c(0.95, 0.05)
)
```

# Figure 4: Simulated response curves

```{r fig-4}
#| code-summary: Code to generate plot
#| warning: false
#| fig-height: 5.25
#| fig-cap: Simulated response curves to environmental variables in the growth model,  including A) uptake rate of nitrate (solid line) and ammonium (dashed line), B) temperature response of A. armata (blue) and A. taxiformis (red), C) light response when grown at depths of 1 m (solid), 2 m (dashed), 5 m (dot-dashed) or 10 m (dotted), and D) salinity response.

N_uptake <- file.path(spec_data, "N_uptake.parquet") %>% 
  read_parquet()
S_response <- file.path(spec_data, "S_response.parquet") %>% 
  read_parquet()
I_response <- file.path(spec_data, "I_response.parquet") %>% 
  read_parquet() %>% 
  mutate(depth_m = as.factor(depth_m))
T_response <- file.path(spec_data, "T_response.parquet") %>% 
  read_parquet() %>% 
  mutate(species = factor(species, 
                          levels = c("taxiformis", "armata"), 
                          labels = c("taxi", "arma")))

p1 <- ggplot(N_uptake, aes(x = uM, y = uptake_uM, linetype = form)) +
  geom_line(linewidth = 0.75) +
  labs(x = expression("Ambient concentration ("*mu*"M)"), y = expression("Uptake rate ("*mu*"mol gDW"^-1*"d"^-1*")")) +
  scale_y_continuous(breaks = seq(0, 3.5, 0.05), limits = c(0, 0.30)) +
  scale_x_continuous(breaks = seq(0, 30, 5), limits = c(0,25)) +
  prettyplot_MS() + 
  theme(aspect.ratio = 0.8)

p2 <- ggplot(T_response, aes(x = temp, y = Tlim, color = species)) +
  geom_line(linewidth = 0.75) +
  labs(x = expression("Temperature ("*degree*"C)"), y = "Relative growth") +
  scale_y_continuous(breaks = seq(0, 1, 0.2), limits = c(0,1)) +
  scale_x_continuous(breaks = seq(0, 40, 5), limits = c(0,35)) +
  prettyplot_MS() + 
  theme(aspect.ratio = 0.8)

p3 <- ggplot(I_response, aes(x = light, y = Ilim, linetype = as.factor(depth_m))) +
  geom_line(linewidth = 0.75) +
  facet_wrap(facets = vars(species)) +
  labs(x = expression("Surface light (PPFD)"), y = "Relative growth") +
  scale_y_continuous(breaks = seq(0, 1, 0.2), limits = c(0,1)) +
  scale_linetype_manual(values = c("solid", "dashed", "dotdash", "dotted")) +
  prettyplot_MS() + 
  theme(aspect.ratio = 0.8, strip.text = element_blank())

p4 <- ggplot(S_response, aes(x = sali, y = Slim)) +
  geom_line(linewidth = 0.75) +
  labs(x = expression("Salinity (g L"^-1*")"), y = "Relative growth") +
  scale_y_continuous(breaks = seq(0, 1, 0.2), limits = c(0,1)) +
  scale_x_continuous(breaks = seq(0, 60, 5), limits = c(19,45)) +
  prettyplot_MS() + 
  theme(aspect.ratio = 0.8)

plot_grid(
  plot_grid(
    p1 + tlabel("A"), 
    p2 + tlabel("B") + theme(axis.title.y = element_text(vjust = 0.5)), 
    p3 + tlabel("C") + theme(axis.title.y = element_text(vjust = 0.5)),  
    p4 + tlabel("D", x = 19) + theme(axis.title.y = element_text(vjust = 0.5)), 
    ncol = 2, align = "hv"
  ),
  ncol = 1,
  rel_heights = c(0.95, 0.05)  # Adjust these values to control spacing
)
```

# Figure 5: N removal in ambient conditions

```{r normal-growth-end-data}
#| code-summary: Get cell growth in ambient (normal) conditions

norm_growth_end <- file.path(runs_data, "cell_growth_end.parquet") %>% 
  read_parquet(mmap = T) %>% 
  merge(cell_coords, by = "cell_no")

# Total N removed at each harvest
norm_growth_end <- norm_growth_end %>% 
  mutate(
    value = TN_end - TN_start,
    value = case_when(value == 0 ~ NA, T ~ value),
    month = as.factor(month(parse_date_time(start, orders = "j"))),
    species = factor(species, levels = c("armata", "taxiformis"), labels = c("A. armata", "A. taxiformis"))
  )
```

```{r fig-5}
#| code-summary: Code to generate plot
#| warning: false
#| fig-width: 6
#| fig-height: 7.75
#| fig-cap: Nitrogen removal per month at harvest by Asparagopsis armata (solid lines) and Asparagopsis taxiformis (dashed lines) in the 12 monthly growth periods in each state in typical environmental conditions. Bold lines show mean values while shaded areas show mean ± standard deviation.

norm_growth_end %>%
  mutate(value = case_when(value <= 0 ~ 0, T ~ value)) %>% 
  group_by(species, state, start) %>%
  reframe(
    mean = meanna(value),
    sd = sdna(value)
  ) %>%
  mutate(
    sd = case_when(sd > mean ~ mean, T ~ sd),
    state = factor(state, levels = states_ord, labels = states_lng)
  ) %>%
  ggplot(
    aes(
      x = start, y = mean, ymin = mean - sd, ymax = mean + sd,
      color = state, fill = state, linetype = species
    )
  ) +
  geom_line(linewidth = 0.75) +
  geom_ribbon(alpha = 0.25, linewidth = 0.25) +
  facet_wrap(facets = vars(state), ncol = 2) +
  # scale_y_continuous(breaks = seq(0, 150, 25), limits = c(0, 175), expand = c(0,0)) +
  scale_color_manual(values = states_pal) +
  scale_fill_manual(values = states_pal) +
  env_plot_MS() +
  labs(x = "Day of the year", y = expression("Total N removed (mg m"^-2 *")"))
```

# Figure 7: Ambient vs supplemented growth

```{r normal-growth-total-data}
#| code-summary: Get total cell growth in ambient (normal) conditions

norm_growth_total <- norm_growth_end %>% 
  mutate(
    value_na = case_when(value <= 0 ~ NA, T ~ value),
    value_0s = case_when(value <= 0 ~ 0, T ~ value)
  ) %>% 
  group_by(state, species, cell_no, latitude, longitude) %>% 
  reframe(
    total = sumna(value),            # good growth months may be outweighed by bad growth months
    total_na = sumna(value_na),      # bad growth months are removed
    total_na = case_when(total_na == 0 ~ NA, T ~ total_na),
    total_0s = sumna(value_0s)       # bad growth months count as zeros
  )
```

```{r supp-growth-total-data}
#| code-summary: Get cell growth in nitrogen supplemented conditions

supp_growth_end_raw <- file.path(runs_data, "cell_growth_supp_end.parquet") %>% 
  read_parquet(mmap = T) %>% 
  merge(cell_coords, by = "cell_no")

supp_growth_end <- supp_growth_end_raw %>% 
  mutate(
    value = TN_end - TN_start,
    value = case_when(value == 0 ~ NA, T ~ value),
    month = as.factor(month(parse_date_time(start, orders = "j"))),
    species = factor(species, levels = c("armata", "taxiformis"), labels = c("A. armata", "A. taxiformis"))
  )

supp_growth_total <- supp_growth_end %>% 
  mutate(
    value_na = case_when(value <= 0 ~ NA, T ~ value),
    value_0s = case_when(value <= 0 ~ 0, T ~ value)
  ) %>% 
  group_by(state, species, cell_no, latitude, longitude) %>% 
  reframe(
    total = sumna(value),            # good growth months may be outweighed by bad growth months
    total_na = sumna(value_na),      # bad growth months are removed
    total_na = case_when(total_na == 0 ~ NA, T ~ total_na),
    total_0s = sumna(value_0s)       # bad growth months count as zeros
  )
```

```{r fig-6}
#| code-summary: Code to generate plot
#| fig-cap: |
#    "Total nitrogen removal by A. armata (top: A, B) and A. taxiformis (bottom: C, D) summed across one year of typical environmental conditions (left: A, C) or with supplemented environmental conditions (right: B, D). Grey cells show areas where growth failed or where nitrogen removal was < 0 across the entire year."

all_growth_total <- rbind(
  norm_growth_total %>% mutate(scenario = "norm_conditions"),
  supp_growth_total %>% mutate(scenario = "supp_conditions")
)

all_growth_total %>%
  ggplot(aes(x = longitude, y = latitude, fill = total_na)) +
  geom_tile() +
  geom_sf(data = ozmap_data(data = "states"), inherit.aes = F) +
  coord_sf(xlim = c(111.5, 155), ylim = c(-44.5, -8.5), expand = F) +
  scale_x_continuous(breaks = seq(100, 160, 10)) +
  scale_y_continuous(breaks = seq(-2.5, -45, -5)) +
  scale_fill_viridis_c(
    limits = c(1, 10000),
    labels = scales::label_number(),
    trans = "log10",
    na.value = "grey",
    guide = guide_colorbar(
      title = expression("Total N removed (mg m"^-2 * ")"),
      barheight = 1,
      barwidth = 20,
      title.position = "top",
      title.hjust = 0.5
    )
  ) +
  labs(x = "Longitude", y = "Latitude") +
  facet_grid(rows = vars(species), cols = vars(scenario)) +
  prettyplot_MS() +
  theme(legend.position = "top", strip.text = element_blank(), strip.background = element_blank())
```

# Figure 7: N removal in supplemented conditions

```{r supplemented-growth-end-data}
#| code-summary: Get cell growth in supplemented conditions

supp_growth_end <- file.path(runs_data, "cell_growth_supp_end.parquet") %>% 
  read_parquet(mmap = T) %>% 
  merge(cell_coords, by = "cell_no")

# Total N removed at each harvest
supp_growth_end <- supp_growth_end %>% 
  mutate(
    value = TN_end - TN_start,
    value = case_when(value == 0 ~ NA, T ~ value),
    month = as.factor(month(parse_date_time(start, orders = "j"))),
    species = factor(species, levels = c("armata", "taxiformis"), labels = c("A. armata", "A. taxiformis"))
  )
```

```{r fig-7}
#| code-summary: Code to generate plot
#| warning: false
#| fig-width: 6
#| fig-height: 7.75
#| fig-cap: Nitrogen removal per month at harvest by Asparagopsis armata (solid lines) and Asparagopsis taxiformis (dashed lines) in the 12 monthly growth periods in each state. Bold lines show mean values while shaded areas show mean ± standard deviation.

supp_growth_end %>%
  mutate(value = case_when(value <= 0 ~ 0, T ~ value)) %>% 
  group_by(species, state, start) %>%
  reframe(
    mean = meanna(value),
    sd = sdna(value)
  ) %>%
  mutate(
    sd = case_when(sd > mean ~ mean, T ~ sd),
    state = factor(state, levels = states_ord, labels = states_lng)
  ) %>%
  ggplot(
    aes(
      x = start, y = mean, ymin = mean - sd, ymax = mean + sd,
      color = state, fill = state, linetype = species
    )
  ) +
  geom_line(linewidth = 0.75) +
  geom_ribbon(alpha = 0.25, linewidth = 0.25) +
  facet_wrap(facets = vars(state), ncol = 2) +
  # scale_y_continuous(breaks = seq(0, 150, 25), limits = c(0, 175), expand = c(0,0)) +
  scale_color_manual(values = states_pal) +
  scale_fill_manual(values = states_pal) +
  env_plot_MS() +
  labs(x = "Day of the year", y = expression("Total N removed (mg m"^-2 *")"))
```

# Supplementary figures

```{r all-cell-input-timeseries}
#| code-summary: Load environmental timeseries

T_input <- file.path(envi_data, "cell_input_all") %>% 
  list.files(full.names = T) %>% 
  purrr::map(function(fnm) {
    fnm %>% arrow::read_parquet(col_select = c("cell_no", "state", "yday", "T_input"))
  }) %>% 
  bind_rows() %>% 
  mutate(state = factor(state, levels = states_ord, labels = states_lng))

I_input <- file.path(envi_data, "cell_input_all") %>% 
  list.files(full.names = T) %>% 
  purrr::map(function(fnm) {
    fnm %>% arrow::read_parquet(col_select = c("cell_no", "state", "yday", "I_input"))
  }) %>% 
  bind_rows() %>% 
  mutate(state = factor(state, levels = states_ord, labels = states_lng))

K_input <- file.path(envi_data, "cell_input_all") %>% 
  list.files(full.names = T) %>% 
  purrr::map(function(fnm) {
    fnm %>% arrow::read_parquet(col_select = c("cell_no", "state", "yday", "Kd_490"))
  }) %>% 
  bind_rows() %>% 
  mutate(state = factor(state, levels = states_ord, labels = states_lng))

S_input <- file.path(envi_data, "cell_input_all") %>% 
  list.files(full.names = T) %>% 
  purrr::map(function(fnm) {
    fnm %>% arrow::read_parquet(col_select = c("cell_no", "state", "yday", "S_input"))
  }) %>% 
  bind_rows() %>% 
  mutate(state = factor(state, levels = states_ord, labels = states_lng))

UV_input <- file.path(envi_data, "cell_input_all") %>% 
  list.files(full.names = T) %>% 
  purrr::map(function(fnm) {
    fnm %>% arrow::read_parquet(col_select = c("cell_no", "state", "yday", "UV_input"))
  }) %>% 
  bind_rows() %>% 
  mutate(state = factor(state, levels = states_ord, labels = states_lng))
```

## Figure S1: Temperature

```{r fig-S1}
#| code-summary: Code to generate plot
#| warning: false
#| fig-width: 6
#| fig-height: 7.75
#| fig-cap: Summary of temperature values across the year within each state, averaged across 2019-2023. Bold lines show the mean across all cells within each state while ribbons show the standard deviation between cells.

T_input_state <- T_input %>%
  group_by(state, yday) %>% 
  reframe(value = mean(T_input, na.rm = T),
          sd = sd(T_input, na.rm = T))

(
  p_input_state_av <- T_input_state %>% 
    ggplot(aes(x = yday, y = value, ymin = value-sd, ymax = value+sd, colour = state, fill = state)) +
    geom_line(linewidth = 0.75) +
    geom_ribbon(alpha = 0.35, linewidth = 0.25) +
    facet_wrap(facets = vars(state), ncol = 2) +
    scale_y_continuous(breaks = seq(0, 40, 5), limits = c(10, 35)) +
    scale_colour_manual(values = states_pal) +
    scale_fill_manual(values = states_pal) +
    env_plot_MS() +
    labs(x = "Day of the year", y = expression("Temperature ("*degree*"C)"))
)
```

## Figure S2: Light

```{r fig-S2}
#| code-summary: Code to generate plot
#| warning: false
#| fig-width: 6
#| fig-height: 7.75
#| fig-cap: Summary of irradiance values across the year within each state, averaged across 2019-2023. Bold lines show the mean across all cells within each state while ribbons show the standard deviation between cells.

I_input_state <- I_input %>%
  group_by(state, yday) %>% 
  reframe(value = mean(I_input, na.rm = T),
          sd = sd(I_input, na.rm = T))

p_input_state_av %+% I_input_state +
  scale_y_continuous(breaks = seq(0, 1600, 150), limits = c(0, 650)) +
  labs(x = "Day of the year", y = expression("Irradiance (photons m"^-2*" s"^-1*")")) 
```

## Figure S3: Light attenuation

```{r fig-S3}
#| code-summary: Code to generate plot
#| warning: false
#| fig-width: 6
#| fig-height: 7.75
#| fig-cap: Summary of light attenuation coefficient values across the year within each state, averaged across 2019-2023. Bold lines show the mean across all cells within each state while ribbons show the standard deviation between cells.

K_input_state <- K_input %>%
  group_by(state, yday) %>% 
  reframe(value = mean(Kd_490, na.rm = T),
          sd = sd(Kd_490, na.rm = T)) %>% 
  mutate(sd = case_when(sd > value ~ value, T ~ sd))

p_input_state_av %+% K_input_state +
  scale_y_continuous(breaks = seq(0, 0.7, 0.1), limits = c(0, 0.62)) +
  labs(x = "Day of the year", y = expression("Light attenuation coefficient (m"^-1*" )"))
```

## Figure S4: Salinity

```{r fig-S4}
#| code-summary: Code to generate plot
#| warning: false
#| fig-width: 6
#| fig-height: 7.75
#| fig-cap: Summary of salinity values across the year within each state, averaged across 2019-2023 and all depths < 15 m. Bold lines show the mean across all cells within each state while ribbons show the standard deviation between cells.

S_input_state <- S_input %>%
  group_by(state, yday) %>% 
  reframe(value = mean(S_input, na.rm = T),
          sd = sd(S_input, na.rm = T))

p_input_state_av %+% S_input_state +
  scale_y_continuous(breaks = seq(30, 40, 0.5), limits = c(33.85, 36.5)) +
  labs(x = "Day of the year", y = expression("Salinity (g L"^-1*")"))
```

## Figure S5: Water velocity

```{r fig-S5}
#| code-summary: Code to generate plot
#| warning: false
#| fig-width: 6
#| fig-height: 7.75
#| fig-cap: Summary of water velocity values across the year within each state, averaged across 2019-2023 and all depths < 15 m. Bold lines show the mean across all cells within each state while ribbons show the standard deviation between cells.

UV_input_state <- UV_input %>%
  group_by(state, yday) %>% 
  reframe(value = mean(UV_input, na.rm = T),
          sd = sd(UV_input, na.rm = T)) %>% 
  mutate(sd_min = case_when(sd > value ~ value, T ~ sd))

(
  p_input_state_av <- UV_input_state %>% 
    ggplot(aes(x = yday, y = value, ymin = value-sd_min, ymax = value+sd, colour = state, fill = state)) +
    geom_line(linewidth = 0.75) +
    geom_ribbon(alpha = 0.35, linewidth = 0.25) +
    facet_wrap(facets = vars(state), ncol = 2) +
    scale_y_continuous(breaks = seq(0, 1, 0.2), limits = c(-0.01, 1)) +
    scale_colour_manual(values = states_pal) +
    scale_fill_manual(values = states_pal) +
    env_plot_MS() +
    labs(x = "Day of the year", y = expression("Water velocity (m s"^-1*")"))
)
```

## Figure S6: Example N input

```{r eg-cell-data}
#| code-summary: Load N data for example cell

eg_cell <- 108843

BARRA_R2_cell_nos <- file.path(envi_data, "BARRA_R2_cell_nos.qs") %>% qs::qread()
ind <- which(BARRA_R2_cell_nos == eg_cell)

N_data <- find_read(file.path(envi_data, "combined_N_data"), "N_data_combined") %>% 
  filter(cell_no == eg_cell) %>% 
  mutate(value = value %>% set_units("mg m-3") %>% set_units("umol L-1") %>% drop_units())

N_curve <- find_read(file.path(envi_data, "combined_N_data"), "parameter") %>% 
  filter(cell_no == eg_cell) 
N_curve <- data.frame(yday = 1:365) %>% 
  mutate(ammonia = N_curve$a[N_curve$measure == "Am"] + N_curve$b[N_curve$measure == "Am"] * 
           sin((yday * pi + N_curve$c[N_curve$measure == "Am"]) / 182.5),
         nitrate_nitrite = N_curve$a[N_curve$measure == "Ni"] + N_curve$b[N_curve$measure == "Ni"] * 
           sin((yday * pi + N_curve$c[N_curve$measure == "Ni"]) / 182.5)) %>% 
  pivot_longer(names_to = "indicator", values_to = "value", cols = c("ammonia", "nitrate_nitrite"), 
               names_transform = list(indicator = as.factor)) %>% 
  mutate(value = value %>% set_units("mg m-3") %>% set_units("umol L-1"))

N_input <- file.path(envi_data, "cell_input_all") %>% 
  list.files(full.names = T) %>% 
  purrr::map(function(fnm){
    read_parquet(fnm) %>% 
      filter(cell_no == eg_cell)
  }) %>% 
  bind_rows() %>% 
  select(yday, Ni_input, Am_input) %>% 
  pivot_longer(names_to = "indicator", values_to = "value", cols = c("Ni_input", "Am_input"), 
               names_transform = list(indicator = as.factor)) %>% 
  mutate(value = value %>% set_units("mg m-3") %>% set_units("umol L-1") %>% drop_units(),
         indicator = factor(indicator, levels = c("Am_input", "Ni_input"), labels = levels(N_data$indicator)))
```

```{r fig-S6}
#| code-summary: Code to generate plot
#| fig-height: 6.5
#| warning: false
#| fig-cap: An example of the daily ammonium (A) and nitrate (B) input and data for a single cell in New South Wales. Black lines show the input driving the growth model (constructed curve), blue points show data from the two nearest reference stations (Port Hacking and North Stradbroke Island) and orange dots show data from outfall points within 48 km (Bombo, Coniston Beach, Penguin Heads, and Shellharbour). Point size indicates the relative data weight (inverse distance from the cell centre) in curve construction. Note the different y-axes.

p1_df <- N_data %>% filter(indicator == "ammonia")
p1_df2 <- N_input %>% filter(indicator == "ammonia")
p2_df <- N_data %>% filter(indicator != "ammonia")
p2_df2 <- N_input %>% filter(indicator != "ammonia")

p1 <- p1_df %>% 
  ggplot(aes(x = yday, y = value, fill = data_source, colour = data_source, size = weight)) +
  geom_point(shape = 21, alpha = 0.15) +
  # facet_wrap(facets = vars(indicator), nrow = 2, scales = "free") +
  geom_line(data = p1_df2, aes(x = yday, y = value), inherit.aes = F) +
  scale_fill_manual(values = c("steelblue", "orange")) +
  scale_colour_manual(values = c("steelblue", "orange")) +
  # scale_y_continuous(limits = c(0, 3.5), breaks = seq(0, 4, 0.5)) +
  env_plot_MS() +
  theme(aspect.ratio = 0.55) +
  scale_x_continuous(breaks = seq(0,360,60)) +
  labs(x = "Day of the year", y = ammonium_lab)

p2 <- p2_df %>% 
  ggplot(aes(x = yday, y = value, fill = data_source, colour = data_source, size = weight)) +
  geom_point(shape = 21, alpha = 0.15) +
  # facet_wrap(facets = vars(indicator), nrow = 2, scales = "free") +
  geom_line(data = p2_df2, aes(x = yday, y = value), inherit.aes = F) +
  scale_fill_manual(values = c("steelblue", "orange")) +
  scale_colour_manual(values = c("steelblue", "orange")) +
  # scale_y_continuous(limits = c(0, 3.5), breaks = seq(0, 4, 0.5)) +
  env_plot_MS() +
  theme(aspect.ratio = 0.55, axis.title.y = element_text(vjust = 0.5)) +
  scale_x_continuous(breaks = seq(0,360,60)) +
  labs(x = "Day of the year", y = nitrate_lab)

plot_grid(
  p1 + tlabel("A") + rm.x(),
  p2 + tlabel("B"),
  nrow = 2, align = "hv"
)
```

## Figure S7: Nitrogen

```{r Ni-input-data}
#| code-summary: Get nitrate input data

Ni_input <- file.path(envi_data, "cell_input_all") %>% 
  list.files(full.names = T) %>% 
  purrr::map(function(fnm) {
    fnm %>% arrow::read_parquet(col_select = c("cell_no", "state", "yday", "Ni_input"))
  }) %>% 
  bind_rows() %>% 
  mutate(state = factor(state, levels = states_ord, labels = states_lng)) %>% 
  mutate(Ni_input = Ni_input %>% set_units("mg m-3") %>% set_units("umol L-1"))
```

```{r fig-S7}
#| code-summary: Code to generate plot
#| warning: false
#| fig-cap: Summary of nitrate concentration values across the year within each state, averaged across 2019-2023. Bold lines show the mean across all cells within each state while ribbons show the standard deviation between cells.

Ni_input_state <- Ni_input %>%
  group_by(state, yday) %>% 
  reframe(value = mean(Ni_input, na.rm = T),
          sd = sd(Ni_input, na.rm = T)) %>% 
  mutate(value = value %>% drop_units())

(
  p_input_state_av <- Ni_input_state %>% 
    ggplot(aes(x = yday, y = value, ymin = value-sd, ymax = value+sd, colour = state, fill = state)) +
    geom_line(linewidth = 0.75) +
    geom_ribbon(alpha = 0.35, linewidth = 0.25) +
    facet_wrap(facets = vars(state), ncol = 2) +
    # scale_y_continuous(breaks = seq(0, 40, 5), limits = c(10, 35)) +
    scale_colour_manual(values = states_pal) +
    scale_fill_manual(values = states_pal) +
    env_plot_MS() +
    labs(x = "Day of the year", y = nitrate_lab)
)
```

## Figure S8: Ammonium

```{r Am-input-data}
#| code-summary: Get ammonium input data

Am_input <- file.path(envi_data, "cell_input_all") %>% 
  list.files(full.names = T) %>% 
  purrr::map(function(fnm) {
    fnm %>% arrow::read_parquet(col_select = c("cell_no", "state", "yday", "Am_input"))
  }) %>% 
  bind_rows() %>% 
  mutate(state = factor(state, levels = states_ord, labels = states_lng)) %>% 
  mutate(Am_input = Am_input %>% set_units("mg m-3") %>% set_units("umol L-1"))
```

```{r fig-S8}
#| code-summary: Code to generate plot
#| warning: false
#| fig-cap: Summary of ammonium concentration values across the year within each state, averaged across 2019-2023. Bold lines show the mean across all cells within each state while ribbons show the standard deviation between cells.

Am_input_state <- Am_input %>%
  group_by(state, yday) %>% 
  reframe(value = mean(Am_input, na.rm = T),
          sd = sd(Am_input, na.rm = T)) %>% 
  mutate(value = value %>% drop_units())

p_input_state_av %+% Am_input_state +
  # scale_y_continuous(breaks = seq(30, 40, 0.5), limits = c(34, 36.5)) +
  labs(x = "Day of the year", y = ammonium_lab)
```

## Figure S9: Model sensitivity to parameter changes

```{r sensitivity-data}
sens_arma <- file.path(spec_data, "sensitivity_arma_calc_combTU.parquet") %>% 
  read_parquet() %>% 
  dplyr::filter(!param %in% c("M_am", "C_am", "M_ot", "C_ot")) %>% 
  mutate(cult_dep = as.factor(cult_dep))
exlude <- unique(sens_arma$param[abs(sens_arma$sens) < 0.01]) %>% as.character
sens_arma <- sens_arma %>% 
  dplyr::filter(!param %in% c(exlude, "M_am", "C_am", "M_ot", "C_ot"))
sens_arma$param <- droplevels(sens_arma$param)

# levels(sens_arma$param)
# exlude

# Add axis labels
param.labs.arma <- c(
  bquote("V"["Am"]), 
  bquote("K"["Am"]),
  # bquote("M"["Ni"]),
  bquote("C"["Ni"]), 
  bquote("Q"["min"]), 
  bquote("Q"["max"]),
  # bquote("K"["c"]),
  bquote(mu), 
  bquote("D"["ve"]),
  # bquote("a"["cs"]), 
  # bquote("I"["o"]),
  bquote("T"["opt"]), 
  bquote("T"["min"]), 
  bquote("T"["max"])
  # bquote("S"["opt"]), 
  # bquote("S"["min"]), 
  # bquote("S"["max"]), 
  # bquote("h"["a"]), 
  # bquote("h"["b"]), 
  # bquote("h"["c"]), 
  # bquote("h"["max"]), 
  # "DW:WW"
)

sens_taxi <- file.path(spec_data, "sensitivity_taxi_calc_combTU.parquet") %>% 
  read_parquet() %>% 
  mutate(cult_dep = as.factor(cult_dep))
exlude <- unique(sens_taxi$param[abs(sens_taxi$sens) < 0.01]) %>% as.character
sens_taxi <- sens_taxi %>% 
  dplyr::filter(!param %in% c(exlude, "M_am", "C_am", "M_ot", "C_ot"))
sens_taxi$param <- droplevels(sens_taxi$param)

# levels(sens_taxi$param)
# exlude

# Add axis labels
param.labs.taxi <- param.labs.arma
```

```{r fig-S9}
#| code-summary: Code to generate plot
#| warning: false
#| fig-height: 7.5
#| fig-cap: Relative impact of increasing each model parameter for Asparagopsis armata (A) and Asparagopsis taxiformis (B) by 10% on total N removed. Colours show the different culture depths of 0.5 m (orange), 2.5 m (blue), and 5 m (green) below the surface. Error bars show variability (SD) across the 12 months of a typical year in each state.

p1_df <- sens_arma %>% filter(!is.na(sens) & cult_dep != "10")
p2_df <- sens_taxi %>% filter(!is.na(sens) & cult_dep != "10")

p1 <- p1_df %>% 
  ggplot(aes(x = rev(param), y = sens, fill = as.factor(cult_dep), ymin = sens-sd, ymax = sens+sd)) +
  geom_col(position = position_dodge(), width = 0.95, colour = "black") +
  geom_errorbar(position = position_dodge(width = 0.95), width = 0.5) +
  coord_flip() +
  scale_y_continuous(limits = c(-1.1, 1.1), breaks = seq(-1.5, 1.5, 0.2)) +
  scale_x_discrete(limits = levels(sens_arma$param), labels = rev(param.labs.arma)) +
  scale_fill_manual(values = depths_pal) +
  labs(y = "Total N sensitivity", x = "Parameter name") +
  prettyplot_MS() +
  theme(axis.title.x = element_blank(), 
        axis.text.x = element_blank(), 
        axis.ticks.x = element_blank(),
        aspect.ratio = 0.7)

p2 <- p2_df %>% 
  ggplot(aes(x = rev(param), y = sens, fill = as.factor(cult_dep), ymin = sens-sd, ymax = sens+sd)) +
  geom_col(position = position_dodge(), width = 0.95, colour = "black") +
  geom_errorbar(position = position_dodge(width = 0.95), width = 0.5) +
  coord_flip() +
  scale_y_continuous(limits = c(-1.1, 1.1), breaks = seq(-1.5, 1.5, 0.2)) +
  scale_x_discrete(limits = levels(sens_arma$param), labels = rev(param.labs.taxi)) +
  scale_fill_manual(values = depths_pal) +
  labs(y = "Total N sensitivity", x = "Parameter name") +
  prettyplot_MS() + 
  theme(aspect.ratio = 0.7)

plot_grid(
  p1 + tlabel("A", x = Inf, y = -1.1), 
  p2 + tlabel("B", x = Inf, y = -1.1), 
  nrow = 2, align = "hv"
)
```

## Figure S10: Factors limiting growth (*A. armata*)

```{r growth-lim-plot-data-armata}
# This has cell_no, lat, lon, month and lims (for maps)
armata_lims <- map(1:12, function(m) {
  file.path(runs_data, str_c("armata_dom_lims_raster_", fixnum(m,2), ".tif")) %>% 
    terra::rast() %>% 
    as.data.frame(xy=TRUE) %>% 
    rename(longitude = x, latitude = y) %>% 
    mutate(lim = factor(lim, levels = 1:4, labels = c("T_lim", "I_lim", "Q_lim", "S_lim")))
  })

# This has only latitude, month and lim (for histograms)
forhist_armata <- find_read(runs_data, "forhist_armata.parquet") %>% split(.$month)
```

```{r growth-lim-plot-templates}
p_dom_hist.1 <- forhist_armata[[1]] %>% 
  ggplot(aes(y = latitude, x = after_stat(100*count/(13207 * 30)), fill = lim, colour = lim)) +
  geom_histogram(position = "stack", alpha = 0.75, binwidth = 0.75) +
  scale_fill_manual(values = lims_pal) +
  scale_colour_manual(values = lims_pal) +
  scale_y_continuous(limits = c(-44.5, -8.5), expand = c(0,0)) +
  scale_x_continuous(limits = c(0, 5.5), breaks = seq(0, 5, 1), expand = c(0.02,0)) +
  labs(x = expression("% of growing days in all cells"), y = "Latitude") +
  prettyplot_MS() +
  rm.y() + 
  theme(plot.margin = margin(0, 0, 0, 0, "pt"), panel.spacing = unit(1, "lines"))

p_dom_map.1 <- armata_lims[[1]] %>% 
  ggplot(aes(x = longitude, y = latitude, fill = lim)) +
  geom_raster() +
  geom_sf(data = ozmap_data(data = "states"), inherit.aes = F) +
  coord_sf(xlim = c(111.5, 155), ylim = c(-44.5, -8.5), expand = F) +
  scale_x_continuous(breaks = seq(112.5, 160, 10)) +
  scale_y_continuous(breaks = seq(-2.5, -45, -5)) +
  scale_color_brewer(palette = "Dark2") +
  scale_fill_manual(values = lims_pal) +
  labs(x = "Longitude", y = "Latitude") +
  prettyplot_MS() + 
  theme(plot.margin = margin(0, 0, 0, 0, "pt"), axis.title.x = element_text(vjust = 1))
```

```{r fig-S10}
#| code-summary: Code to generate plot
#| warning: false
#| fig-height: 3.5
#| fig-width: 6.25

plot_grid(
  p_dom_map.1 %+% armata_lims[[2]] + tlabel(month.name[2], x = 114.5),
  p_dom_hist.1 %+% forhist_armata[[2]],
  ncol = 2, 
  align = "h", 
  rel_widths = c(1, 0.625)
)  

plot_grid(
  p_dom_map.1 %+% armata_lims[[5]] + tlabel(month.name[5], x = 114.5),
  p_dom_hist.1 %+% forhist_armata[[5]],
  ncol = 2, 
  align = "h", 
  rel_widths = c(1, 0.625)
)

plot_grid(
  p_dom_map.1 %+% armata_lims[[8]] + tlabel(month.name[8], x = 114.5),
  p_dom_hist.1 %+% forhist_armata[[8]],
  ncol = 2, 
  align = "h", 
  rel_widths = c(1, 0.625)
)

plot_grid(
  p_dom_map.1 %+% armata_lims[[11]] + tlabel(month.name[11], x = 114.5),
  p_dom_hist.1 %+% forhist_armata[[11]],
  ncol = 2, 
  align = "h", 
  rel_widths = c(1, 0.625)
)
```

## Figure S11: Factors limiting growth (*A. taxiformis*)

```{r growth-lim-plot-data-taxiformis}
# This has cell_no, lat, lon, month and lims (for maps)
taxiformis_lims <- map(1:12, function(m) {
  file.path(runs_data, str_c("taxiformis_dom_lims_raster_", fixnum(m,2), ".tif")) %>% 
    terra::rast() %>% 
    as.data.frame(xy=TRUE) %>% 
    rename(longitude = x, latitude = y) %>% 
    mutate(lim = factor(lim, levels = 1:4, labels = c("T_lim", "I_lim", "Q_lim", "S_lim")))
  })

# This has only latitude, month and lim (for histograms)
forhist_taxiformis <- find_read(runs_data, "forhist_taxiformis.parquet") %>% split(.$month)
```

```{r fig-S11}
#| code-summary: Code to generate plot
#| warning: false
#| fig-height: 3.5
#| fig-width: 6.25

plot_grid(
  p_dom_map.1 %+% taxiformis_lims[[2]] + tlabel(month.name[2], x = 114.5),
  p_dom_hist.1 %+% forhist_taxiformis[[2]],
  ncol = 2, 
  align = "h", 
  rel_widths = c(1, 0.625)
)  

plot_grid(
  p_dom_map.1 %+% taxiformis_lims[[5]] + tlabel(month.name[5], x = 114.5),
  p_dom_hist.1 %+% forhist_taxiformis[[5]],
  ncol = 2, 
  align = "h", 
  rel_widths = c(1, 0.625)
)  

plot_grid(
  p_dom_map.1 %+% taxiformis_lims[[8]] + tlabel(month.name[8], x = 114.5),
  p_dom_hist.1 %+% forhist_taxiformis[[8]],
  ncol = 2, 
  align = "h", 
  rel_widths = c(1, 0.625)
)  

plot_grid(
  p_dom_map.1 %+% taxiformis_lims[[11]] + tlabel(month.name[11], x = 114.5),
  p_dom_hist.1 %+% forhist_taxiformis[[11]],
  ncol = 2, 
  align = "h", 
  rel_widths = c(1, 0.625)
)  
```

