---
title: "Results: Ambient growth limitations"
subtitle: "Modelling <i>Asparagopsis</i> nitrogen bioremediation efficiency in Australian coastal environments"
author: Tormey Reimer
date: last-modified
---

```{r setup}
#| eval: true
#| include: false
#| cache: false

# You need quarto to generate this markdown - install here: https://quarto.org/docs/get-started/
# This markdown uses TinyTex for PDF rendering - install with tinytex::install_tinytex()
# This markdown uses the acronyms extension - install by entering quarto add rchaput/acronyms@master into the terminal

library(stringr)
library(units)
library(targets)
library(tarchetypes)
library(macrogrow)
library(ozmaps)
library(arrow)
library(gganimate)
library(magrittr)
library(ggplot2)
library(dplyr)
library(terra)
library(tidyterra)
library(here)
library(lubridate)
library(tidyr)
library(cowplot)
library(purrr)
library(patchwork)
library(conflicted)
library(gridExtra)
library(grid)

conflicts_prefer(dplyr::select, dplyr::filter)

source(here("R_scripts/00_functions.R"))
source(here("R_scripts/00_global_displays.R"))
```

```{r cell-coords}
#| eval: true
#| code-summary: |
#|   Load in all the BARRA-R2 cells being used, with their coordinates.

cell_coords <- find_read(envi_data, "R2_cell_coords") %>% 
  mutate(state = factor(state, levels = states_ord)) %>% 
  dplyr::select(-layer)
cells_included <- find_read(runs_data, "include")
cells_omitted <- find_read(runs_data, "omit")

totals <- cell_coords %>% 
  filter(!cell_no %in% cells_omitted) %>% 
  group_by(state) %>% 
  reframe(cells = n())
```

The point of this section is to show the most common limiting factor - overall, between species, temporally, geographically.

```{r growth-lim-plot-data}
# These have cell_no, lat, lon, month and lims (for maps)
armata_lims <- map(1:12, function(m) {
  file.path(runs_data, str_c("armata_dom_lims_raster_", fixnum(m,2), ".tif")) %>% 
    terra::rast() %>% 
    as.data.frame(xy=TRUE) %>% 
    rename(longitude = x, latitude = y) %>% 
    mutate(lim = factor(lim, levels = 1:4, labels = c("T_lim", "I_lim", "Q_lim", "S_lim")))
  })
taxiformis_lims <- map(1:12, function(m) {
  file.path(runs_data, str_c("taxiformis_dom_lims_raster_", fixnum(m,2), ".tif")) %>% 
    terra::rast() %>% 
    as.data.frame(xy=TRUE) %>% 
    rename(longitude = x, latitude = y) %>% 
    mutate(lim = factor(lim, levels = 1:4, labels = c("T_lim", "I_lim", "Q_lim", "S_lim")))
  })

# These have only latitude, month and lim (for histograms)
forhist_armata <- find_read(runs_data, "forhist_armata.parquet") %>% split(.$month)
forhist_taxiformis <- find_read(runs_data, "forhist_taxiformis.parquet") %>% split(.$month)
```

```{r growth-lim-plot-templates}
p_dom_hist.1 <- forhist_armata[[1]] %>% 
  ggplot(aes(y = latitude, x = after_stat(100*count/(13207 * 30)), fill = lim, colour = lim)) +
  geom_histogram(position = "stack", alpha = 0.75, binwidth = 0.75) +
  scale_fill_manual(values = lims_pal) +
  scale_colour_manual(values = lims_pal) +
  scale_y_continuous(limits = c(-44.5, -8.5), expand = c(0,0)) +
  scale_x_continuous(limits = c(0, 5.5), breaks = seq(0, 5, 1), expand = c(0.02,0)) +
  labs(x = expression("% of growing days in all cells"), y = "Latitude") +
  prettyplot() +
  rm.y() + 
  theme(plot.margin = margin(0, 0, 0, 0, "pt"), panel.spacing = unit(1, "lines"))

p_dom_map.1 <- armata_lims[[1]] %>% 
  ggplot(aes(x = longitude, y = latitude, fill = lim)) +
  geom_raster() +
  geom_sf(data = ozmap_data(data = "states"), inherit.aes = F) +
  coord_sf(xlim = c(111.5, 155), ylim = c(-44.5, -8.5), expand = F) +
  scale_x_continuous(breaks = seq(112.5, 160, 10)) +
  scale_y_continuous(breaks = seq(-2.5, -45, -5)) +
  scale_color_brewer(palette = "Dark2") +
  scale_fill_manual(values = lims_pal) +
  labs(x = "Longitude", y = "Latitude") +
  prettyplot() + 
  theme(plot.margin = margin(0, 0, 0, 0, "pt"), axis.title.x = element_text(vjust = 1))
```

### _Asparagopsis armata_

```{r armata-lims-stats}
#| eval: false

stat1 <- armata_lims %>% 
  filter(latitude > -30) 
stat1 <- stat1 %>% 
  group_by(cell_no, lim) %>% 
  reframe(perc_days = n()/nrow(stat1))
stat1_T <- sum(stat1$perc_days[stat1$lim == "T_lim"])
stat1_Q <- sum(stat1$perc_days[stat1$lim == "Q_lim"])
stat1_I <- sum(stat1$perc_days[stat1$lim == "I_lim"])

stat2 <- armata_lims %>% 
  filter(latitude <= -30)
stat1 <- stat1 %>% 
  group_by(cell_no, lim) %>% 
  reframe(perc_days = n()/nrow(stat1))
stat2_T <- sum(stat2$perc_days[stat2$lim == "T_lim"])
stat2_Q <- sum(stat2$perc_days[stat2$lim == "Q_lim"])
stat2_I <- sum(stat2$perc_days[stat2$lim == "I_lim"])

# For _A. armata_ grown at latitudes > -30$^\circ$, a mean of
# `r round(100*stat1_T,1)`% of growing days were limited by temperature,
# `r round(100*stat1_Q,1)`% were limited by nutrients, and
# `r round(100*stat1_I,1)`% were limited by light.
# 
# For _A. armata_ grown at latitudes < -30$^\circ$, a mean of
# `r round(100*stat2_T,1)`% of growing days were limited by temperature,
# `r round(100*stat2_Q,1)`% were limited by nutrients, and
# `r round(100*stat2_I,1)`% were limited by light.
```

Factors limiting growth of _A. armata_ in February (A), May (B), August (C) and November (D). Colours indicate limiting factors as temperature (pink), internal nutrient quotient (green) or light availability (gold). Maps show the dominant limiting factor in each cell within the month, while historgram shows the frequency  of limiting factors across all days of the month in all cells.

```{r fig-side-by-side-display-armata}
#| fig-height: 3.5
#| fig-width: 6.25

plot_grid(
  p_dom_map.1 %+% armata_lims[[2]],
  p_dom_hist.1 %+% forhist_armata[[2]],
  ncol = 2, 
  align = "h", 
  rel_widths = c(1, 0.625)
)  

plot_grid(
  p_dom_map.1 %+% armata_lims[[5]],
  p_dom_hist.1 %+% forhist_armata[[5]],
  ncol = 2, 
  align = "h", 
  rel_widths = c(1, 0.625)
)

plot_grid(
  p_dom_map.1 %+% armata_lims[[8]],
  p_dom_hist.1 %+% forhist_armata[[8]],
  ncol = 2, 
  align = "h", 
  rel_widths = c(1, 0.625)
)

plot_grid(
  p_dom_map.1 %+% armata_lims[[11]],
  p_dom_hist.1 %+% forhist_armata[[11]],
  ncol = 2, 
  align = "h", 
  rel_widths = c(1, 0.625)
)
```

### _Asparagopsis taxiformis_

```{r taxiformis lims stats}
#| eval: false

stat1 <- taxiformis_lims %>% 
  filter(latitude > -30) 
stat1 <- stat1 %>% 
  group_by(cell_no, lim) %>% 
  reframe(perc_days = n()/nrow(stat1))
stat1_T <- sum(stat1$perc_days[stat1$lim == "T_lim"])
stat1_Q <- sum(stat1$perc_days[stat1$lim == "Q_lim"])
stat1_I <- sum(stat1$perc_days[stat1$lim == "I_lim"])

stat2 <- taxiformis_lims %>% 
  filter(latitude <= -30)
stat1 <- stat1 %>% 
  group_by(cell_no, lim) %>% 
  reframe(perc_days = n()/nrow(stat1))
stat2_T <- sum(stat2$perc_days[stat2$lim == "T_lim"])
stat2_Q <- sum(stat2$perc_days[stat2$lim == "Q_lim"])
stat2_I <- sum(stat2$perc_days[stat2$lim == "I_lim"])

# For _A. taxiformis_ grown at latitudes > -30$^\circ$, a mean of
# `r round(100*stat1_T,1)`% of growing days were limited by temperature,
# `r round(100*stat1_Q,1)`% were limited by nutrients, and
# `r round(100*stat1_I,1)`% were limited by light.
# 
# For _A. taxiformis_ grown at latitudes < -30$^\circ$, a mean of
# `r round(100*stat2_T,1)`% of growing days were limited by temperature,
# `r round(100*stat2_Q,1)`% were limited by nutrients, and
# `r round(100*stat2_I,1)`% were limited by light.
```

Factors limiting growth of _A. taxiformis_ in February (A), May (B), August (C) and November (D). Colours indicate limiting factors as temperature (pink), internal nutrient quotient (green) or light availability (gold). Maps show the dominant limiting factor in each cell within the month, while historgram shows the frequency  of limiting factors across all days of the month in all cells.

```{r fig-side-by-side-display-taxiformis}
#| fig-height: 3.5
#| fig-width: 6.25

plot_grid(
  p_dom_map.1 %+% taxiformis_lims[[2]],
  p_dom_hist.1 %+% forhist_taxiformis[[2]],
  ncol = 2, 
  align = "h", 
  rel_widths = c(1, 0.625)
)  

plot_grid(
  p_dom_map.1 %+% taxiformis_lims[[5]],
  p_dom_hist.1 %+% forhist_taxiformis[[5]],
  ncol = 2, 
  align = "h", 
  rel_widths = c(1, 0.625)
)  

plot_grid(
  p_dom_map.1 %+% taxiformis_lims[[8]],
  p_dom_hist.1 %+% forhist_taxiformis[[8]],
  ncol = 2, 
  align = "h", 
  rel_widths = c(1, 0.625)
)  

plot_grid(
  p_dom_map.1 %+% taxiformis_lims[[11]],
  p_dom_hist.1 %+% forhist_taxiformis[[11]],
  ncol = 2, 
  align = "h", 
  rel_widths = c(1, 0.625)
)  
```

