---
title: "Results: Supplemented growth and bioremediation"
author: Tormey Reimer
date: last-modified
---

```{r setup}
#| eval: true
#| include: false
#| cache: false

# You need quarto to generate this markdown - install here: https://quarto.org/docs/get-started/
# This markdown uses TinyTex for PDF rendering - install with tinytex::install_tinytex()
# This markdown uses the acronyms extension - install by entering quarto add rchaput/acronyms@master into the terminal

library(stringr)
library(units)
library(targets)
library(tarchetypes)
library(macrogrow)
library(ozmaps)
library(arrow)
library(gganimate)
library(magrittr)
library(ggplot2)
library(dplyr)
library(terra)
library(tidyterra)
library(here)
library(lubridate)
library(tidyr)
library(cowplot)
library(purrr)
library(patchwork)
library(conflicted)
library(gridExtra)
library(grid)

conflicts_prefer(dplyr::select, dplyr::filter)
options(pillar.sigfig = 5)

source(here("R_scripts/00_functions.R"))
source(here("R_scripts/00_global_displays.R"))
```

```{r cell-coords}
#| eval: true
#| code-summary: |
#|   Load in all the BARRA-R2 cells being used, with their coordinates.

cell_coords <- find_read(envi_data, "R2_cell_coords") %>% 
  mutate(state = factor(state, levels = states_ord)) %>% 
  dplyr::select(-layer)
cells_included <- find_read(runs_data, "include")
cells_omitted <- find_read(runs_data, "omit")

totals <- cell_coords %>% 
  filter(!cell_no %in% cells_omitted) %>% 
  group_by(state) %>% 
  reframe(cells = n())
```

```{r normal-growth-end-data}
#| code-summary: |
#|   Get cell growth in ambient (normal) conditions

norm_growth_end <- file.path(runs_data, "cell_growth_end.parquet") %>% 
  read_parquet(mmap = T) %>% 
  merge(cell_coords, by = "cell_no")

# Total N removed at each harvest
norm_growth_end <- norm_growth_end %>% 
  mutate(
    value = TN_end - TN_start,
    value_na = case_when(value <= 0 ~ NA, T ~ value),
    value_0 = case_when(value <= 0 ~ 0, T ~ value),
    month = as.factor(month(parse_date_time(start, orders = "j"))),
    species = factor(species, levels = c("armata", "taxiformis"), labels = c("A. armata", "A. taxiformis"))
  )

norm_growth_total <- norm_growth_end %>% 
  mutate(
    value_na = case_when(value <= 0 ~ NA, T ~ value),
    value_0 = case_when(value <= 0 ~ 0, T ~ value)
  ) %>% 
  group_by(state, species, cell_no, latitude, longitude) %>% 
  reframe(
    total = sumna(value),            # good growth months may be outweighed by bad growth months
    total_na = sumna(value_na),      # bad growth months are removed
    total_na = case_when(total_na == 0 ~ NA, T ~ total_na),
    total_0 = sumna(value_0)       # bad growth months count as zeros
  )
```

```{r supp-growth-end-data}
#| code-summary: |
#|   Get cell growth in nitrogen supplemented conditions

supp_growth_end_raw <- file.path(runs_data, "cell_growth_supp_end.parquet") %>% 
  read_parquet(mmap = T) %>% 
  merge(cell_coords, by = "cell_no")

supp_growth_end <- supp_growth_end_raw %>% 
  mutate(
    value = TN_end - TN_start,
    value_na = case_when(value <= 0 ~ NA, T ~ value),
    value_0 = case_when(value <= 0 ~ 0, T ~ value),
    month = as.factor(month(parse_date_time(start, orders = "j"))),
    species = factor(species, levels = c("armata", "taxiformis"), labels = c("A. armata", "A. taxiformis"))
  )

supp_growth_total <- supp_growth_end %>% 
  mutate(
    value_na = case_when(value <= 0 ~ NA, T ~ value),
    value_0 = case_when(value <= 0 ~ 0, T ~ value)
  ) %>% 
  group_by(state, species, cell_no, latitude, longitude) %>% 
  reframe(
    total = sumna(value),            # good growth months may be outweighed by bad growth months
    total_na = sumna(value_na),      # bad growth months are removed
    total_na = case_when(total_na == 0 ~ NA, T ~ total_na),
    total_0 = sumna(value_0)       # bad growth months count as zeros
  )
```

```{r all-growth-end-data}
#| code-summary: |
#|   Join them together for comparison

supplement_change <- left_join(
  norm_growth_end %>% 
    select(-c("TN_start", "TN_end", "start", "value")) %>% 
    rename(
      norm_success = success, 
      norm_rem = TN_rem, 
      norm_value_na = value_na,
      norm_value_0 = value_0
      ), 
  supp_growth_end %>% 
    select(-c("TN_start", "TN_end", "start", "value")) %>% 
    rename(
      supp_success = success, 
      supp_rem = TN_rem,
      supp_value_na = value_na,
      supp_value_0 = value_0
      ), 
  by = c("state", "cell_no", "longitude", "latitude", "species", "month")
) %>% 
  relocate(state, .before = cell_no) %>% 
  relocate(species, .before = state) %>% 
  relocate(longitude, .after = cell_no) %>% 
  relocate(latitude, .after = longitude) %>% 
  relocate(month, .after = latitude)

supplement_change_total <- left_join(
  norm_growth_total %>% 
    rename(
      norm_total = total, 
      norm_total_na = total_na, 
      norm_total_0 = total_0
      ), 
  supp_growth_total %>% 
    rename(
      supp_total = total, 
      supp_total_na = total_na, 
      supp_total_0 = total_0
      ), 
  by = c("state", "cell_no", "longitude", "latitude", "species")
)
```

# Supplemented success change

Provide summaries of the changes in success (ie whether the modelled seaweed was able to successfully growth at all within a cell) between normal and supplemented conditions.

```{r supp-growth-success-change-1}
#| code-summary: How many/what proportion of cells became successful in each month (and overall) because of supplementation

tc <- supplement_change %>% pull(cell_no) %>% unique() %>% length()

# Which cells changed in success (per month)
norm_success <- supplement_change %>% 
  filter(norm_success) %>% 
  group_by(species, month) %>% 
  reframe(
    norm_success_cells = n(),
    perc_total_cells = 100*n()/tc
  )
success_change <- supplement_change %>% 
  filter(supp_success != norm_success) %>% 
  group_by(species, month) %>% 
  reframe(success_change_cells = n())

norm_success %>% 
  left_join(success_change, by = c("species", "month")) %>% 
  mutate(
    percdiff_change = round(100*success_change_cells/norm_success_cells, 1),
    perctotal_change = round(100*success_change_cells/tc, 1)
  ) %>% 
  select(-norm_success_cells, -success_change_cells) %>% 
  print(n = 24)

# Which cells changed in success (overall)
norm_success <- supplement_change %>% 
  filter(norm_success) %>% 
  group_by(species) %>% 
  reframe(
    norm_success_cells = n(),
    perc_total_cells = 100*n()/(12*tc)
  )
success_change <- supplement_change %>% 
  filter(supp_success != norm_success) %>% 
  group_by(species) %>% 
  reframe(success_change_cells = n())

norm_success %>% 
  left_join(success_change, by = c("species")) %>% 
  mutate(
    percdiff_change = round(100*success_change_cells/norm_success_cells, 1),
    perctotal_change = round(100*success_change_cells/(tc*12), 1)
  ) %>% 
  select(-norm_success_cells, -success_change_cells)
```

```{r area-success-change}
#| code-summary: Calculate the area change (in km2) due to supplementation

library(terra)

rdata_norm <- norm_growth_end %>% 
  select(cell_no, species, success, longitude, latitude, month) %>% 
  group_by(species, cell_no, longitude, latitude) %>% 
  reframe(
    success = as.numeric(any(success)),
    success = case_when(success == 0 ~ NA, T ~ 1)
  )

rdata_norm_arma <- rdata_norm %>% filter(species == "A. armata")
rdata_norm_taxi <- rdata_norm %>% filter(species != "A. armata")

rdata_supp <- supp_growth_end %>% 
  select(cell_no, species, success, longitude, latitude, month) %>% 
  group_by(species, cell_no, longitude, latitude) %>% 
  reframe(
    success = as.numeric(any(success)),
    success = case_when(success == 0 ~ NA, T ~ 1)
  )

rdata_supp_arma <- rdata_supp %>% filter(species == "A. armata")
rdata_supp_taxi <- rdata_supp %>% filter(species != "A. armata")

r_template <- rast(
  extent = ext(c(min(rdata_norm$longitude), max(rdata_norm$longitude), 
                 min(rdata_norm$latitude), max(rdata_norm$latitude))), 
  resolution = 0.12
)

area_success_all <- list()

r <- rasterize(rdata_norm_arma[, c("longitude", "latitude")], r_template, values = rdata_norm_arma$success)
r_area <- cellSize(r, unit = "km")
r_area <- mask(r_area, r)
area_success_all[[1]] <- data.frame(
  species = "A. armata", treatment = "norm", 
  area_km2 = global(r_area, fun = sumna)[[1]]
)

r <- rasterize(rdata_supp_arma[, c("longitude", "latitude")], r_template, values = rdata_supp_arma$success)
r_area <- cellSize(r, unit = "km")
r_area <- mask(r_area, r)
area_success_all[[2]] <- data.frame(
  species = "A. armata", treatment = "supp", 
  area_km2 = global(r_area, fun = sumna)[[1]]
)

r <- rasterize(rdata_norm_taxi[, c("longitude", "latitude")], r_template, values = rdata_norm_taxi$success)
r_area <- cellSize(r, unit = "km")
r_area <- mask(r_area, r)
area_success_all[[3]] <- data.frame(
  species = "A. taxiformis", treatment = "norm", 
  area_km2 = global(r_area, fun = sumna)[[1]]
)

r <- rasterize(rdata_supp_taxi[, c("longitude", "latitude")], r_template, values = rdata_supp_taxi$success)
r_area <- cellSize(r, unit = "km")
r_area <- mask(r_area, r)
area_success_all[[4]] <- data.frame(
  species = "A. taxiformis", treatment = "supp", 
  area_km2 = global(r_area, fun = sumna)[[1]]
)

area_success_all %>% 
  bind_rows() %>% 
  pivot_wider(names_from = treatment, names_prefix = "km2_", values_from = area_km2) %>% 
  mutate(
    km2_diff = km2_supp - km2_norm,
    perc_diff = round(100*km2_diff/km2_norm, 1)
  )
```

```{r supp-growth-success-change-2}
#| code-summary: Find out where cells changed from failure to success under supplemented conditions

supplement_change %>% 
  mutate(success_change = case_when(supp_success != norm_success ~ T, T ~ F)) %>% 
  filter(success_change & species == "A. armata") %>% 
  ggplot(aes(x = longitude, y = latitude, fill = success_change)) +
  geom_tile() +
  geom_sf(data = ozmap_data(data = "states"), inherit.aes = F) +
  coord_sf(
    xlim = c(111.5, 155),
    ylim = c(-44.5, -8.5),
    expand = F
  ) +
  theme_void() + theme(legend.position = "none") +
  facet_wrap(facets = vars(month), nrow = 4) +
  ggtitle("Cells where A. armata became successful in each month under supplementation")

supplement_change %>% 
  mutate(success_change = case_when(supp_success != norm_success ~ T, T ~ F)) %>% 
  filter(success_change & species == "A. taxiformis") %>% 
  ggplot(aes(x = longitude, y = latitude, fill = success_change)) +
  geom_tile() +
  geom_sf(data = ozmap_data(data = "states"), inherit.aes = F) +
  coord_sf(
    xlim = c(111.5, 155),
    ylim = c(-44.5, -8.5),
    expand = F
  ) +
  theme_void() + theme(legend.position = "none") +
  facet_wrap(facets = vars(month), nrow = 4) +
  ggtitle("Cells where A. taxiformis became successful in each month under supplementation")
```

# Supplemented growth and nitrogen removal

```{r supp-total-monthly-means}
#| code-summary: |
#|   Get mean nitrogen removed per each state/month

supp_growth_month <- supp_growth_end %>% 
  group_by(species, state, month) %>% 
  reframe(
    mean = round(meanna(value_0), 1) %>% set_units("mg m-2"),
    sd = round(sdna(value_0), 1) %>% set_units("mg m-2")
  )
  
supp_growth_month %>% 
  filter(species == "A. armata") %>% 
  select(-sd) %>% 
  pivot_wider(names_from = state, values_from = mean)

supp_growth_month %>% 
  filter(species == "A. armata") %>% 
  select(-mean) %>% 
  pivot_wider(names_from = state, values_from = sd, names_prefix = "sd_")

supp_growth_month %>% 
  filter(species != "A. armata") %>% 
  select(-sd) %>% 
  pivot_wider(names_from = state, values_from = mean)

supp_growth_month %>% 
  filter(species != "A. armata") %>% 
  select(-mean) %>% 
  pivot_wider(names_from = state, values_from = sd, names_prefix = "sd_")
```

```{r supp-total-annual-means}
#| code-summary: Get mean monthly nitrogen removed in each state 

supp_growth_annual <- supp_growth_end %>% 
  group_by(species, state) %>% 
  reframe(
    mean = round(meanna(value_0), 1) %>% set_units("mg m-2"),
    sd = round(sdna(value_0), 1) %>% set_units("mg m-2")
  )

supp_growth_annual %>% 
  arrange(species, -mean)
```

```{r supp-total-annual-removal}
#| code-summary: Get total removed in each state 

supp_growth_total <- supp_growth_end %>% 
  group_by(species, state, cell_no) %>% 
  reframe(total = sum(value_0)) %>% 
  group_by(species, state) %>% 
  reframe(total = round(mean(total), 1)) %>% 
  arrange(species, -total)

supp_growth_total
```

```{r fig-supp-growth-end}
plot_data <- supp_growth_month %>% 
  mutate(
    mean = drop_units(mean),
    sd = drop_units(sd),
    mean = case_when(is.na(mean) ~ 0, T ~ mean),
    sd = case_when(sd > mean ~ mean, is.na(sd) ~ 0, T ~ sd),
    state = factor(state, levels = states_ord, labels = states_lng),
    month = as.integer(month)
  ) 
  
plot_data %>%
  ggplot(aes(
      x = month, y = mean, ymin = mean - sd, ymax = mean + sd,
      color = state, fill = state, linetype = species
  )) +
  geom_line(linewidth = 0.75) +
  geom_ribbon(alpha = 0.25, linewidth = 0.25) +
  facet_wrap(facets = vars(state), ncol = 2) +
  # scale_y_continuous(breaks = seq(0, 150, 25), limits = c(0, 175), expand = c(0,0)) +
  scale_color_manual(values = states_pal) +
  scale_fill_manual(values = states_pal) +
  env_plot() +
  labs(x = "Day of the year", y = expression("Total N removed (mg m"^-2 *")"))
```

```{r fig-supp-growth-end-map-totals}
#| fig-cap: Total nitrogen removed by _A. armata_ (top) and _A. taxiformis_ (bottom) across the entire year under supplemented conditions.
#| eval: false

supp_growth_total %>%
  ggplot(aes(x = longitude, y = latitude, fill = total_na)) +
  geom_tile() +
  geom_sf(data = ozmap_data(data = "states"), inherit.aes = F) +
  coord_sf(xlim = c(111.5, 155), ylim = c(-44.5, -8.5), expand = F) +
  scale_x_continuous(breaks = seq(100, 160, 10)) +
  scale_y_continuous(breaks = seq(-2.5, -45, -5)) +
  scale_fill_viridis_c(
    limits = c(0, 10000),
    breaks = seq(0, 10000, 1500),
    na.value = "grey",
    guide = guide_colorbar(
      title = expression("Total N removed (mg m"^-2 * ")"),
      barheight = 1,
      barwidth = 20,
      title.position = "top",
      title.hjust = 0.5
    )
  ) +
  labs(x = "Longitude", y = "Latitude") +
  prettyplot() +
  theme(legend.position = "top", strip.text = element_blank(), strip.background = element_blank()) +
  facet_grid(rows = vars(species))
```

## Difference in nitrogen removal from ambient

```{r diff-total-monthly-means}
#| code-summary: Get difference in mean nitrogen removed per each state/month

norm_growth_month <- norm_growth_end %>% 
  group_by(species, state, month) %>% 
  reframe(
    mean = meanna(value_0) %>% set_units("mg m-2"),
    sd = sdna(value_0) %>% set_units("mg m-2")
  )
  
diff_growth_month <- left_join(
  norm_growth_month %>% select(-sd) %>% rename(norm_mean = mean),
  supp_growth_month %>% select(-sd) %>% rename(supp_mean = mean),
  by = c("species", "state", "month")
) %>% 
  mutate(
    perc_diff = round(100*(supp_mean-norm_mean)/norm_mean, 1),
    times = round(supp_mean/norm_mean, 1)
  )

diff_growth_month %>% 
  select(species, state, month, perc_diff) %>% 
  pivot_wider(names_from = state, values_from = perc_diff, names_prefix = "pd_") %>% 
  print(n=24)

diff_growth_month %>% 
  select(species, state, month, times) %>% 
  pivot_wider(names_from = state, values_from = times, names_prefix = "t_") %>% 
  print(n=24)
```

```{r diff-total-annual-means}
#| code-summary: Get difference in mean monthly nitrogen removed in each state 

norm_growth_annual <- norm_growth_end %>% 
  group_by(species, state) %>% 
  reframe(
    mean = meanna(value_0) %>% set_units("mg m-2"),
    sd = sdna(value_0) %>% set_units("mg m-2")
  )

left_join(
  norm_growth_annual %>% select(-sd) %>% rename(norm_mean = mean),
  supp_growth_annual %>% select(-sd) %>% rename(supp_mean = mean),
  by = c("species", "state")
) %>% 
  mutate(
    perc_diff = round(100*(supp_mean-norm_mean)/norm_mean, 1),
    times = round(supp_mean/norm_mean, 1)
  ) %>% 
  arrange(species, -times)
```

```{r diff-total-annual-removal}
#| code-summary: Get total removed in each state 

options(pillar.sigfig = 5)

norm_growth_total <- norm_growth_end %>% 
  group_by(species, state, cell_no) %>% 
  reframe(total = sum(value_0)) %>% 
  group_by(species, state) %>% 
  reframe(total = mean(total)) %>% 
  arrange(species, -total)

left_join(
  norm_growth_total %>% rename(norm_total = total),
  supp_growth_total %>% rename(supp_total = total),
  by = c("species", "state")
  ) %>% 
  mutate(
    perc_diff = round(100*(supp_total-norm_total)/norm_total, 1),
    times = round(supp_total/norm_total, 1)
  ) %>% 
  arrange(species, -times)
```

```{r fig-compare-growth-map}
#| fig-cap: Total nitrogen removed by _A. armata_ (top) and _A. taxiformis_ (bottom) across the entire year under normal conditions (left) and supplemented conditions (right).
#| eval: false

all_growth_total <- rbind(
  norm_growth_total %>% mutate(scenario = "norm_conditions"),
  supp_growth_total %>% mutate(scenario = "supp_conditions")
)

# hist(all_growth_total$total_na[all_growth_total$scenario == "norm_conditions"])
# hist(all_growth_total$total_na[all_growth_total$scenario != "norm_conditions"])
# minna(all_growth_total$total_na)
# maxna(all_growth_total$total_na)

all_growth_total %>%
  ggplot(aes(x = longitude, y = latitude, fill = total_na)) +
  geom_tile() +
  geom_sf(data = ozmap_data(data = "states"), inherit.aes = F) +
  coord_sf(xlim = c(111.5, 155), ylim = c(-44.5, -8.5), expand = F) +
  scale_x_continuous(breaks = seq(100, 160, 10)) +
  scale_y_continuous(breaks = seq(-2.5, -45, -5)) +
  scale_fill_viridis_c(
    limits = c(1, 10000),
    labels = scales::label_number(),
    trans = "log10",
    na.value = "grey",
    guide = guide_colorbar(
      title = expression("Total N removed (mg m"^-2 * ")"),
      barheight = 1,
      barwidth = 20,
      title.position = "top",
      title.hjust = 0.5
    )
  ) +
  labs(x = "Longitude", y = "Latitude") +
  facet_grid(rows = vars(species), cols = vars(scenario)) +
  prettyplot() +
  theme(legend.position = "top", strip.text = element_blank(), strip.background = element_blank())
```

### Focusing ONLY on removal changes (not range expansion)

```{r diff-total-monthly-means-no-range-exp}
#| code-summary: Get difference in mean nitrogen removed per each state/month, discounting cells that were not successful under normal conditions

cell_filter <- norm_growth_end %>% 
  distinct(cell_no, species, month, success) %>% 
  rename(include = success)

norm_growth_month_nre <- norm_growth_end %>% 
  left_join(cell_filter, by = c("cell_no", "species", "month")) %>% 
  filter(include) %>% 
  group_by(species, state, month) %>% 
  reframe(
    mean = meanna(value_0) %>% set_units("mg m-2"),
    sd = sdna(value_0) %>% set_units("mg m-2")
  )

supp_growth_month_nre <- supp_growth_end %>% 
  left_join(cell_filter, by = c("cell_no", "species", "month")) %>% 
  filter(include) %>% 
  group_by(species, state, month) %>% 
  reframe(
    mean = meanna(value_0) %>% set_units("mg m-2"),
    sd = sdna(value_0) %>% set_units("mg m-2")
  )
  
diff_growth_month <- left_join(
  norm_growth_month_nre %>% select(-sd) %>% rename(norm_mean = mean),
  supp_growth_month_nre %>% select(-sd) %>% rename(supp_mean = mean),
  by = c("species", "state", "month")
) %>% 
  mutate(
    perc_diff = round(100*(supp_mean-norm_mean)/norm_mean, 1),
    times = round(supp_mean/norm_mean, 1)
  )

diff_growth_month %>% 
  select(species, state, month, perc_diff) %>% 
  pivot_wider(names_from = state, values_from = perc_diff, names_prefix = "pd_") %>% 
  print(n=24)

diff_growth_month %>% 
  select(species, state, month, times) %>% 
  pivot_wider(names_from = state, values_from = times, names_prefix = "t_") %>% 
  print(n=24)
```

```{r diff-total-annual-means-no-range-exp}
#| code-summary: Get difference in mean monthly nitrogen removed in each state, discounting cells that were not successful under normal conditions

norm_growth_annual_nre <- norm_growth_end %>% 
  left_join(cell_filter, by = c("cell_no", "species", "month")) %>% 
  filter(include) %>% 
  group_by(species, state) %>% 
  reframe(
    mean = meanna(value_0) %>% set_units("mg m-2"),
    sd = sdna(value_0) %>% set_units("mg m-2")
  )

supp_growth_annual_nre <- supp_growth_end %>% 
  left_join(cell_filter, by = c("cell_no", "species", "month")) %>% 
  filter(include) %>% 
  group_by(species, state) %>% 
  reframe(
    mean = meanna(value_0) %>% set_units("mg m-2"),
    sd = sdna(value_0) %>% set_units("mg m-2")
  )

left_join(
  norm_growth_annual_nre %>% select(-sd) %>% rename(norm_mean = mean),
  supp_growth_annual_nre %>% select(-sd) %>% rename(supp_mean = mean),
  by = c("species", "state")
) %>% 
  mutate(
    norm_mean = round(norm_mean, 1),
    supp_mean = round(supp_mean, 1),
    perc_diff = round(100*(supp_mean-norm_mean)/norm_mean, 1),
    times = round(supp_mean/norm_mean, 1)
  ) %>% 
  arrange(species, -times)
```

```{r diff-total-annual-removal-no-range-exp}
#| code-summary: Get total removed in each state, discounting cells that were not successful under normal conditions 

norm_growth_total_nre <- norm_growth_end %>% 
  left_join(cell_filter, by = c("cell_no", "species", "month")) %>% 
  filter(include) %>% 
  group_by(species, state, cell_no) %>% 
  reframe(total = sum(value_0)) %>% 
  group_by(species, state) %>% 
  reframe(total = mean(total)) %>% 
  arrange(species, -total)

supp_growth_total_nre <- supp_growth_end %>% 
  left_join(cell_filter, by = c("cell_no", "species", "month")) %>% 
  filter(include) %>% 
  group_by(species, state, cell_no) %>% 
  reframe(total = sum(value_0)) %>% 
  group_by(species, state) %>% 
  reframe(total = mean(total)) %>% 
  arrange(species, -total)

left_join(
  norm_growth_total_nre %>% rename(norm_total = total),
  supp_growth_total_nre %>% rename(supp_total = total),
  by = c("species", "state")
  ) %>% 
  mutate(
    norm_total = round(norm_total, 1),
    supp_total = round(supp_total, 1),
    perc_diff = round(100*(supp_total-norm_total)/norm_total, 1),
    times = round(supp_total/norm_total, 1)
  ) %>% 
  arrange(species, -times)
```

# Supplemented bioremediation efficiency

```{r available-nitrogen}
#| code-summary: Get total nitrogen made available in supplemented conditions

N_input <- file.path(envi_data, "cell_input_supp") %>% 
  list.files(full.names = T) %>% 
  purrr::map(function(fnm) {
    fnm %>% arrow::read_parquet(col_select = c("cell_no", "state", "yday", "Ni_input", "Am_input"))
  }) %>% 
  bind_rows() %>% 
  mutate(
    N_input = 5.5*(Ni_input + Am_input),
    month = month(make_date(2023, 12, 31) + days(yday)) %>% as.factor()
    ) %>% 
  group_by(state, cell_no, month) %>% 
  reframe(
    N_input = sum(N_input)
  ) %>% 
  mutate(
    N_input = set_units(N_input, "mg m-2")
  )

biorem_eff <- supp_growth_end %>% 
  select(-contains("TN"), -start, -longitude, -latitude) %>% 
  left_join(N_input, by = c("state", "cell_no", "month")) %>% 
  mutate(
    value_na = set_units(value_na, "mg m-2"),
    biorem_eff = value_na/N_input
  )
```

```{r total-removed-ofsupp-monthly}
#| code-summary: Total nitrogen removed of the supplemented nitrogen available

biorem_eff %>% 
  group_by(species, month) %>% 
  reframe(
    mean_biorem_eff = round(100*meanna(biorem_eff), 1),
    max_biorem_eff = round(100*maxna(biorem_eff), 1)
  ) %>% 
  print(n = 24)

biorem_eff %>% 
  mutate(biorem_eff = 100*biorem_eff) %>% 
  ggplot(aes(x = biorem_eff, fill = species, colour = species)) +
  geom_histogram(position = "identity", alpha = 0.35, binwidth = 0.1) +
  theme_classic()
```

```{r total-removed-ofsupp-states}
#| code-summary: Total nitrogen removed of the ambient nitrogen available (monthly, in each state)

biorem_eff %>% 
  group_by(species, state) %>% 
  reframe(
    mean_biorem_eff = round(100*meanna(biorem_eff), 1),
    max_biorem_eff = round(100*maxna(biorem_eff), 1)
  ) %>% 
  print(n = 24)

biorem_eff %>% 
  group_by(species, state, month) %>% 
  reframe(
    mean_biorem_eff = round(100*meanna(biorem_eff), 1)
  ) %>% 
  pivot_wider(names_from = state, values_from = mean_biorem_eff) %>% 
  print(n = 24)
```

```{r total-removed-ofambi-annual}
#| code-summary: Total nitrogen removed of the ambient nitrogen available across the entire year

biorem_eff_total <- supp_growth_end %>% 
  select(-contains("TN"), -start, -longitude, -latitude) %>% 
  left_join(N_input, by = c("state", "cell_no", "month")) %>% 
  mutate(
    N_input = drop_units(N_input)
  ) %>% 
  group_by(species, state, cell_no) %>% 
  reframe(
    value_na = sumna(value_na),
    N_input = sumna(N_input)
  ) %>% 
  mutate(
    value_na = set_units(value_na, "mg m-2"),
    N_input = set_units(N_input, "mg m-2"),
    biorem_eff = value_na/N_input
    )

biorem_eff %>% 
  group_by(species) %>% 
  reframe(
    mean_biorem_eff = round(100*meanna(biorem_eff), 1),
    max_biorem_eff = round(100*maxna(biorem_eff), 1)
  )

biorem_eff_total %>% 
  mutate(biorem_eff = 100*biorem_eff) %>% 
  ggplot(aes(x = biorem_eff, fill = species, colour = species)) +
  geom_histogram(position = "identity", alpha = 0.35, binwidth = 1) +
  theme_classic()
```


