---
title: (Needs change) Investigating the bioremediation potential of seaweed aquaculture across Australia
author: 
  - name: Tormey E. Reimer
    orcid: 0000-0002-5983-2798
    corresponding: true
    email: tormey.reimer@utas.edu.au
    role: 
      - conceptualization: lead
      - methodology: lead
      - analysis: lead
      - drafting: lead
    affil-id: [1,2]
  - name: Katie Cresswell
    orcid: 0000-0003-1692-4964
    corresponding: false
    roles: 
      - conceptualization: supporting
      - methodology: supporting
      - drafting: supporting
    affil-id: [1]
  - name: Kate Fraser
    orcid: 0000-0002-3057-5257
    corresponding: false
    roles: 
      - methodology: supporting
      - drafting: supporting
    affil-id: [3]
  - name: Camille White
    orcid: 0000-0003-2061-6462
    corresponding: false
    roles: 
      - methodology: supporting
      - analysis: supporting
      - editing: supporting
    affil-id: [3]
  - name: Scott A. Hadley
    orcid: 0000-0001-9925-6804
    corresponding: false
    roles: 
      - methodology: supporting
      - analysis: supporting
      - editing: supporting
    affil-id: [3]
affiliations:
  - id: 1
    name: Institute for Marine and Antarctic Studies
    group: Fisheries & Aquaculture
    address: 29 Nubeena Crescent
    city: Taroona
    region: TAS
    country: Australia
    postal-code: 7053
    url: https://www.imas.utas.edu.au/
  - id: 2
    name: Centre for Marine Socioecology
    url: https://marinesocioecology.org/
  - id: 3
    name: Institute for Marine and Antarctic Studies
    group: Sustainable Marine Research Collaboration
    address: 29 Nubeena Crescent
    city: Taroona
    region: TAS
    country: Australia
    postal-code: 7053
    url: https://www.imas.utas.edu.au/
keywords:
  - keyword
  - keyword
abstract: "Seaweed aquaculture can be a powerful tool for an ecosystem-based approach to nutrient bioremediation. In Australia, where essential nutrients (particularly nitrogen) are naturally low, anthropogenic nutrient inputs can have discernible impacts on the environment. There is currently a large amount of interest in how seaweed aquaculture can mitigate environmental impacts; however, there is currently no large-scale framework guiding the use of seaweed aquaculture for bioremediative purposes in Australia. Few globally farmed seaweed species are native to Australia, and existing culture operations are small-scale and not focused on bioremediation. Modelling approaches can help bridge this gap and provide an adaptable tool for bioremediation assessment and marine spatial planning. This project adapted and parameterised an existing nitrogen-based seaweed growth model for several Australian native species, including species of commercial interest such as *Asparagopsis** spp. Large-scale environmental data were used to model seaweed growth and nitrogen removal potential in coastal waters around Australia, and specific scenarios identified the most suitable conditions for each species. The results of this project can be used by managers, growers and other stakeholders to assess the suitability of sites for growth and bioremediation, and the model itself is freely available for future use and extensions. This project has provided a framework for reducing uncertainty surrounding the potential bioremediation capacity of Australian native seaweed species and highlighted some fundamental gaps that must be addressed in future."

plain-language-summary: |
  Plain language summary goes here....
key-points:
  - Point 1
  - Point 2
date: last-modified
bibliography: [resources/refs.bib, resources/manual-refs.bib]
# citation:
#   container-title: Earth and Space Science

format:
  html:
    theme: simplex
    css: [resources/extra.css]
    code-fold: true
    code-overflow: wrap
    toc: true
    toc-expand: true
    toc-location: left
    toc-depth: 4
    lang: en-GB
    grid:
      sidebar-width: 400px
      body-width: 1200px
      margin-width: 100px
      gutter-width: 2em
  # pdf: 
  #   papersize: a4
  #   linestretch: 1.35
  #   link-citations: true
  #   keep-tex: true
  #   template-partials:
  #     - title.tex
  #   include-in-header:
  #     file: header_extra.tex
        
filters:
  - acronyms

execute:
  eval: true
  cache: true
  fig-height: 6
  fig-width: 4
  freeze: auto

knitr: 
  opts_chunk:
    fig.align: center
    fig.width: 6
---

```{r setup}
#| eval: true
#| include: FALSE

# You need quarto to generate this markdown - install here: https://quarto.org/docs/get-started/
# This markdown uses the acronyms extension - install by entering quarto add rchaput/acronyms@master into the terminal
# This markdown uses TinyTex for PDF rendering - install with tinytex::install_tinytex()

library(stringr)
library(units)
library(targets)
library(tarchetypes)
library(macrogrow)
library(ozmaps)
library(arrow)
library(gganimate)
library(magrittr)
library(ggplot2)
library(dplyr)
library(lubridate)
library(tidyr)
library(conflicted)
conflicts_prefer(dplyr::select(), dplyr::filter(), .quiet = T)
```

```{r global variables and functions}
#| eval: true
#| include: false

base_path <- file.path("C:", "Users", "treimer", "Documents", "R-temp-files", "modelling-Asparagopsis")
# source(file.path(base_path, "R", "other-functions.R"))
cell_store <- file.path(base_path, "targets_outputs", "_spatial_cells")
cell_data <- file.path(base_path, "data", "processed-spatial-cells")
spec_store <- file.path(base_path, "targets_outputs", "_species")
spec_data <- file.path(base_path, "data", "processed-species")
part_store <- file.path(base_path, "targets_outputs", "_partial_lims")
part_data <- file.path(base_path, "data", "processed-partial-lims")
runs_store <- file.path(base_path, "targets_outputs", "_model_running")
runs_data <- file.path(base_path, "data", "processed-model-running")
theo_store <- file.path(base_path, "targets_outputs", "_theo_scens")
theo_data <- file.path(base_path, "data", "processed-theo-scens")

bounds <- tar_read(bbox, store = cell_store)

prettyplot <- theme_classic() +
  theme(text = element_text(family = "sans", size = 12, colour = "black"),
        axis.title = element_text(vjust = 2),
        legend.position = "none")

env_plot <- prettyplot +
  theme(legend.position = "none",
        aspect.ratio = 0.55,
        text = element_text(family = "sans", size = 11, colour = "black"))

nitrogen_lab <- expression("Concentration ("*mu*"M N)")
nitrate_lab <- expression("Nitrate concentration ("*mu*"M NO"[3]*")")
ammonium_lab <- expression("Ammonium concentration ("*mu*"M NH"[4]^+")")

# Allows using "units" package to convert between moles and g of nitrogen ONLY
remove_unit("mol")
install_unit("mol", "14.0067 g")

prettynum <- function(num, sigfig=2) {format(round(as.numeric(num), sigfig), nsmall=1, big.mark=",")}

species_pal_3 <- c("red3", "dodgerblue3", "green3")
species_pal_4 <- c("red3", "pink2", "dodgerblue3", "green3")
depths_pal <- c("green3", "#FF8247", "dodgerblue3", "#292928")
states_pal <- c("#d95f02", "#e6ab02", "#7570b3", "#1b9e77", "#66a61e", "#e7298a", "#a6761d", "#666666")
scens_pal_1 <- c("#292928", "dodgerblue3", "#66a61e", "#e7298a", "#e6ab02")
scens_pal_2 <- c("dodgerblue3", "#66a61e", "#e7298a", "#e6ab02")
states_pal_2 <- c(states_pal, "black")
states <- tar_read(state_levels, store = cell_store)
states_ord <- c("NTE", "QLD", "NSW", "SAU", "TAS", "VIC", "WAN", "WAS")
states_lng <- c("Northern Territory", "Queensland", "New South Wales", "South Australia", "Tasmania", "Victoria", "Western Australia (N)", "Western Australia (S)")
```

# Introduction {#sec-introduction}

Nutrient pollution is increasing globally and can adversely affect marine ecosystem functioning. Excess nitrogen from terrestrial farm run-off, sewage treatment plant outfalls, and marine aquaculture can be particularly impactful, because nitrogen is often a limiting nutrient in coastal ecosystems. Macroalgae species are particularly effective at assimilating dissolved nitrogen (particularly nitrate, NO$_3$, and ammonium, NH$_4^+$) from coastal waters, and farming macroalgae is therefore a potential strategy for bioremediating excess nitrogen and mitigating against associated ecosystem damage [@hadley_investigation_2018].

Macroalgae cultivation is currently a minor industry in Australia, with many uncertainties surrounding its application for bioremediation purposes. One major hurdle has been species selection, as many species farmed globally (e.g. winged kelp *Alaria esculenta*, sugar kelp *Saccharina latissima*) are not native to Australia. However, Australia’s diverse native seaweed taxa provide many promising candidates for both commercial and bioremediative uses. A species’ bioremediative capacity depends on its uptake rate of nitrate and ammonium, its rate of nitrogen assimilation (growth), and its suitability to Australian environmental conditions. Additionally, a species must be suitable for commercial culture in order to reliably bioremediate nutrients at any substantial scale. If the bioremediation potential of promising macroalgae taxa can be estimated across a range of environmental conditions, the seaweed industry can leverage this information to plan cultivation for the purposes of bioremediation in various pollution contexts. 

Nitrogen uptake rates and tolerance to environmental changes vary widely among species, even those which are closely related taxonomically [@hadley_modeling_2015; @reimer_clarifying_2024; @smart_seasonal_2022]. Previous studies have assessed the bioremediative potential of a handful of species at small scales, mostly in Tasmania. Ulva lactuca’s strong preferential uptake of ammonium makes it a good bioremediator in the context of finfish farms with increased flow rate of ammonium waste [@hadley_modeling_2015]. 
Macrocystis pyrifera removes nitrogen very efficiently due to its fast growth and large size, but it has a limited geographic range and a very low tolerance for high temperatures [@reimer_clarifying_2024]. The current study chose species based on a balance of potential industry application and demonstrated bioremediation potential at small scales. 
*Asparagopsis taxiformis*, red seaweeds which have gained much commercial and public attention as an ingredient in feed for cattle, sheep, and other ruminants to substantially decrease the amount of methane they produce [@camer-pesci_opportunities_2023]. 

The current project utilises an established macroalgae growth model to assess bioremediation potential across a broad range of environmental conditions representative of different growing regions in coastal Australian waters, for the species *Asparagopsis armata* and *Asparagopsis taxiformis*. The model estimates uptake rates of nitrate and ammonium and incorporates nitrogen storage capacity, temperature limits, response to light levels, and maximum growth rates in optimal conditions. This approach allows for the identification of strengths and weaknesses specific to bioremediation in different Australian states and to different macroalgae genera, contributing to a comprehensive understanding of the feasibility and implications of this innovative and low-impact approach to nutrient remediation in marine coastal waters.

## Aims {#sec-aims}

The aims of this study were to:

1.	To apply an existing bioremediation model to specified locations across Australia, including Tasmania, New South Wales, Victoria, South Australia, and Western Australia, for *Asparagopsis* and other native macroalgae taxa with potential for cultivation.
2.	To compare bioremediation potential among locations based on different environmental conditions.
3.	To compare the bioremediation potential of several macroalgae taxa using different nutrient sources and different pollution contexts (e.g. marine aquaculture waste, sewage treatment plant runoff).
4.	To identify challenges among locations and macroalgae taxa to achieving full bioremediation potential.

# Methods {#sec-methods}
## Environmental inputs {#sec-envinputs}

```{r cell-coords}
#| eval: true
#| code-summary: |
#|   Load in all the BARRA2 cells being used, with their coordinates.

cell_coords <- tar_read(BARRA_C2_cell_coords, store = cell_store) %>% 
  mutate(state = factor(state, levels = states_ord))

totals <- cell_coords %>% 
  group_by(state) %>% 
  reframe(cells = n())
```

The bioremediation model requires several environmental variables to drive daily macroalgae growth including temperature, light (irradiance), nutrient (nitrogen) concentrations, salinity, water movement, and bathymetry. The waters around Australia were divided into cells based on the BARRA2-R product within the Bureau of Meteorology's \acr{BARRA} dataset [@bureau_of_meteorology_barra_2021; @su_barra_2019]. 
Each cell is 12 km $\times$ 12 km and the cells extend to within ~60 km of the Australian coast (@fig-states-map). This resolution applied to all environmental data in this study. 
There were a total of `r sum(totals$cells)` cells used in this analysis: 
`r as.integer(totals$cells[totals$state == "VIC"])` in Victoria, 
`r as.integer(totals$cells[totals$state == "NSW"])` in New South Wales, 
`r as.integer(totals$cells[totals$state == "QLD"])` in Queensland, 
`r as.integer(totals$cells[totals$state == "SAU"])` in South Australia, 
`r as.integer(totals$cells[totals$state == "WAN"]+totals$cells[totals$state == "WAS"])` in Western Australia (divided into 
`r as.integer(totals$cells[totals$state == "WAN"])` cells in the north and 
`r as.integer(totals$cells[totals$state == "WAS"])` in the south), 
`r as.integer(totals$cells[totals$state == "TAS"])` in Tasmania, and 
`r as.integer(totals$cells[totals$state == "NTE"])` in the Northern Territory. 

```{r fig-states-map}
#| fig-cap: |
#|   Map showing the area included this analysis, with colours indicating the different states. The ~60 km coastline buffer includes 11,709 cells approximately 12 x 12 km each.
#| fig-width: 8
#| fig-height: 7

ggplot(cell_coords, aes(x = lon, y = lat, fill = state)) +
  geom_raster() +
  geom_sf(data = ozmap_data(data = "states"), inherit.aes = F) +
  coord_sf(xlim = c(bounds[["AUS"]]["lonmin"], bounds[["AUS"]]["lonmax"]),
           ylim = c(bounds[["AUS"]]["latmin"], bounds[["AUS"]]["latmax"])) +
  scale_x_continuous(breaks = seq(100, 160, 5)) +
  scale_y_continuous(breaks = seq(-5, -45, -5)) +
  scale_color_brewer(palette = "Dark2") +
  scale_fill_brewer(palette = "Dark2") +
  labs(x = "Longitude", y = "Latitude") +
  prettyplot +
  theme(legend.position = "none")
```

Bathymetry data were sourced from the output of the high-resolution depth model AusBathyTopo [@beaman_ausbathytopo_2023]. Bathymetry data was restricted to ocean topography (< 0 m) and the 250 m grid was mapped onto the BARRA2 12 km grid using weighted means. 

```{r T input data}
#| eval: false

T_input_cell_av <- file.path(cell_data, "T_input_cell_av.parquet") %>% read_parquet()

T_input_cell_av <- T_input_cell_av %>% 
  group_by(state) %>% 
  reframe(mean = mean(value, na.rm = T),
          max = max(value, na.rm = T),
          min = min(value, na.rm = T))

# Sea surface temperature (‘ts’) data were sourced from the BARRA2-R product for 2019-2023 [@su_barra_2019]. Temperature patterns varied across regions, as expected (@fig-T-input). Daily mean temperatures (2019-2023) ranged from 
# `r prettynum(min(T_input_cell_av$mean), 1)`$^{\circ}$C in Tasmania to 
# `r prettynum(max(T_input_cell_av$mean), 1)`$^{\circ}$C in the Northern Territory. 
# Extreme temperatures were recorded at both ends of the spectrum: the highest daily temperature was 
# `r prettynum(max(T_input_cell_av$max), 1)`$^{\circ}$C, in northern Western Australia, and the lowest daily temperature was 
# `r prettynum(min(T_input_cell_av$min), 1)`$^{\circ}$C, in Tasmania. 
```


```{r fig-T-input}
#| fig-height: 8
#| fig-cap: |
#|   Summary of temperature values across the year within each state, averaged across 2019-2023. Bold lines show the mean across all cells within each state while ribbons show the standard deviation between cells.

# # This averages over time and space to create a "typical year" per state - for plotting
# T_input_state_av <- file.path(cell_data, "T_input_state_av.parquet") %>% 
#   read_parquet() %>% 
#   mutate(state = factor(state, levels = states_ord, labels = states_lng))
# 
# p_input_state_av <- T_input_state_av %>% 
#   ggplot(aes(x = yday, y = value, ymin = value-sd, ymax = value+sd, colour = state, fill = state)) + 
#   geom_line(linewidth = 0.75) +
#   geom_ribbon(alpha = 0.35, linewidth = 0.25) +
#   facet_wrap(facets = vars(state), ncol = 2) +
#   scale_y_continuous(breaks = seq(0, 40, 5), limits = c(10, 35))
#   scale_fill_brewer(palette = "Dark2") +
#   scale_color_brewer(palette = "Dark2") +
#   env_plot +
#   labs(x = "Day of the year", y = expression("Temperature ("*degree*"C)"))
# 
# p_input_state_av

ggplot()
```

```{r I input data}
#| eval: true

I_input_cell_av <- file.path(cell_data, "I_input_cell_av.parquet") %>% read_parquet() # For gif making

I_input_cell_av <- I_input_cell_av %>% 
  group_by(state) %>% 
  reframe(mean = mean(value, na.rm = T),
          max = max(value, na.rm = T),
          min = min(value, na.rm = T))

# Sea surface irradiance was estimated using downward short-wave radiation (‘rsdsdir’) data sourced from the BARRA2-R product for 2019-2023 [@su_barra_2019]. Surface irradiance showed strong seasonal variation in the southern states (Tasmania, Victoria, South Australia and southern Western Australia) and less seasonality but more daily variation in the northern states (@fig-I-input). 
# State mean surface irradiance ranged from 
# `r prettynum(min(I_input_cell_av$mean),0)` µmol photons m$^{-2}$ s$^{-1}$ in Tasmania to 
# `r prettynum(max(I_input_cell_av$mean),0)` µmol photons m$^{-2}$ s$^{-1}$ in Queensland. However, the highest daily cell surface irradiance was 
# `r prettynum(max(I_input_cell_av$max),0)` µmol photons m$^{-2}$ s$^{-1}$ in southern Western Australia and the lowest was 
# `r prettynum(min(I_input_cell_av$min),0)` µmol photons m$^{-2}$ s$^{-1}$ in Tasmania. 
```


```{r fig-I-input}
#| fig-height: 8
#| fig-cap: |
#|   Summary of irradiance values across the year within each state, averaged across 2019-2023. Bold lines show the mean across all cells within each state while ribbons show the standard deviation between cells. 

# This averages over time and space to create a "typical year" per state - for plotting
I_input_state_av <- file.path(cell_data, "I_input_state_av.parquet") %>% 
  read_parquet() %>% 
  mutate(state = factor(state, levels = states_ord, labels = states_lng))

p_input_state_av %+% I_input_state_av +
  scale_y_continuous(breaks = seq(0, 1600, 200), limits = c(0,800)) +
  labs(x = "Day of the year", y = expression("Irradiance (photons m"^-2*" s"^-1*")")) 
```

```{r S input data}
#| eval: false

S_input_cell_av <- file.path(cell_data, "S_input_cell_av.parquet") %>% read_parquet() # For gif making
S_input_cell_av_depth <- file.path(cell_data, "S_input_cell_av_depth.parquet") %>% read_parquet()

test_deps <- S_input_cell_av_depth %>% 
  group_by(state, yday, dep) %>% 
  reframe(#sd = sd(value, na.rm = T),
          value = mean(value, na.rm = T)) %>% 
  pivot_wider(names_from = dep, values_from = value, names_prefix = "dep_") %>% 
  mutate(diff1 = abs(dep_2.5-dep_7.5),
         diff2 = abs(dep_12.5-dep_7.5),
         diff3 = abs(dep_12.5-dep_2.5),
         min_diff = min(c(diff1, diff2, diff3)),
         max_diff = max(c(diff1, diff2, diff3)))

S_input_cell_av <- S_input_cell_av %>% 
  group_by(state) %>% 
  reframe(mean = mean(value, na.rm = T),
          max = max(value, na.rm = T),
          min = min(value, na.rm = T))

# Salinity (‘ocean_salt’) data were sourced from the \acr{BRAN} data suite for 2019-2023 [@chamberlain_multiscale_2021; @oke_towards_2013; @oke_bluelink_2008]. \acr{BRAN} data are depth-separated; however, the differences in salinity between layers in surface waters was very small (between `r prettynum(min(test_deps$min_diff),1)` and `r prettynum(max(test_deps$max_diff),1)` g L$^{-1}$), so salinity was depth-averaged over depths < 25 m. 
# Salinity values were very consistent among cells and across all states (@fig-S-input). All states shared the same mean cell daily value of 
# `r prettynum(mean(S_input_cell_av$mean),1)` g L$^{-1}$ and the same minimum daily value of 
# `r prettynum(min(S_input_cell_av$min),1)` g L$^{-1}$. Maximum cell daily value was 
# `r prettynum(max(S_input_cell_av$max),1)` g L$^{-1}$ in New South Wales and southern Western Australia. 
```


```{r fig-S-input}
#| fig-height: 8
#| fig-cap: |
#|   Summary of salinity values across the year within each state, averaged across 2019-2023 and all depths < 25 m. Bold lines show the mean across all cells within each state while ribbons show the standard deviation between cells.

# This averages over time and space to create a "typical year" per state - for plotting
S_input_state_av <- file.path(cell_data, "S_input_state_av.parquet") %>% 
  read_parquet() %>% 
  mutate(state = factor(state, levels = states_ord, labels = states_lng))

p_input_state_av %+% S_input_state_av +
  scale_y_continuous(breaks = seq(30, 40, 0.25), limits = c(34.5, 35.5)) +
  labs(x = "Day of the year", y = expression("Salinity (g L"^-1*")"))
```

```{r UV_input data}
#| eval: true

UV_input_cell_av <- file.path(cell_data, "UV_input_cell_av.parquet") %>% read_parquet() # For gif making
UV_input_cell_av_depth <- file.path(cell_data, "UV_input_cell_av_depth.parquet") %>% read_parquet()

test_deps <- UV_input_cell_av_depth %>% 
  group_by(state, yday, dep) %>% 
  reframe(#sd = sd(value, na.rm = T),
          value = mean(value, na.rm = T)) %>% 
  pivot_wider(names_from = dep, values_from = value, names_prefix = "dep_") %>% 
  mutate(diff1 = abs(dep_2.5-dep_7.5),
         diff2 = abs(dep_12.5-dep_7.5),
         diff3 = abs(dep_12.5-dep_2.5),
         min_diff = min(c(diff1, diff2, diff3)),
         max_diff = max(c(diff1, diff2, diff3)))

# print("Testing deps: min, max")
# min(test_deps$min_diff)
# max(test_deps$max_diff)

UV_input_cell_av <- UV_input_cell_av %>% 
  group_by(state) %>% 
  reframe(mean = mean(value, na.rm = T),
          max = max(value, na.rm = T),
          min = min(value, na.rm = T))
```

Water velocity (‘ocean_u’ and ‘ocean_v’) data were sourced from the \acr{BRAN} data suite for 2019-2023 [@chamberlain_multiscale_2021; @oke_towards_2013; @oke_bluelink_2008]. Water velocity ($\overline{UV} = \sqrt{U^2+V^2}$) showed minimal differences between surface layers and was also depth-averaged over depths < 25 m. Data were very consistent among states but with considerable variation among cells (@fig-UV-input). All states shared similar mean daily values of 
0.13 m s$^{-1}$. Highest daily values were 
0.32 m s$^{-1}$ in southern Western Australia, and minimum daily value was 
0.037 m s$^{-1}$ in Queensland.

```{r fig-UV-input}
#| fig-height: 8
#| fig-cap: |
#|   Summary of water velocity values across the year within each state, averaged across 2019-2023 and all depths < 25 m. Bold lines show the mean across all cells within each state while ribbons show the standard deviation between cells.

# This averages over time and space to create a "typical year" per state - for plotting
UV_input_state_av <- file.path(cell_data, "UV_input_state_av.parquet") %>% 
  read_parquet() %>% 
  mutate(state = factor(state, levels = states_ord, labels = states_lng))

p_I_input_state_av %+% UV_input_state_av +
  scale_y_continuous(breaks = seq(0, 0.50, 0.1), limits = c(-0.01, 0.25)) +
  labs(x = "Day of the year", y = expression("Water velocity (m s"^-1*")"))
```

## Nitrogen {#sec-nitrogen}

```{r N-data sources locations}
# Sys.setenv(TAR_PROJECT = "project_spatial_cells")
# tar_manifest()$name %>% unique() %>% str_subset("refstation") %>% sort()

outfall_locations <- tar_read(outfall_locations, store = cell_store) %>% 
  mutate(state = case_when(
    lon > bounds[["NTE"]]["lonmin"] & lon <= bounds[["NTE"]]["lonmax"] & lat > bounds[["NTE"]]["latmin"] & lat <= bounds[["NTE"]]["latmax"] ~ "NTE",
    lon > bounds[["QLD"]]["lonmin"] & lon <= bounds[["QLD"]]["lonmax"] & lat > bounds[["QLD"]]["latmin"] & lat <= bounds[["QLD"]]["latmax"] ~ "QLD",
    lon > bounds[["NSW"]]["lonmin"] & lon <= bounds[["NSW"]]["lonmax"] & lat > bounds[["NSW"]]["latmin"] & lat <= bounds[["NSW"]]["latmax"] ~ "NSW",
    lon > bounds[["SAU"]]["lonmin"] & lon <= bounds[["SAU"]]["lonmax"] & lat > bounds[["SAU"]]["latmin"] & lat <= bounds[["SAU"]]["latmax"] ~ "SAU",
    lon > bounds[["TAS"]]["lonmin"] & lon <= bounds[["TAS"]]["lonmax"] & lat > bounds[["TAS"]]["latmin"] & lat <= bounds[["TAS"]]["latmax"] ~ "TAS",
    lon > bounds[["VIC"]]["lonmin"] & lon <= bounds[["VIC"]]["lonmax"] & lat > bounds[["VIC"]]["latmin"] & lat <= bounds[["VIC"]]["latmax"] ~ "VIC",
    lon > bounds[["WAN"]]["lonmin"] & lon <= bounds[["WAN"]]["lonmax"] & lat > bounds[["WAN"]]["latmin"] & lat <= bounds[["WAN"]]["latmax"] ~ "WAN",
    lon > bounds[["WAS"]]["lonmin"] & lon <= bounds[["WAS"]]["lonmax"] & lat > bounds[["WAS"]]["latmin"] & lat <= bounds[["WAS"]]["latmax"] ~ "WAS"
    )) %>% 
  mutate(state = factor(state, levels = states_ord)) %>% 
  select(-cell_ID)

refstation_locations <- tar_read(refstation_coords, store = cell_store) %>% 
  mutate(state = NA) %>% 
  rename(name = StationName) %>% 
  relocate(state, .before = name) %>% 
  select(-code)

outfalls <- outfall_locations %>% 
  group_by(state) %>% 
  reframe(outfalls = n())

N_locations <- rbind(
  outfall_locations %>% mutate(type = "outfall"),
  refstation_locations %>% mutate(type = "refstation")
)
```

::: {note}
Revise this section
::: 
Nitrogen inputs (in the form of nitrate and ammonium) are the least precise and most important inputs in the model. This study used the combined biogeochemical parameters taken at the \acr{IMOS} reference stations 
::: {note}
IMOS 2023 reference?
::: 
as a primary source, and supplementary nitrogen data were sourced from the \acr{NOD}
::: {note}
NOD 2025 reference?
::: 
. Each cell was matched to its nearest \acr{IMOS} reference station as a nutrient baseline, while outfalls were matched to multiple cells within a 48 km radius and their influence was considered inversely proportional to the distance between the cell centre and the outfall. @fig-N-data-locations shows the location and distribution of reference stations and outfall recordings. 

```{r fig-N-data-locations}
#| fig-width: 8
#| fig-height: 7
#| fig-cap: Map showing the locations of recorded outfalls (points, coloured by state) and the IMOS reference stations (stars) used in this analysis.

# ggplot(N_locations, aes(x = lon, y = lat, color = state, shape = type)) +
#   geom_sf(data = ozmap_data(data = "states"), inherit.aes = F) +
#   coord_sf(xlim = c(bounds[["AUS"]]["lonmin"], bounds[["AUS"]]["lonmax"]),
#            ylim = c(bounds[["AUS"]]["latmin"], bounds[["AUS"]]["latmax"])) +
#   geom_point(size = 3.5) +
#   scale_shape_manual(values = c(20, 8)) +
#   # scale_colour_manual(values = states_pal, labels = states_lng, na.value = "black") +
#   scale_color_brewer(palette = "Dark2") +
#   scale_x_continuous(breaks = seq(100, 160, 5)) +
#   scale_y_continuous(breaks = seq(-5, -45, -5)) +
#   labs(x = "Longitude", y = "Latitude") +
#   prettyplot +
#   theme(legend.position = "none")

ggplot()
```

Data from the \acr{IMOS} reference stations at Bonney Coast, Esperance and Ningaloo were too sporadic to be useful and their datasets were very similar to their closest neighbour stations, so their data were excluded. This left stations at Darwin, Maria Island, North Stradbroke Island, Port Hacking, Rottnest Island and Yongala for the current analysis. 

```{r}

```

Matching model cells to their nearest reference station resulted in 
2703 cells being matched to Darwin, 
1700 to Kangaroo Island, 
1739 to Maria Island, 
699 to North Stradbroke Island, 
641 to Port Hacking, 
2292 to Rottnest Island, and 
1903 to Yongala.

Outfall data from \acr{NOD} are only recorded as monthly observations of the water being added to coastal conditions, rather than ambient conditions in coastal waters. Data from outfall stations was only included if $\geq$ 20 observations were made within the past 10 years (2014-2023). If $\geq$ 20 observations were made from the period of 2019-2023, then the data were restricted to that period. One outfall station was fully excluded (Pardoe, in northern Tasmania) as the concentrations recorded were ~6 orders of magnitude higher than any other station and it was presumed that this was a data reporting error. While there is at least one outfall in each state, they are much more abundant in Victoria, Tasmania, New South Wales and Queensland. 
There were 
7 outfall stations in South Australia, 
9 in southern Western Australia, 
4 in the Northern Territory, and 
1 in northern Western Australia. 
These two nutrient data sources were combined into a single “typical” year for each nutrient, with outfall data being additionally weighted by the total volume of water recorded from the outfall location relative to a cell’s volume of surface water (12 km $\times$ 12 km $\times$ 10 m). Finally, a small degree of random variation (up to $\pm$10% of standard deviation from the mean) was added to daily values of nitrate and ammonium. This was done to better reflect the high degree of variability generally recorded in nitrate and ammonium levels, forcing the macroalgae in the model to contend with some instability in its nutrient supply. @fig-example-N-input shows an example of the final nitrate and ammonium concentrations driving the growth model in a single cell, located in New South Wales.

```{r fig-example-N-input}
#| fig-width: 6
#| fig-height: 6
#| fig-cap: An example of the daily nitrate (top) and ammonium (bottom) input and data for a single cell in New South Wales. Black lines show the input driving the growth model (constructed curve), blue dots show data from the nearest reference station (Port Hacking) and orange dots show distance-weighted data from the three nearest outfall points (Bateman’s Bay, Bermagui, and Bombo). 

# day15 <- data.frame(month = 1:12, day = rep(15, 12)) %>% 
#   mutate(yday = yday(make_date(year = 2023, month = month, day = day)))
# 
# tar_read(refstation_cell_match, branches = 9500, store = cell_store)
# tar_read(outfall_sites_cellpaired, branches = 9500, store = cell_store)
# 
# Ni_data_prioritised <- tar_read(Ni_data_prioritised, store = cell_store, branches = 9500) %>% 
#   mutate(source = case_when(yday %in% day15$yday ~ "outfall", TRUE ~ "refstation"),
#          form = "1") %>% 
#   merge(cell_coords, by = "cell_ID")
# Am_data_prioritised <- tar_read(Am_data_prioritised, store = cell_store, branches = 9500) %>% 
#   mutate(source = case_when(yday %in% day15$yday ~ "outfall", TRUE ~ "refstation"),
#          form = "2") %>% 
#   merge(cell_coords, by = "cell_ID")
# 
# df <- rbind(Ni_data_prioritised, Am_data_prioritised) %>% 
#   mutate(data_value = set_units(data_value, "mg m$^{-3}$"),
#          data_value = set_units(data_value, "umol L-1"),
#          data_value = drop_units(data_value),
#          value = set_units(value, "mg m$^{-3}$"),
#          value = set_units(value, "umol L-1"),
#          value = drop_units(value))
# unique(df$state)
# unique(df$cell_ID)
# 
# ggplot(df, aes(x = yday, y = data_value, color = as.factor(source))) +
#   geom_point() +
#   facet_wrap(facets = vars(form), nrow = 2) +
#   geom_line(aes(x = yday, y = value), inherit.aes = F) +
#   scale_color_manual(values = c("orange", "steelblue")) +
#   scale_y_continuous(limits = c(0, 3.5), breaks = seq(0, 4, 0.5)) +
#   env_plot +
#   theme(strip.text = element_blank()) +
#   labs(x = "Day of the year", y = nitrogen_lab)

ggplot()
```

Concentrations of both nitrate (@fig-Ni-input) and ammonium (@fig-Am-input) were very low relative to concentrations in coastal waters of other continents, which is expected for Australian coastal waters. However, nitrate around Victoria and Tasmania showed a peak in the middle of the year which was not observed in cells in other states, driven by readings from the BOM reference station located at Maria Island off the east coast of Tasmania. 

```{r fig-Ni-input}
#| fig-width: 8
#| fig-height: 7
#| fig-cap: Summary of nitrate concentration values across the year within each state, averaged across 2019-2023. Bold lines show the mean across all cells within each state while ribbons show the standard deviation between cells.

ggplot()
```

Ammonium concentrations were also low relative to global levels (@fig-Am-input) and generally at least one order of magnitude lower than nitrate but showed greater variability than nitrate. There were no obvious seasonal trends in ammonium, although variability between cells was generally higher than nitrate.

```{r fig-Am-input}
#| fig-width: 8
#| fig-height: 7
#| fig-cap: Summary of ammonium concentration values across the year within each state, averaged across 2019-2023. Bold lines show the mean across all cells within each state while ribbons show the standard deviation between cells.

ggplot()
```

## Model parameterisation {#sec-model-parameterisation}

Growth of the genus *Asparagopsis* is relatively understudied. There are two species (or species complexes) within the genus: *A. armata* is a temperate species, while *A. taxiformis* is more commonly found in tropical areas. Due to the paucity of studies overall, and especially for *A. taxiformis*, the only difference between the species in this analysis is in their temperature response. Additionally, few studies have investigated *Asparagopsis* in its native Australian range – far more work has been done in the northeastern Atlantic and the Mediterranean, where it is an invasive species. Consequently, much of the parameterisation done here was based on studies of invasive *Asparagopsis*, which is almost exclusively *A. armata* [@zanolla_concise_2022].

### Parameter sensitivity

A sensitivity analysis was undertaken by varying each parameter in the model by $\pm$10% and analysing the change in model output according to: 

$$
S = \frac{P[1.1p]-P[0.9p]}{0.2P[p]}
$$

where P is the model output when parameter p is being tested, and $P\propto p^S$ (Murray & Parslow, 1997). If the normalised sensitivity $S=1$ then output is proportional to $p$, and if $S=2$ then the output is proportional to $p^2$. In practice however, the output’s relative sensitivity to each parameter is more important than the absolute value (Everett et al., 2007).

Sensitivity was measured in terms of total nitrogen removed ($TN$, mg m$^{-3}$), defined as 

$$
TN=N_f+N_s
$$

where $N_f$ is nitrogen fixed as macroalgae tissue and $N_s$ is unassimilated nitrogen stored in the tissue. Each species started with the same out-planted biomass (0.005 g L$^{-1}$) and the same relative internal nutrient quotient (0.5) and was grown for 42 days. Since the model does not explicitly include thalli senescence due to age, this growing period was based on previous research and advice from industry.

### Sensitivity conditions

Sensitivity runs started on the first of every month to cover annual variability, for 12 runs per year, with the same mean annual conditions looping from December to January. Sensitivity runs were conducted at culture depths of 0.5 m, 2 m, 3.5 m, and 5 m below the surface with a light attenuation coefficient of 0.65 (clear water). The canopy was always 2.5 m in height, and the total water column depth was always 50 m.
Curves for temperature, surface irradiance, salinity, water velocity, nitrate and ammonium were constructed to capture “typical” conditions based on daily means of each state from 2019-2023. This resulted in two sensitivity curves for temperature (@fig-state-temperature) and a single sensitivity curve for irradiance (@fig-state-irradiance), water velocity (@fig-state-water-velocity), salinity (@fig-state-salinity), nitrate (@fig-state-nitrate) and ammonium (@fig-state-ammonium).

```{r fig-state-temperature}
#| fig-width: 8
#| fig-height: 7
#| fig-cap: Average temperatures in each state region (left, see Figure 2 for detail) and the two temperature curves used for sensitivity analysis (high and low, right).

ggplot()
```

```{r fig-state-irradiance}
#| fig-width: 8
#| fig-height: 7
#| fig-cap: Average irradiance values in each state (left, see Figure 3 for detail) and the irradiance curve used for sensitivity analysis (right).

ggplot()
```

```{r fig-state-water-velocity}
#| fig-width: 8
#| fig-height: 7
#| fig-cap: Average water velocity in each state region (left, see Figure 5 for detail) and the water velocity curve used for sensitivity analysis (right).
#| 
ggplot()
```

```{r fig-state-salinity}
#| fig-width: 8
#| fig-height: 7
#| fig-cap: Average salinity values in each state region (left, see Figure 4 for detail) and the salinity curve used for sensitivity analysis (right).

ggplot()
```

Both nitrate and ammonium are naturally low in Australian coastal waters and this is reflected in the sensitivity analysis (Figure 14). Despite this, the sensitivity curves for nutrients were constructed to encompass a relatively large level of variability. The “typical” ammonium concentration tested was slightly higher than actual reference levels, as it’s suspected that total N removed will be very sensitive to changes in parameters related to ammonium uptake.

```{r fig-state-nitrate}
#| fig-width: 8
#| fig-height: 7
#| fig-cap: Average nitrate concentrations in each state region (left) and the nutrient curves used for sensitivity analysis (right). 

ggplot()
```

```{r fig-state-ammonium}
#| fig-width: 8
#| fig-height: 7
#| fig-cap: Average ammonium concentrations in each state region (left) and the nutrient curves used for sensitivity analysis (right). 

ggplot()
```

## Growth and bioremediation
### Growth limitations
The model is constructed of several functions that individually limit growth based on discreet drivers (light, nutrients, temp., salinity). In the model these functions are combined in a multiplicative relationship. However, each function can be visualised individually to show its impact on growth, in this case we ignore nonlinear interactions between models.. Running these portions of the model across all cells can therefore help determine why a specific area is or is not suitable for seaweed growth and bioremediation. 

Of the model’s components, light and temperature limitation were expected to have the largest impact on species’ growth and bioremediation potential. Temperature limitation depends only on a species’ temperature tolerance and so was run with no biomass specifications. Light limitation depends primarily on PAR availability at culture depth, although biomass-dependent self-shading also has a small impact. Light limitation was run at 0.5 m depth (with a canopy of 5 m) for all species using a very small starting biomass (0.005 g L$^{-1}$) to keep self-shading to a minimum. 

### Total growth and bioremediation
The growth and removal model in its entirety was run for both species (*Asparagopsis armata* and *Asparagopsis taxiformis*) in all cells. The total nitrogen removed was calculated as the difference between total tissue nitrogen (N_f+N_s) at harvest and out-planting. 

## Growth scenarios
The gridded environmental inputs fed into the model will not cover the full range of potential conditions that a seaweed farm might experience. Therefore, we describe “growth scenarios” to encompass some of the variation for comparison (@tbl-scenario-modifications). Temperature and salinity have both been varied by 10% in relevant scenarios. Rather than varying light directly, the light attenuation coefficient (kW) has been varied by 10% to simulate clearer or more cloudy water. Water velocity has been varied by a factor of 4 in each direction to simulate stagnant and exposed waters. Nitrate and ammonium have been modified using fixed additions to their values to match more closely with measured conditions. 

These scenarios were run for each state’s typical conditions and the difference in total N removed and proportional N compared to the “base” scenario, which used state average environmental conditions and a water depth of 50 m. 

|------|------|------|------|------|------|------|------|
| Scenario name | DEPTH (M) | TEMPERATURE | kW           | SALINITY         | WATER VELOCITY  | NITRATE   | AMMONIUM  |
FISH FARM       | 30        |             |              |                  |                 | +3.5 µM   | +6.5 µM   |
DEEP WATER      | 100       | $\times$0.9 | $\times$1.1  |                  | $\times$4       |           |           |
SHALLOW BAY     | 15        | $\times$1.1 | $\times$0.9  | $\times$1.1      | $\times$0.25	  | +3.5 µM   |           |
ESTUARY MOUTH   | 30        |             | $\times$0.9  | $\times$0.9      |                 | +3.5 µM	  |           |
: Summary of modifications made to state average values for each growth scenario. Blank cells indicate that no changes were made to state average values. {#tbl-scenario-modifications}

## Statistical analysis & code availability
All model running, data handling and analysis were conducted in R version 4.4.2 “Pile of Leaves” (R Core Team, 2024). The custom package “macrogrow” (available on Github) was used to run the seaweed growth modelling portions of this analysis. All other packages used are listed in the Github repository.

# Results 
## Species parameterisation 

All species parameters are summarised in Table 2. Parameters for *Asparagopsis armata* and *A. taxiformis* were sourced from studies performed on the individual species where possible, but this was complicated by the limited data available. While *Asparagopsis* is native to Australia, it has been more extensively studied in its invasive range in the northeastern Atlantic and the Mediterranean. Additionally, most *Asparagopsis* studies focus on the tetrasporophyte lifestage (which is easier to cultivate in the lab) but the grow-out phase in oceanic seaweed culture is the gametophyte. Parameters sourced from tetrasporophyte studies but used here for gametophyte growth include the ammonium uptake parameters V_am and K_am, the nitrate uptake parameters M_ni and C_ni, the light response parameter I_o and the temperature response parameters T_opt, T_min,and  T_max. There was no information available to parameterise the salinity tolerance of *Asparagopsis*. This study instead used typical Australian conditions to define “optimum” salinity and gave the species a tolerance range 50% narrower than the range of Ecklonia.

Table 2: Parameters defined for use in the growth model for *Asparagopsis armata*, *Asparagopsis taxiformis*. Note that *A. taxiformis* shares all parameters with *A. armata* except those relating to temperature. Parameters not relevant to a species are left blank.

Modelled species temperature responses (Figure 16) are likely to be very impactful on overall species growth. *Asparagopsis armata* is better-suited to cold water, while *Asparagopsis taxiformis* has a narrower temperature tolerance range than the other species but is better-suited to higher temperatures.

Figure 16: Temperature response of *Asparagopsis armata* (red), *Asparagopsis taxiformis* (pink).

Modelled light responses (Figure 17) shows variation between species, but also in the same species grown at different depths. *Asparagopsis* does very well in shallow waters and low-light situations.

Figure 17: Simulated light response of *Asparagopsis* (red) when grown at depths of 0.5 m (solid), 2 m (dashed), 3.5 m (dot-dashed) or 5 m (dotted).

The modelled response of the different species to changes in salinity (Figure 18) showed similar trends to their temperature responses. *Asparagopsis* had similar optimum salinities.

Figure 18: Salinity response of *Asparagopsis* (red).

Modelled responses to nitrate and ammonium were very different between species (Figure 19). Like most seaweeds, the species studied here prefer ammonia to nitrate. *Asparagopsis* only takes up nitrate at a low, constant rate (Table 2). This is potentially very important, as nitrate is generally the dominant form of nitrogen in Australian coastal waters.

Figure 19: Uptake rate of nitrate (solid) and ammonium (dashed) by *Asparagopsis* (red).

### Parameter sensitivity

Due to physiological differences, the same out-planted biomass (0.005 g L$^{-1}$) and internal relative nutrient quotient (0.5) for each species in the sensitivity runs equated to different starting N_f values: 97.16 mg m−3 for both *Asparagopsis* species.

Both *A. armata* and *A. taxiformis* were most sensitive to changes in parameters controlling their internal nutrient quotient (Q_min and Q_max). This is as expected, as narrowing the possible bounds of the macroalgae internal nutrient reserves decreased the total nitrogen removed, and *Asparagopsis* additionally has a higher minimum internal nutrient quotient requirement than the other species. For more detailed parameter sensitivities, see Appendix section: Species parameter sensitivity.

## Growth and bioremediation
### Temperature limitation

Relative temperature limitation on each species’ growth (T_lim) varied both spatially and temporally. - summarise the temporal trends across a typical year in each state, and - show the cells across Australia that maintained “good” growing temperatures by indicating how many consecutive days in a typical year where T_lim$\geq$0.75. While this value is arbitrary, it provides a consistent reference point for comparing spatial and temporal patterns in growth conditions across regions. Unsurprisingly, the Northern Territory, Queensland and northern Western Australia were almost entirely too warm for the temperate *A. armata*. Tasmania had some of the strongest seasonal variation in temperature limitation, with most of the state providing good conditions from late December to late April (Figure 21). While cells on the east coast of Tasmania were slightly better suited to *A. armata* growth, the state showed little spatial variation overall. New South Wales also showed strong seasonal variation in temperature limitation, showing good growing temperatures for *A. armata* from mid-July to late October, and mostly in the lower half of the state (below ~32˚S).

By contrast, South Australia and southern Western Australia showed weak seasonal variation in temperature limitation but much higher spatial variation. The only area that appeared to have ideal temperature conditions for *A. armata* growth year-round was the area between Esperance and Albany in southern Western Australia (Figure 20). 

Figure 20: The maximum number of days where Tlim $\geq$ 0.75 for *A. armata* in a typical year.

Figure 21: Limitation of relative growth by temperature for *A. armata* across all cells in a typical year (mean $\pm$ SD). 

Within the model, the only difference between the two *Asparagopsis* species was their temperature response. Despite its higher temperature tolerance (see Figure 16) *A. taxiformis* seemed overall more limited by temperature than *A. armata* across Australia (Figure 23). The tropical species was mostly unable to grow below 33˚S on the west coast and 35˚S, and there was only a very small area off the central coast of Western Australia just outside of Shark Bay where temperature conditions were ideal for *A. taxiformis* year-round. 

Figure 22: The maximum number of days where Tlim $\geq$ 0.75 for *A. taxiformis* in a typical year.

Figure 23: Limitation of relative growth by temperature for *A. taxiformis* across all cells in a typical year (mean $\pm$ SD).

### Light limitation

None of the species studied here were sensitive to direct changes in their light-related parameters (Figures 51-54). Additionally, given the similar light responses of the different species (Figure 17) and the available surface irradiance mostly falling between 150-600 µmol photons m-2 s$^{-1}$ (Figure 3), all species were only grown at a depth of 0.5 m. The light attenuation coefficient kW was fixed at 0.6 (clear water) so the following represent light availability at an “optimum” depth.

*Asparagopsis* (both *A. armata* and *A. taxiformis*) were best suited to high-light conditions, showing less limitation due to light availability at higher latitudes (Figure 28). There were very few days where I_lim dropped below 0.75 in the northern states, however in Tasmania and Victoria limitation due to light availability was consistently <0.75 from mid-April to early October and <0.5.from early May to early August (Figure 29).

Figure 28: The maximum number of days where Ilim $\geq$ 0.75 for *Asparagopsis* spp. in a typical year.

Figure 29: Limitation of relative growth by changes in seasonal light availability for *Asparagopsis* spp. across all cells grown at 0.5 m depth in a typical year (mean $\pm$ SD).

### Total growth and bioremediation

After running the model in all cells, there were 1034 cells (of 11677) where the run failed due to missing environmental data or the cell being too shallow to accommodate the 5.5 m canopy. Overall, nitrogen removal by *A. armata* and *A. taxiformis* was less than other species. 

Nitrogen removal by *A. armata* was fairly consistent across a typical year in Tasmania with a mean of 204.0 g m-2 of farmed area, Victoria with a mean of 256.0 g m-2 of farmed area, and South Australia with a mean of 327.0 g m-2 of farmed area (Figure 34). The highest total removal across the year was also in South Australia. Nitrogen removal was most temporally variable in New South Wales, ranging from 8.5 g m-2 of farmed area to 540.0 g m-2 of farmed area, peaking when the simulated macroalgae was out-planted in August.

Figure 34: Total nitrogen removed (mean $\pm$ SD) by *A. armata* planted in different months across all cell in each state.
 
Nitrogen removal in South Australia was relatively spatially consistent, while the east coast of Australia showed greater variability (Figure 35). Some areas of Victoria on the borders of New South Wales and South Australia showed nitrogen removal comparable to the neighbouring state, likely reflecting the influence of data from the different reference stations. The potential “hotspot” area between Esperance and Albany in southern Western Australia, identified in temperature limitation maps above (see Figure 20), did not show higher nitrogen removal rates than the rest of the state. Similarly, Tasmania showed moderate nitrogen removal rates year-round (), despite the species being relatively limited by temperature from May to November (see Figure 21).

Figure 35: Southern half of Australia showing the total nitrogen removed by *A. armata* planted in March (top) or August (bottom).

Nitrogen removal by *A. taxiformis* was highest in New South Wales, with a mean nitrogen removal rate of 304 g m-2 of farmed area in each month of outplanting (Figure 36). Nitrogen removal was less consistent across the year in South Australia (peaking at a mean of 158 g m-2 of farmed area in February), northern Western Australia (peaking at 143 g m-2 of farm in June) and the Northern Territory (peaking at 142 g m-2 of farmed area in June).

Figure 36: Total nitrogen removed (mean $\pm$ SD) by *A. taxiformis* planted in different months across all cell in each state. 

Total nitrogen removal by *A. taxiformis* was relatively spatially consistent in most states except for Queensland, where the southern coast showed higher removal rates (Figure 37). This was likely driven by ambient nitrogen data, particularly the differences between reference stations (Figures 8 and 9). By contrast, the spatial variability seen in the north-west of the country (Figure 37) is more likely driven by temperature and light variability, given that there are fewer reference stations and outfall points (Figure 6).

Figure 37: Total nitrogen removed by *A. taxiformis* planted in May in the Northern Territory and northern Western Australia (top), and in October in New South Wales and Queensland (bottom).

## Growth scenarios

When compared with the base scenario, all species were able to remove the most nitrogen in the fish farm scenario, almost certainly because of the increase in ammonium concentration from (very low) background levels (Figure 42). *Asparagopsis armata* and *A. taxiformis* consistently removed more nitrogen in the fish farm scenario.

Differences in nitrogen removed by *Asparagopsis* were minimal in the estuary mouth scenario across all states (ranging from -2.1% to +2.5% of base). Similarly, in the shallow bay scenario, *Asparagopsis* was able to remove from 0.4 to 9.5$\times$ more nitrogen. The deep water scenario showed mostly negative but relatively small changes from the base removal, with changes ranging from -36% removal by *A. taxiformis* in New South Wales to +16.2% removed by *A. armata* in northern Western Australia. 

Figure 42: Nitrogen removal for different scenarios compared to the base scenario. Colours show the different scenarios as fish farm (purple), deep water (green), estuary mouth (orange) and shallow bay (pink).

# Discussion {#sec-discussion}

This study demonstrates that seaweed aquaculture has significant bioremediation potential across Australia, with scenario modelling suggesting that *Asparagopsis* species can be effective at removing excess nitrogen from the environment. Across a range of conditions, these species showed nitrogen uptake potential that varied by region and season. Our model tested different pollution contexts, including marine aquaculture waste and estuarine runoff, highlighting the ability of seaweed cultivation to mitigate nutrient pollution.

Our analysis highlighted variability in bioremediation potential among different macroalgae species under varying environmental conditions across Australia. The local availability of nitrogen, specifically ammonium and nitrate, emerged as the primary driver of this variability. Nitrate, generally in very low concentrations, is the predominant form of nitrogen in Australian coastal waters, which can be considered as oligotrophic (e.g. Chen et al., 2020; Pritchard et al., 2001; Twomey et al., 2007; Wild-Allen et al., 2010). This poses a clear challenge for the use of macroalgae for bioremediation in Australia. While all species studied showed higher uptake efficiency of ammonium, the low ambient levels of this compound limit their overall nitrogen uptake. Anthropogenic nutrient inputs, such as those from fish farms and sewage treatment outfalls, can create high nitrogen environments conducive to effective macroalgae cultivation and therefore bioremediation. The highest mean concentration of nitrate reported here from available data was 3.9 µM, and the fish farm scenario almost doubled this value (7.4 µM). While this is expected within Australian waters, it is still very low in a global context. For example, in Chile, mean coastal nitrate concentrations near salmon farms range from 3.2 µM in summer to 22.1 µM in winter (Varela et al., 2018). 

All species studied here demonstrated a preference for ammonium uptake over nitrate, with *Asparagopsis armata* and *A. taxiformis* showing the highest uptake efficiency of ammonium. However, in the context of bioremediation, a species’ ability to take advantage of surges in nutrient concentrations is perhaps more important than its absolute uptake rate. By contrast, *Asparagopsis*, has a constant nitrate uptake rate which was minimally impacted by changes in nitrate concentrations. It is worth nothing that minimal seasonal variation in nutrient concentration was included in the present study. However, seasonal trends in ammonium and nitrate concentrations are often observed in coastal waters due to natural process (e.g. phytoplankton growth) and anthropogenic inputs (e.g. urban runoff from rainfall events).

Temperature played a secondary role in influencing nitrogen uptake efficiency, generally reflecting the natural geographic ranges of the species. *Asparagopsis armata* showed the potential for year-round growth (indicating year-round nitrogen uptake) in cooler regions of Australia, reflecting its temperate to subtropical range and wide temperature tolerance. *Asparagopsis taxiformis* showed more restricted temperature ranges, limiting their bioremediation potential to specific regions and seasons. 

Light-related parameters did not appear to significantly influence the nitrogen uptake of any macroalgae species, possibly driven by the low resolution in the available light attenuation data. Given the similar light responses among species and the consistent surface irradiance among all regions, all species were only modelled as growing at a depth of 0.5 m and in uniformly clear water. This uniformity in light response and growth depth simplified the model, ensuring that light availability did not confound the results. There is, however, a clear need for accurate site-specific light attenuation data to assess the impact of depth on macroalgae growth, as light is crucial for photosynthesis and growth. Smart et al (2022) used a laboratory setting to examine the nitrogen uptake kinetics of four kelp species under different ammonium concentrations. The nitrogen uptake rate of all species varied seasonally, increasing with higher temperatures and light availability.

*Asparagopsis* species are currently the most commercially valuable for cultivation than the other species studied here due to their use in reducing methane emissions from livestock. However, modelling revealed that, despite the high nitrogen tissue content of *Asparagopsis* spp., these species are significantly less efficient in nitrogen uptake. This inefficiency is likely due to their lower rates of nitrate uptake compared to the other species combined with Australia’s low nitrate waters, their high nitrogen tissue requirements, and their more narrow temperature tolerance. These assumptions, however, are largely based on research conducted in regions outside Australia (e.g. Torres et al., 2021; Zanolla et al., 2019). More Australian local-scale studies are necessary to better understand the nitrogen uptake kinetics of *Asparagopsis* spp. in Australian coastal waters. 

The study’s methodology, which integrated reference station and outfall data, provided a broad overview of spatial and temporal variation in nutrient concentrations in Australian coastal waters, However, the actual impact of outfalls and other nutrient sources on nitrogen concentration is more complex and site-specific, often stratified by depth and influenced by local hydrodynamic conditions (Pritchard et al., 2001). This suggests that detailed knowledge of ammonium and nitrate concentrations at specific sites, along with species-specific uptake kinetics, is crucial for accurately assessing the bioremediation potential of macroalgae in different coastal environments. Furthermore, individual site conditions could provide valuable insights for growers on the suitability of particular species for cultivation. For example, while the present model captured some thalli degradation due to high water velocities, it did not include breakage off the lines due to high swells or storm events, which would have a greater impact on more fragile seaweeds like *Asparagopsis* spp.

Given *Asparagopsis* is the only genus studied here that currently has significant commercial value, future research should focus on identifying ‘hotspots’, considering temperature thresholds, time of year and placement, where *Asparagopsis* can thrive. Additionally, understanding the nitrogen uptake rate of the *Asparagopsis* gametophyte life stage, rather than assuming it is identical to the tetrasporophyte life stage, is essential for advancing open-water *Asparagopsis* culture for bioremediation. 

Overall, this study provides valuable insights into the bioremediation potential of native Australian macroalgae species, highlighting the importance of nitrogen availability and temperature in shaping their effectiveness. The findings highlight the need for species selection and consideration of site-specific conditions when integrating seaweed aquaculture for bioremediation in Australian waters. As the industry explores seaweed farming's role in coastal management, future research should focus on refining spatial models, incorporating finer-scale environmental data, and evaluating economic feasibility. With the right approach, seaweed farming could reduce water pollution, support sustainable aquaculture, and enhance coastal ecosystem health. This study lays the groundwork for guiding site selection and key considerations in expanding seaweed farming across Australia.

# Acknowledgments

The project authors would like to thank Jens Knauer and Margie Rule of the \acr{ASSA} for their input and expertise on *Asparagopsis** husbandry, and Professor Rocky de Nys for his feedback on study methodology. Data for this study were sourced from Australia’s \acr{IMOS}, which is enabled by the \acr{NCRIS}. Data were also sourced from the \acr{NOD}. Work for \acr{NOD} is undertaken for the Marine and Coastal Hub, a collaborative partnership supported through funding from the Australian Government's \acr{NESP}.

# References