---
title: Investigating the bioremediation potential of seaweed aquaculture across Australia
date: last-modified
bibliography: ["refs.bib", "manual-refs.bib"]

format:
  html:
    theme: simplex
    css: ["extra.css"]
    code-fold: true
    code-overflow: wrap
    toc: true
    toc-expand: true
    toc-location: left
    toc-depth: 4
    lang: en-GB
    grid:
      sidebar-width: 300px
      body-width: 1250px
      margin-width: 200px
      gutter-width: 2em
        
execute:
  eval: false
  cache: true

knitr: 
  opts_chunk:
    fig.align: center
    fig.height: 9.5
    fig.width: 6.5
---

```{r setup}
#| eval: true
#| include: false
#| cache: false

# You need quarto to generate this markdown - install here: https://quarto.org/docs/get-started/
# This markdown uses the acronyms extension - install by entering quarto add rchaput/acronyms@master into the terminal
# This markdown uses TinyTex for PDF rendering - install with tinytex::install_tinytex()

library(stringr)
library(units)
library(targets)
library(tarchetypes)
library(macrogrow)
library(ozmaps)
library(arrow)
library(gganimate)
library(magrittr)
library(ggplot2)
library(dplyr)
library(terra)
library(tidyterra)
library(here)
library(lubridate)
library(tidyr)
library(cowplot)
library(purrr)
library(conflicted)
conflicts_prefer(dplyr::select, dplyr::filter)
```

```{r global variables and functions}
#| eval: true
#| include: false
#| cache: false

# source(file.path(base_path, "R", "other-functions.R"))
spec_store <- here() %>% file.path("targets_outputs", "_species")
envi_store <- here() %>% file.path("targets_outputs", "_env_inputs")
runs_store <- here() %>% file.path("targets_outputs", "_model_running")
spec_data <- here() %>% file.path("data", "processed_species")
envi_data <- here() %>% file.path("data", "processed_env_inputs")
runs_data <- here() %>% file.path("data", "processed_model_running")

bounds <- file.path(envi_data, "states_bbox.qs") %>% qs::qread()

prettyplot <- theme_classic() +
  theme(text = element_text(family = "serif", size = 11, colour = "black"),
        legend.position = "none",
        axis.title.y = element_text(vjust = 1.5),
        axis.title.x = element_text(vjust = 1.5),
        legend.title = element_blank())
env_plot <- prettyplot + theme(aspect.ratio = 0.55)

find_read <- function(path, pattern) {
  fnm <- list.files(path, full.names = T) %>% str_subset(pattern)
  if (all(str_detect(fnm, ".qs"))) {dat <- purrr::map(fnm, qs::qread)}
  if (all(str_detect(fnm, ".parquet"))) {dat <- purrr::map_dfr(fnm, arrow::read_parquet)}
  return(dat)
}

nitrogen_lab <- expression("Concentration ("*mu*"M N)")
nitrate_lab <- expression("Nitrate concentration ("*mu*"M NO"[3]*")")
ammonium_lab <- expression("Ammonium concentration ("*mu*"M NH"[4]^+")")

# Allows using "units" package to convert between moles and g of nitrogen ONLY
remove_unit("mol")
install_unit("mol", "14.0067 g")

prettynum <- function(num, sigfig=2) {format(round(as.numeric(num), sigfig), nsmall=1, big.mark=",")}

species_pal <- c("red3", "dodgerblue3")
depths_pal <- c("green3", "#FF8247", "dodgerblue3", "grey56")
scens_pal_1 <- c("#292928", "dodgerblue3", "#66a61e", "#e7298a", "#e6ab02")
scens_pal_2 <- c("dodgerblue3", "#66a61e", "#e7298a", "#e6ab02")
states_pal <- c(
  "NTE" = "#d95f02", "Northern Territory" = "#d95f02",
  "QLD" = "#e6ab02", "Queensland" = "#e6ab02",
  "NSW" = "#7570b3", "New South Wales" = "#7570b3",
  "SAU" = "#1b9e77", "South Australia" = "#1b9e77",
  "TAS" = "#66a61e", "Tasmania" = "#66a61e",
  "VIC" = "#e7298a", "Victoria" = "#e7298a",
  "WAN" = "#a6761d", "Western Australia (N)" = "#a6761d",
  "WAS" = "#666666", "Western Australia (S)" = "#666666"
)
lims_pal <- c("No_limit" = "black", "T_lim" = "#e7298a", "I_lim" = "#e6ab02", "S_lim" = "#7570b3", 
              "Q_lim" = "#66a61e", "Nf_loss" = "#1b9e77", "Ns_loss" = "#e7298a")
states <- tar_read(states, store = runs_store)
states_ord <- c("NTE", "QLD", "NSW", "SAU", "TAS", "VIC", "WAN", "WAS")
states_lng <- c("Northern Territory", "Queensland", "New South Wales", "South Australia", "Tasmania", "Victoria", "Western Australia (N)", "Western Australia (S)")
```

# Methods {#sec-methods}

```{r cell-coords}
#| eval: true
#| code-summary: |
#|   Load in all the BARRA-C2 cells being used, with their coordinates.

cell_coords <- find_read(envi_data, "cell_coords") %>% 
  mutate(state = factor(state, levels = states_ord)) %>% 
  dplyr::select(-layer)
cells_included <- find_read(runs_data, "include")
cells_omitted <- find_read(runs_data, "omit")

totals <- cell_coords %>% 
  filter(!cell_no %in% cells_omitted) %>% 
  group_by(state) %>% 
  reframe(cells = n())
```

The waters around Australia were divided into `r as.integer(sum(totals$cells))` cells for this analysis: 
`r as.integer(totals$cells[totals$state == "VIC"])` in Victoria, 
`r as.integer(totals$cells[totals$state == "NSW"])` in New South Wales, 
`r as.integer(totals$cells[totals$state == "QLD"])` in Queensland, 
`r as.integer(totals$cells[totals$state == "SAU"])` in South Australia, 
`r as.integer(totals$cells[totals$state == "WAN"]+totals$cells[totals$state == "WAS"])` in Western Australia (divided into 
`r as.integer(totals$cells[totals$state == "WAN"])` cells in the north and 
`r as.integer(totals$cells[totals$state == "WAS"])` in the south), 
`r as.integer(totals$cells[totals$state == "TAS"])` in Tasmania, and 
`r as.integer(totals$cells[totals$state == "NTE"])` in the Northern Territory. 

```{r fig-states-map}
#| eval: true
#| fig-cap: |
#|   Map showing the area included this analysis, with colours indicating the different states. The ~60 km coastline buffer includes 95,418 cells approximately 4 x 4 km each.

ggplot(cell_coords, aes(x = longitude, y = latitude, fill = state)) +
  geom_raster() +
  geom_sf(data = ozmap_data(data = "states"), inherit.aes = F) +
  coord_sf(xlim = c(bounds[["AUS"]]["lonmin"], bounds[["AUS"]]["lonmax"]),
           ylim = c(bounds[["AUS"]]["latmin"], bounds[["AUS"]]["latmax"])) +
  scale_x_continuous(breaks = seq(100, 160, 5)) +
  scale_y_continuous(breaks = seq(-5, -45, -5)) +
  scale_colour_manual(values = states_pal) +
  scale_fill_manual(values = states_pal) +
  labs(x = "Longitude", y = "Latitude") +
  prettyplot +
  theme(legend.position = "none")
```

## Environmental inputs {#sec-envinputs}

```{r all-cell-input-timeseries}
#| eval: true
#| cache: false

cell_input_all <- find_read(envi_data, "cell_input_all") %>% 
  mutate(state = factor(state, levels = states_ord, labels = states_lng))

T_input <- cell_input_all %>% 
  select(cell_no, state, yday, T_input)

(T_means <- T_input %>% 
  group_by(state) %>% 
  reframe(mean = mean(T_input, na.rm = T),
          max = max(T_input, na.rm = T),
          min = min(T_input, na.rm = T)) %>% 
  mutate(variable = "T_input"))

I_input <- cell_input_all %>% 
  select(cell_no, state, yday, I_input)

(I_means <- I_input %>% 
  group_by(state) %>% 
  reframe(mean = mean(I_input, na.rm = T),
          max = max(I_input, na.rm = T),
          min = min(I_input, na.rm = T)) %>% 
  mutate(variable = "I_input"))

K_input <- cell_input_all %>% 
  select(cell_no, state, yday, Kd_490)

(K_means <- K_input %>% 
  group_by(state) %>% 
  reframe(mean = mean(Kd_490, na.rm = T),
          max = max(Kd_490, na.rm = T),
          min = min(Kd_490, na.rm = T)) %>% 
  mutate(variable = "Kd_490"))

S_input <- cell_input_all %>% 
  select(cell_no, state, yday, S_input)

(S_means <- S_input %>% 
  group_by(state) %>% 
  reframe(mean = mean(S_input, na.rm = T),
          max = max(S_input, na.rm = T),
          min = min(S_input, na.rm = T)) %>% 
  mutate(variable = "S_input"))

UV_input <- cell_input_all %>% 
  select(cell_no, state, yday, UV_input)

(UV_means <- UV_input %>% 
  group_by(state) %>% 
  reframe(mean = mean(UV_input, na.rm = T),
          max = max(UV_input, na.rm = T),
          min = min(UV_input, na.rm = T)) %>% 
  mutate(variable = "UV_input"))
```

Mean daily temperature ranged from 
`r round(min(T_means$mean), 1)` in `r as.character(T_means$state[T_means$mean == min(T_means$mean)])` to 
`r round(max(T_means$mean), 1)` in `r as.character(T_means$state[T_means$mean == max(T_means$mean)])`,
with a mean across states of `r round(mean(T_means$mean), 1)`. 

```{r fig-T-input}
#| eval: true
#| fig-cap: |
#|   Summary of temperature values across the year within each state, averaged across 2019-2023. Bold lines show the mean across all cells within each state while ribbons show the standard deviation between cells.

T_input_state <- T_input %>%
  group_by(state, yday) %>% 
  reframe(value = mean(T_input, na.rm = T),
          sd = sd(T_input, na.rm = T))

(
  p_input_state_av <- T_input_state %>% 
    ggplot(aes(x = yday, y = value, ymin = value-sd, ymax = value+sd, colour = state, fill = state)) +
    geom_line(linewidth = 0.75) +
    geom_ribbon(alpha = 0.35, linewidth = 0.25) +
    facet_wrap(facets = vars(state), ncol = 2) +
    scale_y_continuous(breaks = seq(0, 40, 5), limits = c(10, 35)) +
    scale_colour_manual(values = states_pal) +
    scale_fill_manual(values = states_pal) +
    env_plot +
    labs(x = "Day of the year", y = expression("Temperature ("*degree*"C)"))
)
```

Mean daily irradiance ranged from 
`r round(min(I_means$mean), 0)` in `r as.character(I_means$state[I_means$mean == min(I_means$mean)])` to 
`r round(max(I_means$mean), 0)` in `r as.character(I_means$state[I_means$mean == max(I_means$mean)])`,
with a mean across states of `r round(mean(I_means$mean), 0)`. 

```{r fig-I-input}
#| eval: true
#| fig-cap: |
#|   Summary of irradiance values across the year within each state, averaged across 2019-2023. Bold lines show the mean across all cells within each state while ribbons show the standard deviation between cells. 

I_input_state <- I_input %>%
  group_by(state, yday) %>% 
  reframe(value = mean(I_input, na.rm = T),
          sd = sd(I_input, na.rm = T))

p_input_state_av %+% I_input_state +
  scale_y_continuous(breaks = seq(0, 1600, 150), limits = c(0, 750)) +
  labs(x = "Day of the year", y = expression("Irradiance (photons m"^-2*" s"^-1*")")) 
```

Mean daily values for the light attenuation coefficient in water ranged from 
`r round(100*min(K_means$mean), 2)` $\times 10^2$ in 
`r as.character(K_means$state[K_means$mean == min(K_means$mean)])` to 
`r round(10*max(K_means$mean), 2)` $\times 10$ in 
`r as.character(K_means$state[K_means$mean == max(K_means$mean)])`,
with a mean across states of `r round(100*mean(K_means$mean), 2)` $\times 10^2$. 

```{r fig-kW-input}
#| eval: true
#| fig-cap: |
#|   Summary of irradiance values across the year within each state, averaged across 2019-2023. Bold lines show the mean across all cells within each state while ribbons show the standard deviation between cells. 

K_input_state <- K_input %>%
  group_by(state, yday) %>% 
  reframe(value = mean(Kd_490, na.rm = T),
          sd = sd(Kd_490, na.rm = T)) %>% 
  mutate(sd = case_when(sd > value ~ value, T ~ sd))

p_input_state_av %+% K_input_state +
  scale_y_continuous(breaks = seq(0, 0.6, 0.1), limits = c(0, 0.6)) +
  labs(x = "Day of the year", y = expression("Light attenuation coefficient (m"^-1*" )"))
```

Mean daily salinity ranged from 
`r round(min(S_means$mean), 1)` in `r as.character(S_means$state[S_means$mean == min(S_means$mean)])` to 
`r round(max(S_means$mean), 1)` in `r as.character(S_means$state[S_means$mean == max(S_means$mean)])`,
with a mean across states of `r round(mean(S_means$mean), 1)`. 

```{r fig-S-input}
#| eval: true
#| fig-cap: |
#|   Summary of salinity values across the year within each state, averaged across 2019-2023 and all depths < 25 m. Bold lines show the mean across all cells within each state while ribbons show the standard deviation between cells.

S_input_state <- S_input %>%
  group_by(state, yday) %>% 
  reframe(value = mean(S_input, na.rm = T),
          sd = sd(S_input, na.rm = T))

p_input_state_av %+% S_input_state +
  scale_y_continuous(breaks = seq(30, 40, 0.5), limits = c(34, 36.5)) +
  labs(x = "Day of the year", y = expression("Salinity (g L"^-1*")"))
```

Mean daily water velocity ranged from 
`r round(min(UV_means$mean), 2)` in `r as.character(UV_means$state[UV_means$mean == min(UV_means$mean)])` to 
`r round(max(UV_means$mean), 2)` in `r as.character(UV_means$state[UV_means$mean == max(UV_means$mean)])`,
with a mean across states of `r round(mean(UV_means$mean), 2)`. 

```{r fig-UV-input}
#| eval: true
#| fig-cap: |
#|   Summary of water velocity values across the year within each state, averaged across 2019-2023 and all depths < 25 m. Bold lines show the mean across all cells within each state while ribbons show the standard deviation between cells.

UV_input_state <- UV_input %>%
  group_by(state, yday) %>% 
  reframe(value = mean(UV_input, na.rm = T),
          sd = sd(UV_input, na.rm = T)) %>% 
  mutate(sd_min = case_when(sd > value ~ value, T ~ sd))

(
  p_input_state_av <- UV_input_state %>% 
    ggplot(aes(x = yday, y = value, ymin = value-sd_min, ymax = value+sd, colour = state, fill = state)) +
    geom_line(linewidth = 0.75) +
    geom_ribbon(alpha = 0.35, linewidth = 0.25) +
    facet_wrap(facets = vars(state), ncol = 2) +
    scale_y_continuous(breaks = seq(0, 1, 0.2), limits = c(-0.01, 1)) +
    scale_colour_manual(values = states_pal) +
    scale_fill_manual(values = states_pal) +
    env_plot +
    labs(x = "Day of the year", y = expression("Water velocity (m s"^-1*")"))
)
```

## Nitrogen {#sec-nitrogen}

```{r N-data sources locations}
#| eval: true

refstation_locations <- here() %>% 
  file.path("data", "nitrogen", "refstation_locations.parquet") %>% 
  arrow::read_parquet() %>% 
  rename(name = StationName) %>% 
  dplyr::select(-code)

outfall_locations <- here() %>% 
  file.path("data_raw", "national-outfall-database", "data-output", "outflow_site_locations.parquet") %>% 
  arrow::read_parquet() %>% 
  rename(lon = longitude,
         lat = latitude) %>% 
  dplyr::select(-state)

N_locations <- rbind(
  outfall_locations %>% mutate(type = "outfall"),
  refstation_locations %>% mutate(type = "refstation")
) %>% 
  mutate(name = as.character(name),
         state = case_when(
    lon > bounds[["SAU"]]["lonmin"] & lon <= bounds[["SAU"]]["lonmax"] & 
      lat > bounds[["SAU"]]["latmin"] & lat <= bounds[["SAU"]]["latmax"]  ~ "SAU",
    lon > bounds[["QLD"]]["lonmin"] & lon <= bounds[["QLD"]]["lonmax"] & 
      lat > bounds[["QLD"]]["latmin"] & lat <= bounds[["QLD"]]["latmax"]  ~ "QLD",
    lon > bounds[["WAS"]]["lonmin"] & lon <= bounds[["WAS"]]["lonmax"] & 
      lat > bounds[["WAS"]]["latmin"] & lat <= bounds[["WAS"]]["latmax"]  ~ "WAS",
    lon > bounds[["WAN"]]["lonmin"] & lon <= bounds[["WAN"]]["lonmax"] & 
      lat > bounds[["WAN"]]["latmin"] & lat <= bounds[["WAN"]]["latmax"]  ~ "WAN",
    lon > bounds[["VIC"]]["lonmin"] & lon <= bounds[["VIC"]]["lonmax"] & 
      lat > bounds[["VIC"]]["latmin"] & lat <= bounds[["VIC"]]["latmax"]  ~ "VIC",
    lon > bounds[["NSW"]]["lonmin"] & lon <= bounds[["NSW"]]["lonmax"] & 
      lat > bounds[["NSW"]]["latmin"] & lat <= bounds[["NSW"]]["latmax"]  ~ "NSW",
    lon > bounds[["NTE"]]["lonmin"] & lon <= bounds[["NTE"]]["lonmax"] & 
      lat > bounds[["NTE"]]["latmin"] & lat <= bounds[["NTE"]]["latmax"]  ~ "NTE",
    lon > bounds[["TAS"]]["lonmin"] & lon <= bounds[["TAS"]]["lonmax"] & 
      lat > bounds[["TAS"]]["latmin"] & lat <= bounds[["TAS"]]["latmax"]  ~ "TAS",
    TRUE ~ NA
  ),
  state = factor(state, levels = states_ord)) %>% 
  relocate(state, .before = name) %>% 
  rename(latitude = lat,
         longitude = lon)

refstation_locations <- filter(N_locations, type == "refstation")
outfall_locations <- filter(N_locations, type == "outfall")
```

```{r fig-N-data-locations}
#| eval: false
#| fig-cap: Map showing the locations of recorded outfalls (points, coloured by state) and the IMOS reference stations (stars) used in this analysis.

ggplot(N_locations, aes(x = longitude, y = latitude, color = state, shape = type)) +
  geom_sf(data = ozmap_data(data = "states"), inherit.aes = F) +
  coord_sf(xlim = c(bounds[["AUS"]]["lonmin"], bounds[["AUS"]]["lonmax"]),
           ylim = c(bounds[["AUS"]]["latmin"], bounds[["AUS"]]["latmax"])) +
  geom_point(size = 3.5) +
  scale_shape_manual(values = c(20, 13)) +
  # scale_colour_manual(values = states_pal, labels = states_lng, na.value = "black") +
  scale_color_brewer(palette = "Dark2") +
  scale_x_continuous(breaks = seq(100, 160, 5)) +
  scale_y_continuous(breaks = seq(-5, -45, -5)) +
  labs(x = "Longitude", y = "Latitude") +
  prettyplot +
  theme(legend.position = "none")
```

```{r cell-matching}
#| eval: true
#| cache: true

all_cells <- sum(totals$cells)

cells_by_station_refstations <- file.path(envi_data, "cells_by_station_refstations.parquet") %>% read_parquet() %>% mutate(ncp = 100*nc/all_cells)
# stations_by_cell_refstations <- file.path(envi_data, "stations_by_cell_refstations.parquet") %>% read_parquet() %>% mutate(nsp = 100*ns/all_cells)
cells_by_station_outfalls <- file.path(envi_data, "cells_by_station_outfalls.parquet") %>% read_parquet() %>% mutate(ncp = 100*nc/all_cells)
stations_by_cell_outfalls <- file.path(envi_data, "stations_by_cell_outfalls.parquet") %>% read_parquet() %>% 
  group_by(ns) %>% 
  reframe(nc = n()) %>% 
  mutate(ncp = 100*nc/all_cells)

# levels(cells_by_station_refstations$StationName)
```

Matching model cells to their two nearest reference stations resulted in 
`r round(cells_by_station_refstations$ncp[cells_by_station_refstations$StationName == "Darwin"], 1)`% of cells using data from Darwin, 
`r round(cells_by_station_refstations$ncp[cells_by_station_refstations$StationName == "Kangaroo Island"], 1)`% using data from Kangaroo Island, 
`r round(cells_by_station_refstations$ncp[cells_by_station_refstations$StationName == "Maria Island"], 1)`% using data from Maria Island, 
`r round(cells_by_station_refstations$ncp[cells_by_station_refstations$StationName == "North Stradbroke Island"], 1)`% from North Stradbroke Island, 
`r round(cells_by_station_refstations$ncp[cells_by_station_refstations$StationName == "Port Hacking"], 1)`% using data from Port Hacking, 
`r round(cells_by_station_refstations$ncp[cells_by_station_refstations$StationName == "Rottnest Island"], 1)`% using data from Rottnest Island, and 
`r round(cells_by_station_refstations$ncp[cells_by_station_refstations$StationName == "Yongala"], 1)`% using data from Yongala. 

There were 
`r nrow(outfall_locations[outfall_locations$state == "SAU", ])` outfall stations in South Australia, 
`r nrow(outfall_locations[outfall_locations$state == "WAS", ])` in southern Western Australia, 
`r nrow(outfall_locations[outfall_locations$state == "WAN", ])` in northern Western Australia, 
`r nrow(outfall_locations[outfall_locations$state == "NSW", ])` in southern New South Wales, 
`r nrow(outfall_locations[outfall_locations$state == "QLD", ])` in Queensland, 
`r nrow(outfall_locations[outfall_locations$state == "VIC", ])` in Victoria, 
`r nrow(outfall_locations[outfall_locations$state == "TAS", ])` in Tasmania, and
`r nrow(outfall_locations[outfall_locations$state == "NTE", ])` in the Northern Territory.

Matching cells to outfall stations within 48 km resulted in 
`r round(stations_by_cell_outfalls$ncp[stations_by_cell_outfalls$ns == 1],1)`% of cells using data from one outfall site,
`r round(stations_by_cell_outfalls$ncp[stations_by_cell_outfalls$ns %in% c(2,3)],1)`% of cells using data from two or three sites,
`r round(stations_by_cell_outfalls$ncp[stations_by_cell_outfalls$ns %in% c(4,5)],1)`% of cells using data from three or four sites,
`r round(stations_by_cell_outfalls$ncp[stations_by_cell_outfalls$ns %in% c(6:8)],1)`% of cells using data from six to eight sites, and 
`r round(stations_by_cell_outfalls$ncp[stations_by_cell_outfalls$ns %in% c(9:12)],1)`% of cells using data from nine to twelve sites.
The remaining 
`r round(100-sum(stations_by_cell_outfalls$ncp),1)`% of cells did not use data from any outfall sites. 

```{r fig-example-N-input}
#| eval: true
#| fig-cap: An example of the daily nitrate (top) and ammonium (bottom) input and data for a single cell in New South Wales. Black lines show the input driving the growth model (constructed curve), blue dots show data from the nearest reference station (Port Hacking) and orange dots show distance-weighted data from the three nearest outfall points (Bateman’s Bay, Bermagui, and Bombo). 

BARRA_C2_cell_nos <- file.path(envi_data, "BARRA_C2_cell_nos.qs") %>% qs::qread()
ind <- which(BARRA_C2_cell_nos == 753076)

combined_data <- envi_data %>% 
  list.files(full.names = T) %>% 
  str_subset("combined_N_data") %>% 
  purrr::map_dfr(arrow::read_parquet) %>% 
  filter(cell_no == 753076) %>% 
  mutate(data_source = as.factor(data_source),
         indicator = factor(indicator, levels = c("nitrate_nitrite", "ammonia")),
         value = value %>% set_units("mg m-3") %>% set_units("umol L-1"))

cell_input_eg <- cell_input_all %>% 
  filter(cell_no == 753076) %>% 
  select(yday, Ni_input, Am_input) %>% 
  pivot_longer(names_to = "indicator", values_to = "value", cols = c(Ni_input, Am_input)) %>% 
  mutate(indicator = factor(indicator, 
                            levels = c("Ni_input", "Am_input"), 
                            labels = c("nitrate_nitrite", "ammonia")),
         value = value %>% set_units("mg m-3") %>% set_units("umol L-1"))

# combined_mean_sd <- envi_data %>% 
#   list.files(full.names = T) %>% 
#   str_subset("combined_N_mean_sd") %>% 
#   purrr::map_dfr(arrow::read_parquet) %>% 
#   filter(cell_no == 753076)
# 
# combined_parameters <- envi_data %>% 
#   list.files(full.names = T) %>% 
#   str_subset("combined_N_parameters") %>% 
#   purrr::map_dfr(arrow::read_parquet) %>% 
#   filter(cell_no == 753076)
  
ggplot(combined_data, 
       aes(x = yday, y = value, fill = as.factor(data_source), colour = as.factor(data_source), size = weight)) +
  geom_point(shape = 21, alpha = 0.35) +
  facet_wrap(facets = vars(indicator), nrow = 2) +
  geom_line(data = cell_input_eg, aes(x = yday, y = value), inherit.aes = F) +
  scale_fill_manual(values = c("orange", "steelblue")) +
  scale_colour_manual(values = c("orange", "steelblue")) +
  # scale_y_continuous(limits = c(0, 3.5), breaks = seq(0, 4, 0.5)) +
  env_plot +
  theme(strip.text = element_blank(), aspect.ratio = 0.55) +
  labs(x = "Day of the year", y = nitrogen_lab)
```

```{r Ni input data}
#| eval: true
#| cache: false

Ni_input <- cell_input_all %>% 
  select(cell_no, state, yday, Ni_input)

(Ni_means <- Ni_input %>% 
  group_by(state) %>% 
  reframe(mean = mean(Ni_input, na.rm = T),
          max = max(Ni_input, na.rm = T),
          min = min(Ni_input, na.rm = T)) %>% 
  mutate(variable = "Ni_input"))
```

Mean daily water velocity ranged from 
`r round(min(Ni_means$mean), 2)` in `r as.character(Ni_means$state[Ni_means$mean == min(Ni_means$mean)])` to 
`r round(max(Ni_means$mean), 2)` in `r as.character(Ni_means$state[Ni_means$mean == max(Ni_means$mean)])`,
with a mean across states of `r round(mean(Ni_means$mean), 2)`. 

```{r fig-Ni-input}
#| eval: true
#| cache: false
#| fig-cap: |
#|   Summary of nitrate concentration values across the year within each state, averaged across 2019-2023. Bold lines show the mean across all cells within each state while ribbons show the standard deviation between cells.

Ni_input_state <- Ni_input %>%
  group_by(state, yday) %>% 
  reframe(value = mean(Ni_input, na.rm = T),
          sd = sd(Ni_input, na.rm = T))

(
  p_input_state_av <- Ni_input_state %>% 
    ggplot(aes(x = yday, y = value, ymin = value-sd, ymax = value+sd, colour = state, fill = state)) +
    geom_line(linewidth = 0.75) +
    geom_ribbon(alpha = 0.35, linewidth = 0.25) +
    facet_wrap(facets = vars(state), ncol = 2) +
    # scale_y_continuous(breaks = seq(0, 40, 5), limits = c(10, 35)) +
    scale_colour_manual(values = states_pal) +
    scale_fill_manual(values = states_pal) +
    env_plot +
    labs(x = "Day of the year", y = nitrate_lab)
)
```

```{r Am input data}
#| eval: true
#| cache: false

Am_input <- cell_input_all %>% 
  select(cell_no, state, yday, Am_input)

(Am_means <- Am_input %>% 
  group_by(state) %>% 
  reframe(mean = mean(Am_input, na.rm = T),
          max = max(Am_input, na.rm = T),
          min = min(Am_input, na.rm = T)) %>% 
  mutate(variable = "Am_input"))
```

Mean daily water velocity ranged from 
`r round(min(Am_means$mean), 2)` in `r as.character(Am_means$state[Am_means$mean == min(Am_means$mean)])` to 
`r round(max(Am_means$mean), 2)` in `r as.character(Am_means$state[Am_means$mean == max(Am_means$mean)])`,
with a mean across states of `r round(mean(Am_means$mean), 2)`. 

```{r fig-Am-input}
#| eval: true
#| cache: false
#| fig-cap: |
#|   Summary of ammonium concentration values across the year within each state, averaged across 2019-2023. Bold lines show the mean across all cells within each state while ribbons show the standard deviation between cells.

Am_input_state <- Am_input %>%
  group_by(state, yday) %>% 
  reframe(value = mean(Am_input, na.rm = T),
          sd = sd(Am_input, na.rm = T))

p_input_state_av %+% Am_input_state +
  # scale_y_continuous(breaks = seq(30, 40, 0.5), limits = c(34, 36.5)) +
  labs(x = "Day of the year", y = ammonium_lab)
```

## Sensitivity conditions

```{r sens-conditions}
#| eval: true
#| cache: false

library(patchwork)

culture_depths <- tar_read(culture_depths, store = spec_store)
factors <- tar_read(factors, store = spec_store)

state_input_timeseries <- envi_data %>% 
  list.files(full.names = T) %>% 
  str_subset("state_input_timeseries") %>% 
  purrr::map_dfr(arrow::read_parquet) %>% 
  filter(yday <= 365)

sens_cond <- file.path(spec_data, "sensitivity_conditions.parquet") %>% 
  arrow::read_parquet()
```

Culture depths tested were `r culture_depths` m below the surface, with a 

Sensitivity runs started on the first of every month to cover annual variability, for 12 runs per year, with the same mean annual conditions looping from December to January. Sensitivity runs were conducted at culture depths of 0.5 m, 2 m, 3.5 m, and 5 m below the surface with a light attenuation coefficient of 0.65 (clear water). The canopy was always 2.5 m in height, and the total water column depth was always 50 m.
Curves for temperature, surface irradiance, salinity, water velocity, nitrate and ammonium were constructed to capture “typical” conditions based on daily means of each state from 2019-2023. This resulted in two sensitivity curves for temperature and a single sensitivity curve for irradiance, water velocity, salinity, nitrate and ammonium (fig-sens-altogether). 

```{r fig-sens-altogether}
#| eval: true
#| cache: false
#| fig-cap: |
#|   Average temperatures in each state region (left, see Figure 2 for detail) and the two temperature curves used for sensitivity analysis (high and low, right).

p1 <- sens_cond %>% 
  dplyr::filter(t <= 365) %>% 
  ggplot(aes(x = t, y = T_input, linetype = as.factor(T_level))) +
  geom_line(linewidth = 0.75) +
  geom_line(data = state_input_timeseries, aes(x = yday, y = T_input_mean, colour = state), linetype = "dotted", inherit.aes = F) +
  scale_y_continuous(breaks = seq(10, 35, 5), limits = c(10, 33)) +
  labs(y = expression("Temperature ("*degree*"C)")) +
  prettyplot + 
  annotate("text", x = Inf, y = Inf, label = "A", hjust = 1.5, vjust = 1.5, size = 4, fontface = "plain") +
  theme(legend.position = "none", aspect.ratio = 0.65, 
        axis.text.x = element_blank(), axis.ticks.x = element_blank(), axis.title.x = element_blank())

p2 <- sens_cond %>% 
  dplyr::filter(t <= 365 & T_level == 1) %>% 
  ggplot(aes(x = t, y = I_input)) +
  geom_line(linewidth = 0.75) +
  geom_line(data = state_input_timeseries, aes(x = yday, y = I_input_mean, colour = state), linetype = "dotted", inherit.aes = F) +
  scale_y_continuous(breaks = seq(0, 700, 100), limits = c(0, 610)) +
  labs(y = expression("Irradiance ("*mu*"mol m"^-2*" s"^-1*")")) +
  prettyplot + 
  annotate("text", x = Inf, y = Inf, label = "B", hjust = 1.5, vjust = 1.5, size = 4, fontface = "plain") +
  theme(legend.position = "none", aspect.ratio = 0.65, 
        axis.text.x = element_blank(), axis.ticks.x = element_blank(), axis.title.x = element_blank())

p3 <- sens_cond %>% 
  dplyr::filter(t <= 365 & T_level == 1) %>% 
  ggplot(aes(x = t, y = U_input)) +
  geom_line(linewidth = 0.75) +
  geom_line(data = state_input_timeseries, aes(x = yday, y = UV_input_mean, colour = state), linetype = "dotted", inherit.aes = F) +
  scale_y_continuous(breaks = seq(0, 0.6, 0.1), limits = c(0, 0.55)) +
  labs(y = expression("Water velocity (m"^-1*" s"^-1*")")) +
  prettyplot + 
  annotate("text", x = Inf, y = Inf, label = "C", hjust = 1.5, vjust = 1.5, size = 4, fontface = "plain") +
  theme(legend.position = "none", aspect.ratio = 0.65, 
        axis.text.x = element_blank(), axis.ticks.x = element_blank(), axis.title.x = element_blank())

p4 <- sens_cond %>% 
  dplyr::filter(t <= 365 & T_level == 1) %>% 
  ggplot(aes(x = t, y = S_input)) +
  geom_line(linewidth = 0.75) +
  geom_line(data = state_input_timeseries, aes(x = yday, y = S_input_mean, colour = state), linetype = "dotted", inherit.aes = F) +
  scale_y_continuous(breaks = seq(33.1, 36.8, 0.3), limits = c(34, 36.1)) +
  labs(y = expression("Salinity (g L"^-1*")")) +
  prettyplot + 
  annotate("text", x = Inf, y = Inf, label = "D", hjust = 1.5, vjust = 1.5, size = 4, fontface = "plain") +
  theme(legend.position = "none", aspect.ratio = 0.65, 
        axis.text.x = element_blank(), axis.ticks.x = element_blank(), axis.title.x = element_blank())

pdata1 <- sens_cond %>% 
  dplyr::filter(t <= 365 & T_level == 1) %>% 
  mutate(Ni_input = set_units(Ni_input, "mg m-3") %>% set_units("umol L-1") %>% drop_units(),
         Am_input = set_units(Am_input, "mg m-3") %>% set_units("umol L-1") %>% drop_units())
pdata2 <- state_input_timeseries %>% 
  mutate(Ni_input_mean = set_units(Ni_input_mean, "mg m-3") %>% set_units("umol L-1") %>% drop_units(),
         Am_input_mean = set_units(Am_input_mean, "mg m-3") %>% set_units("umol L-1") %>% drop_units())

p5 <- pdata1 %>% 
  ggplot(aes(x = t, y = Ni_input)) +
  geom_line(linewidth = 0.75) +
  geom_line(data = pdata2, aes(x = yday, y = Ni_input_mean, colour = state), linetype = "dotted", inherit.aes = F) +
  scale_y_continuous(breaks = seq(0, 5, 0.5), limits = c(0, 3)) +
  labs(y = nitrate_lab) +
  prettyplot + 
  annotate("text", x = Inf, y = Inf, label = "E", hjust = 1.5, vjust = 1.5, size = 4, fontface = "plain") +
  theme(legend.position = "none", aspect.ratio = 0.65, axis.title.x = element_blank())

p6 <- pdata1 %>% 
  ggplot(aes(x = t, y = Am_input)) +
  geom_line(linewidth = 0.75) +
  geom_line(data = pdata2, aes(x = yday, y = Am_input_mean, colour = state), linetype = "dotted", inherit.aes = F) +
  scale_y_continuous(breaks = seq(0, 10, 0.25), limits = c(0, 1)) +
  labs(y = ammonium_lab) +
  prettyplot + 
  annotate("text", x = Inf, y = Inf, label = "F", hjust = 1.5, vjust = 1.5, size = 4, fontface = "plain") +
  theme(legend.position = "none", aspect.ratio = 0.65, axis.title.x = element_blank())

plot_grid(
  plot_grid(p1, p2, p3, p4, p5, p6, ncol = 2, align = "hv"),
  ggdraw() + draw_label("Day of the year", size = 12),
  ncol = 1,
  rel_heights = c(0.95, 0.05)
)
```

```{r Ilim-testing}
#| eval: false

d_top_PL <- c(0.5, 1, 1.5, 2, 2.5, 3) # tar_read(d_top_PL, store = runs_store)

Ilim_cell <- runs_data %>% 
  list.files(full.names = T) %>% 
  str_subset("Ilim_cell") %>% 
  purrr::map_dfr(arrow::read_parquet)

# Ilim_means <- Ilim_cell %>% 
#   group_by(state, depth, yday) %>% 
#   reframe(I_lim = mean(I_lim, na.rm = T))
# 
# ggplot(Ilim_means, aes(x = yday, y = I_lim, colour = depth)) +
#   geom_line(linewidth = 0.75) +
#   facet_wrap(facets = vars(state)) +
#   theme_classic()

Ilim_means <- Ilim_cell %>% 
  group_by(state, depth) %>% 
  reframe(I_sd = sd(I_lim, na.rm = T),
          I_lim = mean(I_lim, na.rm = T))

ggplot(Ilim_means, aes(x = state, y = I_lim, fill = depth, ymin = I_lim-I_sd, ymax = I_lim+I_sd)) +
  geom_col(colour = "black", position = position_dodge(), width = 0.95) +
  geom_errorbar(position = position_dodge(width = 0.95), width = 0.5) +
  theme_classic()
```

```{r Ilim_cell-at-single-depth}
#| eval: false

Ilim_cell <- Ilim_cell %>% 
  dplyr::filter(depth == "2.5") %>% 
  dplyr::select(-c(irradiance, depth, state)) %>% 
  merge(cell_coords, by = "cell_no")

# var_rast <- terra::rast(list(x = Ilim_cell$longitude, y = Ilim_cell$latitude, z = Ilim_cell$I_lim))
# crs(var_rast) <- "+proj=longlat +datum=WGS84 +ellps=WGS84 +towgs84=0,0,0"))
```

Extra stuff?

# Results 

A total of `r length(cells_omitted[[1]])` were omitted for being too shallow to accommodate the 5.5 m macroalgae canopy, leaving `r length(cells_included[[1]])` cells.

## Species parameterisation

```{r fig-all-responses}
#| eval: true
#| fig-cap: |
#|   all responses

library(gridExtra)
library(grid)

N_uptake <- file.path(spec_data, "N_uptake.parquet") %>% 
  read_parquet()
S_response <- file.path(spec_data, "S_response.parquet") %>% 
  read_parquet()
I_response <- file.path(spec_data, "I_response.parquet") %>% 
  read_parquet() %>% 
  mutate(depth_m = as.factor(depth_m))
T_response <- file.path(spec_data, "T_response.parquet") %>% 
  read_parquet() %>% 
  mutate(species = factor(species, levels = c("taxi", "arma")))

p1 <- ggplot(N_uptake, aes(x = uM, y = uptake_uM, linetype = form)) +
  geom_line(linewidth = 0.75) +
  labs(x = expression("Ambient concentration ("*mu*"M)"), y = expression("Uptake rate ("*mu*"mol gDW"^-1*"d"^-1*")")) +
  scale_y_continuous(breaks = seq(0, 3.5, 0.05), limits = c(0, 0.30)) +
  scale_x_continuous(breaks = seq(0, 30, 5), limits = c(0,25)) +
  prettyplot + 
  theme(aspect.ratio = 0.75)

p2 <- ggplot(T_response, aes(x = temp, y = Tlim, colour = species)) +
  geom_line(linewidth = 0.75) +
  labs(x = expression("Temperature ("*degree*"C)"), y = "Relative growth") +
  scale_y_continuous(breaks = seq(0, 1, 0.2), limits = c(0,1)) +
  scale_x_continuous(breaks = seq(0, 40, 5), limits = c(0,35)) +
  prettyplot + 
  theme(aspect.ratio = 0.75)

p3 <- ggplot(I_response, aes(x = light, y = Ilim, linetype = as.factor(depth_m))) +
  geom_line(linewidth = 0.75) +
  facet_wrap(facets = vars(species)) +
  labs(x = expression("Surface light (PPFD)"), y = "Relative growth") +
  scale_y_continuous(breaks = seq(0, 1, 0.2), limits = c(0,1)) +
  scale_linetype_manual(values = c("solid", "dashed", "dotdash", "dotted")) +
  prettyplot + 
  theme(aspect.ratio = 0.75, strip.text = element_blank())

p4 <- ggplot(S_response, aes(x = sali, y = Slim)) +
  geom_line(linewidth = 0.75) +
  labs(x = expression("Salinity (g L"^-1*")"), y = "Relative growth") +
  scale_y_continuous(breaks = seq(0, 1, 0.2), limits = c(0,1)) +
  scale_x_continuous(breaks = seq(0, 60, 5), limits = c(19,45)) +
  prettyplot + 
  theme(aspect.ratio = 0.75)

p1 <- p1 + annotate("text", x = Inf, y = Inf, label = "A", hjust = 1.5, vjust = 1.5, size = 4)
p2 <- p2 + annotate("text", x = Inf, y = Inf, label = "B", hjust = 1.5, vjust = 1.5, size = 4)
p3 <- p3 + annotate("text", x = Inf, y = Inf, label = "C", hjust = 1.5, vjust = 1.5, size = 4)
p4 <- p4 + annotate("text", x = Inf, y = Inf, label = "D", hjust = 1.5, vjust = 1.5, size = 4)

plot_grid(
  plot_grid(p1, p2, p3, p4, ncol = 2, align = "hv"),
  ncol = 1,
  rel_heights = c(0.95, 0.05)  # Adjust these values to control spacing
)
```

### Parameter sensitivity

```{r sensitivity data}
#| eval: true

sens_arma <- file.path(spec_data, "sensitivity_arma_calc_combTU.parquet") %>% 
  read_parquet() #%>% 
  # dplyr::filter(!param %in% c("M_am", "C_am", "M_ot", "C_ot"))
exlude <- unique(sens_arma$param[abs(sens_arma$sens) < 0.01]) %>% as.character
sens_arma <- sens_arma %>% 
  dplyr::filter(!param %in% c(exlude, "M_am", "C_am", "M_ot", "C_ot"))
sens_arma$param <- droplevels(sens_arma$param)

levels(sens_arma$param)
exlude

# Add axis labels
param.labs.arma <- c(
  bquote("V"["Am"]), bquote("K"["Am"]),
  # bquote("M"["Ni"]),
  bquote("C"["Ni"]), bquote("Q"["min"]), bquote("Q"["max"]),
  # bquote("K"["c"]),
  bquote(mu), bquote("D"["ve"]),
  # bquote("a"["cs"]), bquote("I"["o"]),
  bquote("T"["opt"]), bquote("T"["min"]), bquote("T"["max"])
  # bquote("S"["opt"]), bquote("S"["min"]), bquote("S"["max"]), bquote("h"["a"]), bquote("h"["b"]), bquote("h"["c"]), bquote("h"["max"]), "DW:WW"
)

sens_taxi <- file.path(spec_data, "sensitivity_taxi_calc_combTU.parquet") %>% 
  read_parquet()
exlude <- unique(sens_taxi$param[abs(sens_taxi$sens) < 0.01]) %>% as.character
sens_taxi <- sens_taxi %>% 
  dplyr::filter(!param %in% c(exlude, "M_am", "C_am", "M_ot", "C_ot"))
sens_taxi$param <- droplevels(sens_taxi$param)

levels(sens_taxi$param)
exlude

# Add axis labels
param.labs.taxi <- c(
  bquote("V"["Am"]), bquote("K"["Am"]), 
  # bquote("M"["Ni"]), 
  bquote("C"["Ni"]), bquote("Q"["min"]), bquote("Q"["max"]), bquote("K"["c"]), bquote(mu), bquote("D"["ve"]), 
  # bquote("a"["cs"]), bquote("I"["o"]), 
  bquote("T"["opt"]), bquote("T"["min"]), bquote("T"["max"])
  # bquote("S"["opt"]), bquote("S"["min"]), bquote("S"["max"]), bquote("h"["a"]), bquote("h"["b"]), bquote("h"["c"]), bquote("h"["max"]), "DW:WW"
)
```

```{r fig-sensitivity-plot}
#| eval: true
#| fig-cap: |
#|   Relative impact of increasing each model parameter for _Asparagopsis armata_ (A) and _Asparagopsis taxiformis_ (B) by 10% on total N removed. Colours show the different culture depths of 0.5 m (orange), 2.5 m (blue) and 5.0 m (grey) below the surface. Error bars show variability (SD) across the 12 months of a typical year in each state. 

p1 <- ggplot(dplyr::filter(sens_arma, !is.na(sens)), 
       aes(x = rev(param), y = sens, fill = as.factor(cult_dep), ymin = sens-sd, ymax = sens+sd), 
       alpha = 0.75) +
  geom_col(position = position_dodge(), width = 0.95, colour = "black") +
  geom_errorbar(position = position_dodge(width = 0.95), width = 0.5) +
  coord_flip() +
  scale_y_continuous(limits = c(-1.1, 1.1), breaks = seq(-1.5, 1.5, 0.2)) +
  scale_x_discrete(limits = levels(sens_arma$param), labels = rev(param.labs.arma)) +
  scale_fill_manual(values = rev(depths_pal)) +
  labs(y = "Total N sensitivity", x = "Parameter name") +
  prettyplot +
  theme(axis.title.x = element_blank(), 
        axis.text.x = element_blank(), 
        axis.ticks.x = element_blank(),
        aspect.ratio = 0.75) + 
  annotate("text", x = Inf, y = Inf, label = "A", hjust = 1.5, vjust = 1.5, size = 4)

p2 <- ggplot(dplyr::filter(sens_taxi, !is.na(sens)), 
       aes(x = rev(param), y = sens, fill = as.factor(cult_dep), ymin = sens-sd, ymax = sens+sd), 
       alpha = 0.75) +
  geom_col(position = position_dodge(), width = 0.95, colour = "black") +
  geom_errorbar(position = position_dodge(width = 0.95), width = 0.5) +
  coord_flip() +
  scale_y_continuous(limits = c(-1.1, 1.1), breaks = seq(-1.5, 1.5, 0.2)) +
  scale_x_discrete(limits = levels(sens_arma$param), labels = rev(param.labs.taxi)) +
  scale_fill_manual(values = rev(depths_pal)) +
  labs(y = "Total N sensitivity", x = "Parameter name") +
  prettyplot + 
  theme(aspect.ratio = 0.75) + 
  annotate("text", x = Inf, y = Inf, label = "B", hjust = 1.5, vjust = 1.5, size = 4)

plot_grid(p1, p2, nrow = 2, align = "hv")
```

## Growth and bioremediation

The point of this section is to:

* Show the most common limiting factor - overall, between species, temporally, geographically
* Show growth/nitrogen removal - overall, between species, temporally, geographically

### Growth lims

```{r growth lim data}
#| eval: true

cell_growth_lims_fnms <- file.path(runs_data, "cell_growth") %>% 
  list.files(full.names = T) %>% 
  str_subset("cell_growth_lims_abbrev")

# This gives you a daily growth limiter for every cell
armata_lims <- map_dfr(cell_growth_lims_fnms, function(fnm) {
  arrow::read_parquet(fnm) %>%
    filter(species == "armata") %>%
    mutate(t = as.integer(t), t = case_when(t > 365 ~ t-365, T ~ t)) %>% 
    distinct(cell_no, species, t, lim)
})  %>% 
  merge(cell_coords, by = "cell_no")

# This gives you the dominant growth limiters per cell per month
dom_lims <- file.path(runs_data, "cell_growth_dom_lims.parquet") %>% 
  arrow::read_parquet() %>% 
  merge(cell_coords, by = "cell_no")

# cell_growth_rate <- file.path(runs_data, "cell_growth_rate.parquet") %>% 
#   arrow::read_parquet()
```

#### _Asparagopsis armata_

```{r armata lims stats}
#| eval: true
#| code-summary: See if there are any days where different lims are conflicting (different batches)

stat1 <- armata_lims %>% 
  filter(latitude > -32) %>% 
  group_by(species, cell_no, lim) %>% 
  reframe(perc_days = n()/365)
stat1_T <- mean(stat1$perc_days[stat1$lim == "T_lim"])
stat1_Q <- mean(stat1$perc_days[stat1$lim == "Q_lim"])
stat1_I <- mean(stat1$perc_days[stat1$lim == "I_lim"])

stat2 <- armata_lims %>% 
  filter(latitude <= -32) %>% 
  group_by(species, cell_no, lim) %>% 
  reframe(perc_days = n()/365)
stat2_T <- mean(stat2$perc_days[stat2$lim == "T_lim"])
stat2_Q <- mean(stat2$perc_days[stat2$lim == "Q_lim"])
stat2_I <- mean(stat2$perc_days[stat2$lim == "I_lim"])
```

For _A. armata_ grown at latitudes > -32$^\circ$, a mean of 
`r round(100*stat1_T,1)`% of growing days were limited by temperature, 
`r round(100*stat1_Q,1)`% were limited by nutrients, and 
`r round(100*stat1_I,1)`% were limited by light. 

For _A. armata_ grown at latitudes < -32$^\circ$, a mean of 
`r round(100*stat2_T,1)`% of growing days were limited by temperature, 
`r round(100*stat2_Q,1)`% were limited by nutrients, and 
`r round(100*stat2_I,1)`% were limited by light. 

```{r fig-growth-lim-armata-hist}
(
  p_dom_xy <- armata_lims %>% 
    filter(species == "armata") %>% 
    filter(latitude <= -30) %>% 
    # slice_sample(n = 3000) %>%
    mutate(month = parse_date_time(t, orders = "j"),
           month = factor(month(month), labels = month.name)) %>% 
    ggplot(aes(y = latitude, fill = lim, colour = lim)) +
    geom_histogram(position = "stack", alpha = 0.75) +
    facet_wrap(facets = vars(month)) +
    theme_classic() +
    scale_fill_manual(values = lims_pal) +
    scale_colour_manual(values = lims_pal) +
    labs(x = expression("Cells"*%*%*"number of days"), y = "Latitude")
)
```


```{r fig-growth-lim-armata}
#| eval: true

p_dom <- filter(dom_lims, species == "armata" & start == 1) %>% 
  ggplot(aes(x = longitude, y = latitude, fill = lim)) +
  geom_tile() +
  geom_sf(data = ozmap_data(data = "states"), inherit.aes = F) +
  coord_sf(xlim = c(bounds[["AUS"]]["lonmin"], bounds[["AUS"]]["lonmax"]),
           ylim = c(bounds[["AUS"]]["latmin"], bounds[["AUS"]]["latmax"])) +
  scale_x_continuous(breaks = seq(100, 160, 5)) +
  scale_y_continuous(breaks = seq(-5, -45, -5)) +
  scale_color_brewer(palette = "Dark2") +
  scale_fill_manual(values = lims_pal) +
  labs(x = "Longitude", y = "Latitude") +
  prettyplot + 
  theme(legend.position = "top")

p_dom
p_dom %+% filter(dom_lims, species == "armata" & start == 2)
p_dom %+% filter(dom_lims, species == "armata" & start == 3)
p_dom %+% filter(dom_lims, species == "armata" & start == 4)
p_dom %+% filter(dom_lims, species == "armata" & start == 5)
p_dom %+% filter(dom_lims, species == "armata" & start == 6)
p_dom %+% filter(dom_lims, species == "armata" & start == 7)
p_dom %+% filter(dom_lims, species == "armata" & start == 8)
p_dom %+% filter(dom_lims, species == "armata" & start == 9)
p_dom %+% filter(dom_lims, species == "armata" & start == 10)
p_dom %+% filter(dom_lims, species == "armata" & start == 11)
p_dom %+% filter(dom_lims, species == "armata" & start == 12)
```

#### _Asparagopsis taxiformis_

```{r fig-growth-lim-taxiformis}
#| eval: true

p_dom <- filter(dom_lims, species == "taxiformis" & start == 1) %>% 
  ggplot(aes(x = longitude, y = latitude, fill = lim)) +
  geom_tile() +
  geom_sf(data = ozmap_data(data = "states"), inherit.aes = F) +
  coord_sf(xlim = c(bounds[["AUS"]]["lonmin"], bounds[["AUS"]]["lonmax"]),
           ylim = c(bounds[["AUS"]]["latmin"], bounds[["AUS"]]["latmax"])) +
  scale_x_continuous(breaks = seq(100, 160, 5)) +
  scale_y_continuous(breaks = seq(-5, -45, -5)) +
  scale_color_brewer(palette = "Dark2") +
  scale_fill_manual(values = lims_pal) +
  labs(x = "Longitude", y = "Latitude") +
  prettyplot + 
  theme(legend.position = "top")

p_dom
p_dom %+% filter(dom_lims, species == "taxiformis" & start == 2)
p_dom %+% filter(dom_lims, species == "taxiformis" & start == 3)
p_dom %+% filter(dom_lims, species == "taxiformis" & start == 4)
p_dom %+% filter(dom_lims, species == "taxiformis" & start == 5)
p_dom %+% filter(dom_lims, species == "taxiformis" & start == 6)
p_dom %+% filter(dom_lims, species == "taxiformis" & start == 7)
p_dom %+% filter(dom_lims, species == "taxiformis" & start == 8)
p_dom %+% filter(dom_lims, species == "taxiformis" & start == 9)
p_dom %+% filter(dom_lims, species == "taxiformis" & start == 10)
p_dom %+% filter(dom_lims, species == "taxiformis" & start == 11)
p_dom %+% filter(dom_lims, species == "taxiformis" & start == 12)
```

### Growth ends

```{r growth end data}
#| eval: true

cell_growth_end <- file.path(runs_data, "cell_growth_end.parquet") %>% 
  read_parquet() %>% 
  mutate(t_start = as.integer(lubridate::month(parse_date_time(x = t_start, orders = "j"))+1),
         t_start = case_when(t_start == 13 ~ 1, t_start == 14 ~ 2, T ~ t_start)) %>% 
  mutate(species = as.factor(species)) %>% 
  merge(cell_coords, by = "cell_no")
```

## Growth scenarios

```{r scen lim data}
#| eval: false

cell_growth_lims <- find_read(file.path(runs_data, "cell_growth"), "growth_lims") %>% 
  dplyr::select(cell_no, start, species, lims) %>% 
  mutate(lims = as.factor(lims),
         species = as.factor(species)) %>% 
  merge(cell_coords, by = "cell_no")
```

When compared with the base scenario, all species were able to remove the most nitrogen in the fish farm scenario, almost certainly because of the increase in ammonium concentration from (very low) background levels (Figure 42). *Asparagopsis armata* and *A. taxiformis* consistently removed more nitrogen in the fish farm scenario.

Differences in nitrogen removed by *Asparagopsis* were minimal in the estuary mouth scenario across all states (ranging from -2.1% to +2.5% of base). Similarly, in the shallow bay scenario, *Asparagopsis* was able to remove from 0.4 to 9.5$\times$ more nitrogen. The deep water scenario showed mostly negative but relatively small changes from the base removal, with changes ranging from -36% removal by *A. taxiformis* in New South Wales to +16.2% removed by *A. armata* in northern Western Australia. 

Figure 42: Nitrogen removal for different scenarios compared to the base scenario. Colours show the different scenarios as fish farm (purple), deep water (green), estuary mouth (orange) and shallow bay (pink).

# Acknowledgments

# Supplementary material

# References