---
title: "Results: Supplemented growth and bioremediation"
author: Tormey Reimer
date: last-modified
---

```{r setup}
#| eval: true
#| include: false
#| cache: false

# You need quarto to generate this markdown - install here: https://quarto.org/docs/get-started/
# This markdown uses TinyTex for PDF rendering - install with tinytex::install_tinytex()
# This markdown uses the acronyms extension - install by entering quarto add rchaput/acronyms@master into the terminal

library(stringr)
library(units)
library(targets)
library(tarchetypes)
library(macrogrow)
library(ozmaps)
library(arrow)
library(gganimate)
library(magrittr)
library(ggplot2)
library(dplyr)
library(terra)
library(tidyterra)
library(here)
library(lubridate)
library(tidyr)
library(cowplot)
library(purrr)
library(patchwork)
library(conflicted)
library(gridExtra)
library(grid)

conflicts_prefer(dplyr::select, dplyr::filter)

source(here("R_scripts/00_functions.R"))
source(here("R_scripts/00_global_displays.R"))
```

```{r cell-coords}
#| eval: true
#| code-summary: |
#|   Load in all the BARRA-R2 cells being used, with their coordinates.

cell_coords <- find_read(envi_data, "R2_cell_coords") %>% 
  mutate(state = factor(state, levels = states_ord)) %>% 
  dplyr::select(-layer)
cells_included <- find_read(runs_data, "include")
cells_omitted <- find_read(runs_data, "omit")

totals <- cell_coords %>% 
  filter(!cell_no %in% cells_omitted) %>% 
  group_by(state) %>% 
  reframe(cells = n())
```

```{r normal-growth-end-data}
#| code-summary: |
#|   Get cell growth in ambient (normal) conditions

norm_growth_end <- file.path(runs_data, "cell_growth_end.parquet") %>% 
  read_parquet(mmap = T) %>% 
  merge(cell_coords, by = "cell_no")

# Total N removed at each harvest
norm_growth_end <- norm_growth_end %>% 
  mutate(
    value = TN_end - TN_start,
    value = case_when(value == 0 ~ NA, T ~ value),
    month = as.factor(month(parse_date_time(start, orders = "j"))),
    species = factor(species, levels = c("armata", "taxiformis"), labels = c("A. armata", "A. taxiformis"))
  )

norm_growth_total <- norm_growth_end %>% 
  mutate(
    value_na = case_when(value <= 0 ~ NA, T ~ value),
    value_0s = case_when(value <= 0 ~ 0, T ~ value)
  ) %>% 
  group_by(state, species, cell_no, latitude, longitude) %>% 
  reframe(
    total = sumna(value),            # good growth months may be outweighed by bad growth months
    total_na = sumna(value_na),      # bad growth months are removed
    total_na = case_when(total_na == 0 ~ NA, T ~ total_na),
    total_0s = sumna(value_0s)       # bad growth months count as zeros
  )
```

```{r supp-growth-end-data}
#| code-summary: |
#|   Get cell growth in nitrogen supplemented conditions

supp_growth_end_raw <- file.path(runs_data, "cell_growth_supp_end.parquet") %>% 
  read_parquet(mmap = T) %>% 
  merge(cell_coords, by = "cell_no")

supp_growth_end <- supp_growth_end_raw %>% 
  mutate(
    value = TN_end - TN_start,
    value = case_when(value == 0 ~ NA, T ~ value),
    month = as.factor(month(parse_date_time(start, orders = "j"))),
    species = factor(species, levels = c("armata", "taxiformis"), labels = c("A. armata", "A. taxiformis"))
  )

supp_growth_total <- supp_growth_end %>% 
  mutate(
    value_na = case_when(value <= 0 ~ NA, T ~ value),
    value_0s = case_when(value <= 0 ~ 0, T ~ value)
  ) %>% 
  group_by(state, species, cell_no, latitude, longitude) %>% 
  reframe(
    total = sumna(value),            # good growth months may be outweighed by bad growth months
    total_na = sumna(value_na),      # bad growth months are removed
    total_na = case_when(total_na == 0 ~ NA, T ~ total_na),
    total_0s = sumna(value_0s)       # bad growth months count as zeros
  )
```

```{r all-growth-end-data}
#| code-summary: |
#|   Join them together for comparison

supplement_change <- left_join(
  norm_growth_end %>% 
    select(-c("TN_start", "TN_end", "start", "value")) %>% 
    rename(norm_success = success, norm_rem = TN_rem), 
  supp_growth_end %>% 
    select(-c("TN_start", "TN_end", "start", "value")) %>% 
    rename(supp_success = success, supp_rem = TN_rem), 
  by = c("state", "cell_no", "longitude", "latitude", "species", "month")
) %>% 
  relocate(state, .before = cell_no) %>% 
  relocate(species, .before = state) %>% 
  relocate(longitude, .after = cell_no) %>% 
  relocate(latitude, .after = longitude) %>% 
  relocate(month, .after = latitude)

supplement_change_total <- left_join(
  norm_growth_total %>% 
    rename(norm_total = total, norm_total_na = total_na, norm_total_0s = total_0s), 
  supp_growth_total %>% 
    rename(supp_total = total, supp_total_na = total_na, supp_total_0s = total_0s), 
  by = c("state", "cell_no", "longitude", "latitude", "species")
)
```

# Supplemented success change

Provide summaries of the changes in success (ie whether the modelled seaweed was able to successfully growth at all within a cell) between normal and supplemented growth.

```{r supp-growth-success-change-1}
#| code-summary: |
#|   How many/what proportion of cells became successful in each month (and overall) because of supplementation

# Which cells changed in success (per month)
norm_success <- supplement_change %>% 
  filter(norm_success) %>% 
  group_by(species, month) %>% 
  reframe(
    norm_success_cells = n(),
    perc_total_cells = 100*n()/11748
  )
success_change <- supplement_change %>% 
  filter(supp_success != norm_success) %>% 
  group_by(species, month) %>% 
  reframe(success_change_cells = n())

norm_success %>% 
  left_join(success_change, by = c("species", "month")) %>% 
  mutate(
    percdiff_normal_change = 100*success_change_cells/norm_success_cells,
    perc_total_change = 100*success_change_cells/11748
  ) %>% 
  select(-norm_success_cells, -success_change_cells) %>% 
  print(n = 24)

# Which cells changed in success (overall)
norm_success <- supplement_change %>% 
  filter(norm_success) %>% 
  group_by(species) %>% 
  reframe(
    norm_success_cells = n(),
    perc_total_cells = 100*n()/(12*11748)
  )
success_change <- supplement_change %>% 
  filter(supp_success != norm_success) %>% 
  group_by(species) %>% 
  reframe(success_change_cells = n())

norm_success %>% 
  left_join(success_change, by = c("species")) %>% 
  mutate(
    percdiff_normal_ch = 100*success_change_cells/norm_success_cells,
    perc_total_ch = 100*success_change_cells/(11748*12)
  ) %>% 
  select(-norm_success_cells, -success_change_cells)
```

```{r area-success-change}
#| eval: false

library(terra)

rdata_norm <- norm_growth_end %>% 
  select(cell_no, species, success, longitude, latitude, month) %>% 
  group_by(species, cell_no, longitude, latitude) %>% 
  reframe(
    success = as.numeric(any(success)),
    success = case_when(success == 0 ~ NA, T ~ 1)
  ) %>% 
  filter(species != "A. armata")

rdata_supp <- supp_growth_end %>% 
  select(cell_no, species, success, longitude, latitude, month) %>% 
  group_by(species, cell_no, longitude, latitude) %>% 
  reframe(
    success = as.numeric(any(success)),
    success = case_when(success == 0 ~ NA, T ~ 1)
  ) %>% 
  filter(species != "A. armata")

r_template <- rast(
  extent = ext(c(min(rdata_norm$longitude), max(rdata_norm$longitude), 
                 min(rdata_norm$latitude), max(rdata_norm$latitude))), 
  resolution = 0.12
)

r <- rasterize(rdata_norm[, c("longitude", "latitude")], r_template, values = rdata_norm$success)
r_area <- cellSize(r, unit = "km")
r_area <- mask(r_area, r)
area_norm_success <- global(r_area, fun = sumna)[[1]]

r <- rasterize(rdata_supp[, c("longitude", "latitude")], r_template, values = rdata_supp$success)
r_area <- cellSize(r, unit = "km")
r_area <- mask(r_area, r)
area_supp_success <- global(r_area, fun = sumna)[[1]]

area_supp_success-area_norm_success
100*(area_supp_success-area_norm_success)/area_norm_success
```

```{r supp-growth-success-change-2}
#| fig-height: 7
#| code-summary: |
#|   Find out where cells changed from failure to success under supplemented conditions

supplement_change %>% 
  mutate(success_change = case_when(supp_success != norm_success ~ T, T ~ F)) %>% 
  filter(success_change & species == "A. armata") %>% 
  ggplot(aes(x = longitude, y = latitude, fill = success_change)) +
  geom_tile() +
  geom_sf(data = ozmap_data(data = "states"), inherit.aes = F) +
  coord_sf(
    xlim = c(111.5, 155),
    ylim = c(-44.5, -8.5),
    expand = F
  ) +
  theme_void() + theme(legend.position = "none") +
  facet_wrap(facets = vars(month), nrow = 3) +
  ggtitle("Cells where A. armata became successful in each month under supplementation")

supplement_change %>% 
  mutate(success_change = case_when(supp_success != norm_success ~ T, T ~ F)) %>% 
  filter(success_change & species == "A. taxiformis") %>% 
  ggplot(aes(x = longitude, y = latitude, fill = success_change)) +
  geom_tile() +
  geom_sf(data = ozmap_data(data = "states"), inherit.aes = F) +
  coord_sf(
    xlim = c(111.5, 155),
    ylim = c(-44.5, -8.5),
    expand = F
  ) +
  theme_void() + theme(legend.position = "none") +
  facet_wrap(facets = vars(month), nrow = 3) +
  ggtitle("Cells where A. taxiformis became successful in each month under supplementation")
```

# Supplemented nitrogen removal 

```{r total-removed-with-supplementation}
#| code-summary: |
#|   Differences in nitrogen removed because of supplementation

rem_diff_total <- supplement_change %>% 
  select(-contains("success")) %>% 
  mutate(
    norm_rem = case_when(norm_rem <= 0 ~ NA, T ~ norm_rem),
    supp_rem = case_when(supp_rem <= 0 ~ NA, T ~ supp_rem),
    rem_diff = supp_rem - norm_rem
  )

# Show difference in nitrogen removed per month (all cells)
rem_diff_total %>%
  group_by(species, month) %>% 
  reframe(
    norm_rem = sumna(norm_rem),
    supp_rem = sumna(supp_rem),
    rem_diff = sumna(rem_diff)/norm_rem,
    perc_rem_diff = 100*rem_diff
  ) %>% 
  print(n = 24)

# Show difference in nitrogen removed (overall)
supplement_change_total %>%
  mutate(
    rem_diff = supp_total_0s-norm_total_0s,
    rem_diff = case_when(rem_diff <= 0 ~ NA, T ~ rem_diff)
  ) %>% 
  group_by(species) %>% 
  reframe(
    rem_diff = sumna(rem_diff),
    rem_diff_kg = rem_diff %>% set_units("mg") %>% set_units("kg"),
    norm_total_na = sumna(norm_total_na),
    supp_total_na = sumna(supp_total_na),
    prop_rem_diff = rem_diff/norm_total_na,
    perc_rem_diff = 100*prop_rem_diff
  )
```

```{r fig-supp-growth-end-map-totals}
#| fig-cap: Total nitrogen removed by _A. armata_ (top) and _A. taxiformis (bottom) across the entire year under supplemented conditions.

supp_growth_total %>%
  ggplot(aes(x = longitude, y = latitude, fill = total_na)) +
  geom_tile() +
  geom_sf(data = ozmap_data(data = "states"), inherit.aes = F) +
  coord_sf(xlim = c(111.5, 155), ylim = c(-44.5, -8.5), expand = F) +
  scale_x_continuous(breaks = seq(100, 160, 10)) +
  scale_y_continuous(breaks = seq(-2.5, -45, -5)) +
  scale_fill_viridis_c(
    limits = c(0, 10000),
    breaks = seq(0, 10000, 1500),
    na.value = "grey",
    guide = guide_colorbar(
      title = expression("Total N removed (mg m"^-2 * ")"),
      barheight = 1,
      barwidth = 20,
      title.position = "top",
      title.hjust = 0.5
    )
  ) +
  labs(x = "Longitude", y = "Latitude") +
  prettyplot() +
  theme(legend.position = "top", strip.text = element_blank(), strip.background = element_blank()) +
  facet_grid(rows = vars(species))
```

## Comparing normal and supplemented removal

```{r compare-growth-map-armata}
#| fig-cap: Total nitrogen removed by _A. armata_ (top) and _A. taxiformis (bottom) across the entire year under normal conditions (left) and supplemented conditions (right).

all_growth_total <- rbind(
  norm_growth_total %>% mutate(scenario = "norm_conditions"),
  supp_growth_total %>% mutate(scenario = "supp_conditions")
)

# hist(all_growth_total$total_na[all_growth_total$scenario == "norm_conditions"])
# hist(all_growth_total$total_na[all_growth_total$scenario != "norm_conditions"])
minna(all_growth_total$total_na)
maxna(all_growth_total$total_na)

all_growth_total %>%
  ggplot(aes(x = longitude, y = latitude, fill = total_na)) +
  geom_tile() +
  geom_sf(data = ozmap_data(data = "states"), inherit.aes = F) +
  coord_sf(xlim = c(111.5, 155), ylim = c(-44.5, -8.5), expand = F) +
  scale_x_continuous(breaks = seq(100, 160, 10)) +
  scale_y_continuous(breaks = seq(-2.5, -45, -5)) +
  scale_fill_viridis_c(
    limits = c(1, 10000),
    labels = scales::label_number(),
    trans = "log10",
    na.value = "grey",
    guide = guide_colorbar(
      title = expression("Total N removed (mg m"^-2 * ")"),
      barheight = 1,
      barwidth = 20,
      title.position = "top",
      title.hjust = 0.5
    )
  ) +
  labs(x = "Longitude", y = "Latitude") +
  facet_grid(rows = vars(species), cols = vars(scenario)) +
  prettyplot() +
  theme(legend.position = "top", strip.text = element_blank(), strip.background = element_blank())
```

## Supplemented removed OF supplementation

Overall, the supplemented growth had an extra 10uM of nitrogen per day, or an extra 3650 uM per year. How much of that was actually removed by supplemented growth?

```{r total-removed-OF-supplementation}
#| eval: false

supplementation <- set_units(3650, "umol L-1")
supplementation <- supplementation %>% set_units("mg m-3") 
supplementation <- (supplementation * set_units(5.5, "m")) %>% drop_units()

rem_diff_ofsupp <- rem_diff_total %>% 
  select(-norm_rem, -rem_diff) %>% 
  group_by(species, state, cell_no, longitude, latitude) %>% 
  reframe(supp_rem = sumna(supp_rem)) %>% 
  mutate(
    supp_rem = supp_rem/supplementation,
    # supp_rem = case_when(supp_rem <= 0 ~ NA, T ~ supp_rem)
  )

rem_diff_ofsupp %>% 
  group_by(species) %>% 
  reframe(
    supp_rem_mean = 100*meanna(supp_rem),
    supp_rem_max = 100*maxna(supp_rem)
  )
```

# Monthly supplemented growth results

```{r supp-total-means}
#| code-summary: |
#|   Get mean nitrogen removed per each state/month

supp_growth_month_arma <- supp_growth_end %>% 
  filter(success & species == "A. armata") %>% 
  group_by(species, state, month) %>% 
  reframe(
    mean = meanna(value) %>% set_units("mg m-2"),
    sd = sdna(value) %>% set_units("mg m-2")
  ) 
  
supp_growth_month_arma %>% 
  select(-sd) %>% 
  pivot_wider(names_from = state, values_from = mean) %>% 
  arrange(month)

supp_growth_month_taxi <- supp_growth_end %>% 
  filter(success & species != "A. armata") %>% 
  group_by(species, state, month) %>% 
  reframe(
    mean = meanna(value) %>% set_units("mg m-2"),
    sd = sdna(value) %>% set_units("mg m-2")
  )
  
supp_growth_month_taxi %>% 
  select(-sd) %>% 
  pivot_wider(names_from = state, values_from = mean) %>% 
  arrange(month)
```

```{r fig-supp-growth-end}
#| eval: false

plot_data <- rbind(supp_growth_month_arma, supp_growth_month_taxi) %>% 
  mutate(
    mean = drop_units(mean),
    sd = drop_units(sd),
    mean = case_when(is.na(mean) ~ 0, T ~ mean),
    sd = case_when(sd > mean ~ mean, is.na(sd) ~ 0, T ~ sd),
    state = factor(state, levels = states_ord, labels = states_lng),
    month = as.integer(month)
  ) 
  
plot_data %>%
  ggplot(aes(
      x = month, y = mean, ymin = mean - sd, ymax = mean + sd,
      color = state, fill = state, linetype = species
  )) +
  geom_line(linewidth = 0.75) +
  geom_ribbon(alpha = 0.25, linewidth = 0.25) +
  facet_wrap(facets = vars(state), ncol = 2) +
  # scale_y_continuous(breaks = seq(0, 150, 25), limits = c(0, 175), expand = c(0,0)) +
  scale_color_manual(values = states_pal) +
  scale_fill_manual(values = states_pal) +
  env_plot() +
  labs(x = "Day of the year", y = expression("Total N removed (mg m"^-2 *")"))
```

